### 2012-09-08 paper.tex
Initial commit.

\documentclass{llncs}

\begin{document}

\end{document}

### 2012-09-11 paper.tex
Boilerplate.
Sections.
Dump.

@@
 \documentclass{llncs}
+
+\title{A Correct Compiler for Continuation Marks}
+
+\author{Kimball Germane \and Jay McCarthy}
+\institute{Brigham Young University}
 
@@
 
+\maketitle
+
+\begin{abstract}
+
+\end{abstract}
+
+\section{Introduction}
+
+\section{Related Work}
+
+\section{Continuation Marks}
+
+There are certain tools that are indispensable to some programmers that concern the
+behavior of their programs: debuggers, profilers, steppers, etc. Without these tools,
+these programmers cannot justify the adoption of a language, however compelling it might
+otherwise be. Traditionally, these tools are developed at the same level the 
+language is, privy to incidental implementation detail, precisely because that detail 
+enables these tools to function. This is problematic for at least two reasons. First, 
+it couples the implementation of the tool with the implementation of the language, which
+increases the cost to port to other platforms. If users become dependent upon these tools,
+it can stall the advancement of the language and the adoption of new language features.
+Second and more critical, it makes these tools unsound. For instance, debuggers typically
+examine programs which have been compiled without optimizations. In general, this means 
+that the debugged program has different behavior than the deployed program. This is 
+obviously undesirable.
+
+It is desirable to implement such tools at the same level as the language, removing
+dependency upon the implementation an instead relying on definitional and behavioral
+invariants. Continuation marks are a language-level feature that provide the information
+necessary for these tools to function. Furthermore, languages which require stack
+inspection to enforce security policies (\emph{Java}, \emph{C\#}) or support aspect
+oriented programming (\emph{aspectj}) can be defined in terms of a simpler language with
+continuation marks \cite{clements2004tail}.
+
+Continuation marks originated in PLT Scheme (now Racket \cite{plt-tr1}) as a stack 
+inspection mechanism. In fact, the \emph{Java} and \emph{C\#} languages rely on a similar 
+stack inspection to enforce security policies of which continuation marks can be seen as 
+a generalization. Surprisingly, continuation marks can be encoded in any language with 
+exception facilities \cite{pettyjohn2005continuations} which fact has led to their 
+experimental addition to Javascript \cite{clements2008implementing}.
+
+\section{$\lambda_{cm}$}
+
+\section{Transformation}
+
+\section{Testing}
+
+\section{Proof}
+
+\section{Future Work}
+
+\section{Conclusion}
+
 \end{document}

### 2012-09-11 paper.tex
Up to date.

@@
 these programmers cannot justify the adoption of a language, however compelling it might
-otherwise be. Traditionally, these tools are developed at the same level the 
-language is, privy to incidental implementation detail, precisely because that detail 
-enables these tools to function. This is problematic for at least two reasons. First, 
-it couples the implementation of the tool with the implementation of the language, which
-increases the cost to port to other platforms. If users become dependent upon these tools,
-it can stall the advancement of the language and the adoption of new language features.
-Second and more critical, it makes these tools unsound. For instance, debuggers typically
-examine programs which have been compiled without optimizations. In general, this means 
-that the debugged program has different behavior than the deployed program. This is 
-obviously undesirable.
+otherwise be. Traditionally, these tools are developed at the same level the language is,
+privy to incidental implementation detail, precisely because that detail enables these
+tools to function. This is problematic for at least two reasons. First, it couples the
+implementation of the tool with the implementation of the language, which increases the
+cost to port to other platforms. If users become dependent upon these tools, it can stall
+the advancement of the language and the adoption of new language features. Second and more
+critical, it makes these tools unsound. For instance, debuggers typically examine programs
+which have been compiled without optimizations. In general, this means that the debugged
+program has different behavior than the deployed program. This is obviously undesirable.
 
 It is desirable to implement such tools at the same level as the language, removing
-dependency upon the implementation an instead relying on definitional and behavioral
+dependency upon the implementation and instead relying on definitional and behavioral
 invariants. Continuation marks are a language-level feature that provide the information
@@
 exception facilities \cite{pettyjohn2005continuations} which fact has led to their 
-experimental addition to Javascript \cite{clements2008implementing}.
+experimental addition to \emph{JavaScript} \cite{clements2008implementing}.
 
@@
 
+\bibliographystyle{plain}
+\bibliography{paper}
+
 \end{document}

### 2012-09-20 paper.tex
A bit of a dump but new prose and ideas in there.

--- 2012-09-08 paper.tex
@@
 \documentclass{llncs}
+
+\usepackage{amsmath}
+\usepackage{slatex}
+
+\title{A Correct Compiler for Continuation Marks}
+
+\author{Kimball Germane \and Jay McCarthy}
+\institute{Brigham Young University}
+
+% Look at the many other papers on c marks to justify them (see mine for
+% example). Talk about how your transformation is a compiler and right
+% now no language but Racket has C-marks, so we could use this to get
+% them in that language. For Racket though, now they don't need to be
+% built in to the compiler if we don't want. In addition, it makes it
+% easier to do a type system.
+
+\newcommand{\cm}[0]{$\lambda_{cm}$}
+\newcommand{\lv}[0]{$\lambda_v$}
+\newcommand{\lc}[0]{$\lambda$-calculus}
+\newcommand{\wcm}[2]{(\mathrm{wcm}\,#1\,#2)}
+\newcommand{\ccm}[0]{(\mathrm{ccm})}
+\newcommand{\app}[2]{(#1\,#2)}
+\newcommand{\abs}[2]{(\lambda\,(#1)\,#2)}
+\newcommand{\hole}[0]{\bullet}
+\newcommand{\rr}[0]{\rightarrow}
+\newcommand{\lrrs}[0]{\rightarrow^{*}}
+\newcommand{\lvrr}[0]{\rightarrow_v}
+\newcommand{\lvrrs}[0]{\rightarrow_v^{*}}
+\newcommand{\cmrr}[0]{\rightarrow_{cm}}
+\newcommand{\cmrrs}[0]{\rightarrow_{cm}^{*}}
 
@@
 
+\maketitle
+
+\begin{abstract}
+Continuation marks, a programming language feature which generalizes stack inspection, do not yet enjoy a transformation to the plain $\lambda$-calculus.
+We present a CPS-like transformation from the call-by-value $\lambda$-calculus augmented with continuation marks to the pure call-by-value $\lambda$-calculus.
+We discuss how such a transformation would simplify the construction of compilers which treat continuation marks correctly.
+We document an iterative, feedback-based approach using Redex.
+We accompany Finally, we verify the transformation with a meaning-preservation theorem.
+\end{abstract}
+
+\section{Introduction}
+
+Language transformation is useful for a variety of purposes.
+It can be used to give meaning to the source language in terms of the target language [cite something], provide a characterization more amenable to analysis.
+Language transformation can also be used to \emph{desugar} a language into a simpler core.
+This
+Transforming has the following benefits:
+reduce core code (desugaring, simplifies implementations)
+give meaning to constructs (semantically, as well as, say, macros)
+local transformations - macros, for example
+global transformations - compilers of all kinds, for example
+
+talk about the common practice of analyzing an extension of the lambda calculus, and a subpractice of doing so via a transformation to the core.
+
+While continuation marks have had significant treatment, they do not currently enjoy a transformation to the plain $\lambda$-calculus.
+
+\section{Continuation Marks}
+
+Continuation marks generalize stack inspection, allowing a bevy of features that rely on some form of it to be defined at the language level.
+This includes algebraic steppers \cite{clements2001modeling}, security policies \cite{clements2004tail}, and aspects \cite{tucker2003pointcuts}.
+
+Few languages offer continuation marks.
+Their inception lay in Racket \cite{plt-tr1} and they have been added experimentally to \emph{JavaScript} \cite{clements2008implementing}, but their adoption by other languages has been nonexistent.
+
+Racket offers desugaring facilities to the user via a powerful macro system. This requires and is limited to local transformations limited to expression boundaries.
+
+A transformation-based compiler, such as a CPS compiler \cite{appel2007compiling}, effects a global transformation of the program.
+This effect discards original source or transforms it so radically that it is suitable only for the levels beneath the programmer's regular interaction.
+
+\section{\cm}
+
+Terms $e$ in \cm\ are defined inductively by
+\begin{align}
+e=\, &\wcm{e}{e}\\
+     &\ccm\\
+     &\app{e}{e}\\
+     &v\\
+     &x
+\end{align}
+where 
+\begin{equation}
+v=\abs{x}{e}
+\end{equation}
+describes the values of the \lc.
+
+The sole novelty of \cm\ above the plain \lc\ is the ability to annotate and observe the continuation.
+Our evaluation model of \cm, introduced by Pettyjohn et al.
+\cite{pettyjohn2005continuations}, utilizes the current evaluation context for that purpose.
+The mutually inductive valuation contexts $E$, $F$ are defined by
+\begin{align}
+E=\, &\wcm{v}{F}\\
+     &F\\
+F=\, &\app{E}{e}\\
+     &\app{v}{E}\\
+     &\wcm{E}{e}\\
+     &\hole
+\end{align}
+where $\hole$ denotes a ``hole'' in the evaluation context, the ultimate destination of the evaluation of the expression that previously resided there.
+
+The semantics of \cm\ are
+\begin{align}
+E[\app{\abs{x}{e}}{v}]  &\rr E[e[x\leftarrow v]]\\
+E[\wcm{v}{\wcm{v'}{e}}] &\rr E[\wcm{v'}{e}]\\
+E[\wcm{v}{v'}]          &\rr E[v']\\
+E[\ccm]                 &\rr E[\chi(E,\abs{x}{\abs{y}{y}})]
+\end{align}
+where the metafunction $\chi$ is defined by
+\begin{align}
+\chi(\wcm{v}{F},vs) &= \chi(F,\abs{z}{\app{\app{z}{v}}{vs}})\\
+\chi(\app{E}{e},vs) &= \chi(E,vs)\\
+\chi(\app{v}{E},vs) &= \chi(E,vs)\\
+\chi(\wcm{E}{e},vs) &= \chi(E,vs)\\
+\chi(\hole,vs)      &= vs
+\end{align}.
+
+\section{Transformation}
+
+\section{Testing}
+
+A pragmatic approach to the discovery of a correct transformation involves consistent feedback and testing to validate candidate transforms.
+Testing is no substitute for proof, but, as Klein et al.
+\cite{klein2012run} show, proof is no substitute for testing.
+Lightweight mechanization is a fruitful middle ground between pencil-and-paper analysis and fully-mechanized formal proof.
+We use Redex \cite{findler2010redex}, a domain-specific language for exploring language semantics, to provide feedback, thoroughly exercise candidates, and perform exploratory analysis.
+
+\setkeyword{define-language define-extended-language define-metafunction reduction-relation extend-reduction-relation define-metafunction/extension in-hole hole variable-not-otherwise-mentioned}
+\setkeyword{wcm ccm}
+
+\setspecialsymbol{lambda}{$\lambda$}
+\setspecialsymbol{->}{$\to$}
+\setspecialsymbol{-->}{$\rightarrow$}
+\setspecialsymbol{betav}{$\beta$v}
+\setconstant{error}
+\setkeyword{chi}
+
+\setspecialsymbol{lambdav}{$\lambda$v}
+\setspecialsymbol{lambdav-rr}{$\lambda$v-rr}
+\setspecialsymbol{lambdav-subst}{$\lambda$v-subst}
+
+\begin{schemedisplay}
+(define-language lambdav
+  (e (e e) x v error)
+  (x variable-not-otherwise-mentioned)
+  (v (lambda (x) e))
+  (E (E e) (v E) hole))
+\end{schemedisplay}
+
+\begin{schemedisplay}
+(define lambdav-rr
+  (reduction-relation lambdav
+   (--> (in-hole E ((lambda (x) e) v))
+        (in-hole E (lambdav-subst x v e))
+        "betav")
+   (--> (in-hole E x)
+        (in-hole E error)
+        "error: unbound identifier")
+   (--> (in-hole E (error e))
+        (in-hole E error)
+        "error in operator")
+   (--> (in-hole E (v error))
+        (in-hole E error)
+        "error in operand")))
+\end{schemedisplay}
+
+\begin{schemedisplay}
+(define-metafunction lambdav
+  lambdav-subst : x v e -> e
+  ;; 1. substitute in application
+  [(lambdav-subst x_1 v_1 (e_1 e_2))
+   ((lambdav-subst x_1 v_1 e_1) (lambdav-subst x_1 v_1 e_2))]
+  ;; 2a. substitute in variable (same)
+  [(lambdav-subst x_1 v_1 x_1)
+   v_1]
+  ;; 2b. substitute in variable (different)
+  [(lambdav-subst x_1 v_1 x_2)
+   x_2]
+  ;; 3a. substitute in abstraction (bound)
+  [(lambdav-subst x_1 v_1 (lambda (x_1) e_1))
+   (lambda (x_1) e_1)]
+  ;; 3b. substitute in abstraction (free)
+  [(lambdav-subst x_1 v_1 (lambda (x_2) e_1))
+   (lambda (x_2) (lambdav-subst x_1 v_1 e_1))]
+  ;; 4. substitute in error
+  [(lambdav-subst x_1 v_1 error)
+   error])
+\end{schemedisplay}
+
+\setspecialsymbol{lambdacm}{$\lambda$cm}
+\setspecialsymbol{lambdacm-rr}{$\lambda$cm-rr}
+\setspecialsymbol{lambdacm-subst}{$\lambda$cm-subst}
+
+Since \cm\ is a superset of \lv, we need only extend the definition of the \lv\ interpreter to accommodate the additions \cm\ brings.
+
+\begin{schemedisplay}
+(define-extended-language lambdacm lambdav
+  (e .... (wcm e e) (ccm))
+  (E (wcm v F) F)
+  (F (E e) (v E) (wcm E e) hole))
+\end{schemedisplay}
+
+Redex allows us to easily define a proper extension of a language, inheriting anything left unspecified.
+As similar as the \lv\ interpeter definition is to the \lv\ definition in figure \ref{lv-language-forms}, this \cm\ interpreter definition is to the \cm\ definition in figure \ref{cm-language-forms}.
+
+\begin{schemedisplay}
+(define lambdacm-rr
+  (extend-reduction-relation lambdav-rr lambdacm
+   (--> (in-hole E ((lambda (x) e) v))
+        (in-hole E (lambdacm-subst x v e))
+        "betav")
+   (--> (in-hole E (wcm v_1 (wcm v_2 e)))
+        (in-hole E (wcm v_2 e))
+        "wcm-collapse")
+   (--> (in-hole E (wcm v_1 v_2))
+        (in-hole E v_2)
+        "wcm")
+   (--> (in-hole E (ccm))
+        (in-hole E (chi E (lambda (x) (lambda (y) y))))
+        "chi")
+   (--> (in-hole E (wcm error e))
+        (in-hole E error)
+        "error in wcm mark expression")
+   (--> (in-hole E (wcm v error))
+        (in-hole E error)
+        "error in wcm body expression")))
+\end{schemedisplay}
+
+The first three rules in the reduction relation correspond with the three additional semantic rules found in \ref{cm-language-semantics}.
+The remaining handle cases introduced by the the new language forms' interaction with \scheme'error'.
+
+\begin{schemedisplay}
+(define-metafunction/extension lambdav-subst lambdacm
+  lambdacm-subst : x v e -> e
+  ;; 1. substitute in wcm form
+  [(lambdacm-subst x_1 v_1 (wcm e_1 e_2))
+   (wcm (lambdacm-subst x_1 v_1 e_1) (lambdacm-subst x_1 v_1 e_2))]
+  ;; 2. substitute in ccm form
+  [(lambdacm-subst x_1 v_1 (ccm))
+   (ccm)])
+\end{schemedisplay}
+
+The \scheme{lambdacm-subst} metafunction is extended to accommodate the additional forms in \cm.
+
+\begin{schemedisplay}
+(define-metafunction lambdacm
+  chi : E v -> v
+  [(chi hole v_ms)      v_ms]
+  [(chi (E e) v_ms)     (chi E v_ms)]
+  [(chi (v E) v_ms)     (chi E v_ms)]
+  [(chi (wcm E e) v_ms) (chi E v_ms)]
+  [(chi (wcm v E) v_ms) (chi E (lambda (p) ((p v) v_ms)))])
+\end{schemedisplay}
+
+\section{Proof}
+
+\section{Related Work}
+
+\section{Conclusion}
+
+\bibliographystyle{plain}
+\bibliography{paper}
+
 \end{document}

### 2012-09-21 paper.tex
Big dump of stuff.

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{slatex}
@@
 \newcommand{\cmrrs}[0]{\rightarrow_{cm}^{*}}
+\newcommand{\C}[1]{\mathcal{C}[#1]}
 
@@
 
+Continuation marks are good.
+Languages don't have them.
+This transformation should help.
+
+% uses of language transformation:
+% meaning of one language in terms of another (desugaring)
+% - reduce core code which simplifies implementations
+% easier to reason/optimize
+
+Language transformations are used for many reasons, but they all have one thing in common: the transformation must be meaningful: it must either preserve meaning or provide it.
+
+Language transformations either provide meaning or preserve meaning.
+
 Language transformation is useful for a variety of purposes.
-It can be used to give meaning to the source language in terms of the target language [cite something], provide a characterization more amenable to analysis.
+It can be used to define one language in terms of another, It can be used to give meaning to the source language in terms of the target language [cite something], provide a characterization more amenable to analysis or optimization.
 Language transformation can also be used to \emph{desugar} a language into a simpler core.
@@
 This includes algebraic steppers \cite{clements2001modeling}, security policies \cite{clements2004tail}, and aspects \cite{tucker2003pointcuts}.
+By defining such tools at the language level, such tools are more robust, cheaper to port, more easily changed, etc.
 
-Few languages offer continuation marks.
-Their inception lay in Racket \cite{plt-tr1} and they have been added experimentally to \emph{JavaScript} \cite{clements2008implementing}, but their adoption by other languages has been nonexistent.
+Despite their profound utility, few languages offer continuation marks.
+Their inception lays in Racket \cite{plt-tr1} and they have been added experimentally to \emph{JavaScript} \cite{clements2008implementing}, but their earnest adoption by other languages has been nonexistent.
+
+% transformation means that they can be compiled away and any higher order language can have them
+% transformation also means that they do not need to be built into the language
+
+% talk about composability of this transform to others
 
@@
 
-Terms $e$ in \cm\ are defined inductively by
-\begin{align}
-e=\, &\wcm{e}{e}\\
-     &\ccm\\
-     &\app{e}{e}\\
-     &v\\
-     &x
-\end{align}
+We define an extension of the call-by-value \lc\ with facilities for continuation marks. To begin, we briefly formalize the definition of the call-by-value \lc which we refer to as \lv.
+
+Terms $e$ in \lv\ are the familiar terms of the \lc, defined by
+\begin{equation}
+e=\app{e}{e}\,|\,v\,|\,x
+\end{equation}
 where 
 \begin{equation}
-v=\abs{x}{e}
+v=\abs{x}{e}\,|\,\mathbb{N}
 \end{equation}
-describes the values of the \lc.
+describes the values of the \lc\ extended with the natural numbers.
+
+The evaluation model of \lv\ requires a definition of evaluation contexts. We define evaluation contexts $E$ by
+\begin{equation}
+E=\app{E}{e}\,|\,\app{v}{E}\,|\,\hole
+\end{equation}
+where $\hole$ denotes a ``hole'' in the evaluation context, the ultimate destination of the evaluation of the expression that previously resided there.
+
+The semantics of \lv\ are defined simply by
+\begin{align}
+E[\app{\abs{x}{e}}{v}]  &\rr E[e[x\leftarrow v]]
+\end{align}
+where $e[x\leftarrow v]$ denotes a capture-avoiding substitution of every free occurrence of $x$ in $e$ with $v$.
+
+Because \cm\ is an extension of \lv, the definition of terms $e$ in \cm\ 
+\begin{equation}
+e=\wcm{e}{e}\,|\,\ccm\,|\,\app{e}{e}\,|\,v\,|\,x
+\end{equation}
+is identical to that of \lv\ with the addition of two forms for manipulating continuation marks: \scheme'wcm', short for \scheme'with-continuation-mark', which annotates the continuation with the evaluation of the given mark expression; and \scheme'ccm', short for \scheme'current-continuation-marks', which retrieves the current continuation marks.
 
@@
 \begin{align}
-E=\, &\wcm{v}{F}\\
-     &F\\
-F=\, &\app{E}{e}\\
-     &\app{v}{E}\\
-     &\wcm{E}{e}\\
-     &\hole
+E=\, &\wcm{v}{F}\,|\,F\\
+F=\, &\app{E}{e}\,|\,\app{v}{E}\,|\,\wcm{E}{e}\,|\,\hole
 \end{align}
@@
 E[\wcm{v}{v'}]          &\rr E[v']\\
-E[\ccm]                 &\rr E[\chi(E,\abs{x}{\abs{y}{y}})]
+E[\ccm]                 &\rr E[\chi(E)]
 \end{align}
-where the metafunction $\chi$ is defined by
+where $\chi$ is defined by
+\begin{equation}
+\chi(E)=\chi'(E,\mathbf{nil})
+\end{equation}
+and where $\chi'$ is defined in turn by
 \begin{align}
-\chi(\wcm{v}{F},vs) &= \chi(F,\abs{z}{\app{\app{z}{v}}{vs}})\\
-\chi(\app{E}{e},vs) &= \chi(E,vs)\\
-\chi(\app{v}{E},vs) &= \chi(E,vs)\\
-\chi(\wcm{E}{e},vs) &= \chi(E,vs)\\
-\chi(\hole,vs)      &= vs
-\end{align}.
+\chi'(\wcm{v}{F},vs) &= \chi'(F,\abs{z}{\app{\app{z}{v}}{vs}})\\
+\chi'(\app{E}{e},vs) &= \chi'(E,vs)\\
+\chi'(\app{v}{E},vs) &= \chi'(E,vs)\\
+\chi'(\wcm{E}{e},vs) &= \chi'(E,vs)\\
+\chi'(\hole,vs)      &= vs
+\end{align}
+Notice that $\chi'$ is defined in accumulator-passing style which has the effect of ordering the marks from most- to least-recently placed.
+The direct style formulation reverses this order.
+In the end, the ordering chosen is immaterial; we use this formulation to make the transform simpler.
+
+\subsection{\cm\ Example}
+
+\scheme|(wcm 4|
 
 \section{Transformation}
+
+A meaning-preserving transformation should commute with evaluation. By this, we mean that the following property should hold.
+
+\[
+\begin{array}{ccc}
+p & \cmrrs & v\\
+\downarrow_\mathcal{C} & & \downarrow_\mathcal{C}\\
+\C{p} & \lvrrs & \C{v}
+\end{array}
+\]
+
+This diagram captures the property that a meaning-preserving transformation should commute with evaluation. More succinctly, 
+\begin{equation}
+\label{meaning-preservation-property}
+p\cmrrs v\implies\C{p}\lvrrs\C{v}
+\end{equation}
+
+Our first approach was to use CPS because it treats all data and control flow uniformly [cite Sabry].
+We passed an additional parameter (or two) in the spirit of double-barrelled continuations (but only in spirit).
+Our first attempt passed a pair as the second argument that needed to be deconstructed and constructed in every recursive definition.
+This was a lot of mindless manipulation that had little to do with the essence of the transform.
+In what took far too long a time, we realized the extent of this manipulation and finally confronted the subconscious aversion to passing \emph{two} arguments which we found was baseless.
+
+Almost immediately after, we realized that CPS was wholly unnecessary and landed back where we started with a direct-style definition.
+
+A comprehensive language transformation need only be defined over each syntactic form of the language.
+Thus, our full transformation definition will comprise transformation definitions for each of the five syntactic forms of \cm.
+
+As variables, abstractions, and applications exist in the target language, we imagine that their transformation will be relatively straightforward.
+The essence of \cm\ is that programs within can apply information to and observe information about the context in which they are evaluated.
+The work of the transform is then to propagate contextual information to each expression, manipulating it appropriately as the context changes.
+
+The treatment of \scheme'wcm' will require care as it is especially sensitive to evaluation order. Assuming termination for all constituent evaluations, consider
+\begin{align*}
+       &E[\wcm{e_0}{\wcm{e_1}{e_2}}]\\
+\cmrrs &E[\wcm{v_0}{\wcm{e_1}{e_2}}]\\
+\cmrrs &E[\wcm{v_0}{\wcm{v_1}{e_2}}]\\
+\cmrr  &E[\wcm{v_1}{e_2}]\\
+\cmrrs &E[\wcm{v_1}{v_2}]\\
+\cmrr  &E[v_2]
+\end{align*}
+We will need to arrange arguments in such a way that the natural call-by-value evaluation order of \lv\ accomplishes correct evaluation order according to the semantics of \cm.
+
+The treatment of \scheme'ccm' simply needs access to the evaluation of $\chi$ on the current context.
+As \lv\ lacks any facility to observe the context, the transformation needs to explicitly provide this information to each \scheme'ccm' expression.
+The $\chi$ metafunction is continuous in the sense that a small change in the prefix of the input effects only a small change in the prefix of the output, a consequence of its particular recursive definition.
+This fact means that it is possible to compute $\chi$ incrementally.
+The transform would then need to ensure that each expression had access to information about its own context.
+
+Finally, we consider the preservation of proper tail-call behavior.
+Even though tail calls exist in \lv--indeed, they are prevalent--we cannot rely on their behavior to honor proper tail-call behavior through the transform.
+Thus, we would need to maintain information about the immediate context of the program: specifically, whether evaluation is occurring directly within the body of a \scheme'wcm' expression or not.
+
+To accomplish each of these, the transform is defined in the spirit of CPS. Recall that in
+CPS, every function receives an additional formal parameter and every call site receives a
+corresponding argument. The definition of $\mathcal{C}$ takes this a step further under
+which every function receives \emph{two} additional formal parameters and every call site
+two additional arguments. As in CPS, the first parameter encapsulates the continuation or,
+for our purposes, the evaluation context. The second encapsulates all contextual
+information observable to the original \cm\ program.
+
+% talk about continuity of chi
+
+\subsection{Example: Factorial}
 
@@
   (x variable-not-otherwise-mentioned)
-  (v (lambda (x) e))
+  (v (lambda (x) e) number)
   (E (E e) (v E) hole))
@@
         "betav")
+   (--> (in-hole E (number_1 v))
+        (in-hole E error)
+        "error: number in operator position")
    (--> (in-hole E x)
@@
    (lambda (x_2) (lambdav-subst x_1 v_1 e_1))]
-  ;; 4. substitute in error
+  ;; 4. substitute in number
+  [(lambdav-subst x_1 v_1 number_1)
+   number_1]
+  ;; 5. substitute in error
   [(lambdav-subst x_1 v_1 error)
@@
 \end{schemedisplay}
+
+This definition we must redefines evaluation contexts instead of merely extending them. This is to enforce the tail-call semantics [make sure those are mentioned earlier].
 
@@
 
+We can test that the property described by equation \ref{meaning-preservation-property} holds for a given program \scheme'p' with 
+\begin{schemedisplay}
+(define (meaning-preserved? p)
+  (alpha-eq? (transform (eval lambdacm p)) (eval lambdav (init (transform p)))))
+\end{schemedisplay}
+where \scheme'alpha-eq?' determines $\alpha$-equivalence between two \lc\ terms and \scheme'eval' is an alias for the Redex native \scheme'apply-reduction-relation*'.
+
+Redex provides convenient functions to initiate random testing.
+
+\begin{schemedisplay}
+(redex-check lambdacm e (meaning-preserved? e))
+\end{schemedisplay}
+
+\scheme'redex-check' generates random terms according to the grammar of the given language (\scheme'lambdacm') and category (\scheme'e') in search of counterexamples to the predicate.
+
 \section{Proof}
@@
 
+[Other people] have made a strong case for continuation marks.
+
 \section{Conclusion}
+
+We have provided and verified a correct transformation for continuation marks which, in effect, compiles them away.
 

### 2012-09-24 paper.tex
Reworked intro.

@@
 \begin{abstract}
-Continuation marks, a programming language feature which generalizes stack inspection, do not yet enjoy a transformation to the plain $\lambda$-calculus.
-We present a CPS-like transformation from the call-by-value $\lambda$-calculus augmented with continuation marks to the pure call-by-value $\lambda$-calculus.
-We discuss how such a transformation would simplify the construction of compilers which treat continuation marks correctly.
-We document an iterative, feedback-based approach using Redex.
-We accompany Finally, we verify the transformation with a meaning-preservation theorem.
+Continuation marks are a programming language feature which generalizes stack inspection.
+Despite its usefulness, this feature has not been adopted by languages which rely on stack inspection, e.g., for dynamic security checks.
+One reason for this neglect may be that continuation marks do not yet enjoy a transformation to the plain $\lambda$-calculus which would allow higher-order languages to provide continuation marks at little cost.
+
+We present direct and CPS-like transformations from the call-by-value $\lambda$-calculus augmented with continuation marks to the pure call-by-value $\lambda$-calculus.
+We discuss how such transformations would simplify the construction of compilers which treat continuation marks correctly.
+We document the aide of Redex, a domain-specific language for exploring language semantics, in the discovery of correct transformations.
+Finally, we verify the transformations with the proof of a meaning-preservation theorem.
 \end{abstract}
@@
 
-Continuation marks are good.
-Languages don't have them.
-This transformation should help.
+There are numerous programming language instruments which rely on stack inspection to function.
+For example, statistical profilers sample the stack regularly to record active functions, algebraic steppers observe the stack to represent the evaluation context of an expression, and debuggers naturally require consistent access to the stack.
+In essence, each of these examples rely on implementation-specific information and must be maintained accordingly as the instrumented language undergoes optimizations and ventures across platforms.
+This makes these tools brittle and increases the porting cost of the language ecosystem.
+Each of these examples would benefit from a generalized stack-inspection mechanism available within the instrumented language itself.
+If written in the enhanced instrumented language, this would increase their robustness and remove their porting cost entirely.
 
+Continuation marks \cite{clements2006portable} are a programming language feature which generalizes stack inpection.
+Not only do they dramatically simplify instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail} and to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}.
+
+Continuation marks originated in PLT Scheme (now Racket \cite{plt-tr1}) and have been added experimentally to \emph{JavaScript} \cite{clements2008implementing}.
+However, in spite of their usefulness, continuation marks have remained absent from programming languages at large.
+One reason for this may be that continuation marks do not yet enjoy a transformation to the \lc.
+With a correct transformation, higher-order language semanticists would not need to reconcile the feature with other features in the language (provided they had already done so with $\lambda$) and their compiler-writers would not need to worry about complicating their implementations (for the same reason).
+
+Language transformations must be meaningful: they must preserve meaning or provide it. Continuation marks have special meaning in the presence of tail-call optimization.
 % uses of language transformation:
@@
 
-Language transformations are used for many reasons, but they all have one thing in common: the transformation must be meaningful: it must either preserve meaning or provide it.
+Language transformations are used for many reasons, but they he transformation must be meaningful: it must either preserve meaning or provide it.
 
@@
 \begin{equation}
-v=\abs{x}{e}\,|\,\mathbb{N}
+v=\abs{x}{e}
 \end{equation}
-describes the values of the \lc\ extended with the natural numbers.
+describes the values of the \lc.
 
@@
 \subsection{Example: Factorial}
+
+For convenience, we will augment the base language with direct forms for expressing factorial. Define terms $e$ in \lv\ by
+\begin{equation}
+e=\app{e}{e}\,|\,(*\,e\,e)\,|\,\app{\mathrm{pred}}{e}\,|\,\app{\mathrm{zero?}}{e}\,|\,v\,|\,x
+\end{equation}
+and values $v$ by
+\begin{equation}
+v=\abs{x}{e}\,|\,\mathbb{N}
+\end{equation}
+.
+
+We can express the factorial function by the following
+\begin{schemedisplay}
+(Y (lambda (f)
+     (lambda (n)
+       (((zero? n)
+         1)
+        (* n (f (pred n)))))))
+\end{schemedisplay}
+
+and with continuation marks by
+\begin{schemedisplay}
+(Y (lambda (f)
+     (lambda (n)
+       (((zero? n)
+         (print (ccm) 1))
+        (* n (wcm n (f (pred n))))))))
+\end{schemedisplay}
 

### 2012-09-24 paper.tex
Good things.

@@
 % easier to do a type system.
+
+\setkeyword{wcm ccm with-continuation-mark current-continuation-marks}
+\setkeyword{print fac zero?}
 
@@
 Each of these examples would benefit from a generalized stack-inspection mechanism available within the instrumented language itself.
-If written in the enhanced instrumented language, this would increase their robustness and remove their porting cost entirely.
+If written in such an enhanced language, each instrument would be more robust and port for free.
 
 Continuation marks \cite{clements2006portable} are a programming language feature which generalizes stack inpection.
-Not only do they dramatically simplify instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail} and to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}.
+Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail} and to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}.
 
@@
 With a correct transformation, higher-order language semanticists would not need to reconcile the feature with other features in the language (provided they had already done so with $\lambda$) and their compiler-writers would not need to worry about complicating their implementations (for the same reason).
+
+We present two transformations of continuation marks--one direct and one CPS-like--to the call-by-value \lc.
+We employ Redex, a domain-specific semantic meta-language, to provide feedback through randomized testing and reduction trace visualization.
+
+\section{Background}
+
+\subsection{Continuation Marks}
+
+Continuation marks allow the program to annotate and observe the stack (or continuation).
+This is accomplished via two surface-level syntactic forms in the language: \scheme'with-continuation-mark', abbreviated \scheme'wcm'; and \scheme'current-continuation-marks', abbreviated \scheme'ccm'.
+
+A \scheme'wcm' expression of the form \scheme|(wcm mark-expr body-expr)| annotates the youngest portion of the continuation with \scheme'mark-value', the evaluation of \scheme'mark-expr', before evaluating \scheme'body-expr' to \scheme'body-value'.
+The entire expression takes on \scheme'body-value'.
+
+Conceptually, a \scheme'ccm' expression of the form \scheme|(ccm)| traverses the continuation accumulating a list of annotations ascending in age.
+
+For example, the program \scheme|(wcm (+ 1 2) (ccm))| evaluates as
+\begin{schemedisplay}
+(wcm (+ 1 2) (ccm))
+(wcm 3 (ccm))
+(wcm 3 '(3))
+'(3)
+\end{schemedisplay}
+
+Continuation marks respect tail-call optimization.
+
+\begin{schemedisplay}
+(define (fac n)
+  (if (zero? n)
+      (begin
+        (print (ccm))
+        1)
+      (wcm n (* n (fac (- n 1))))))
+\end{schemedisplay}
+
+\begin{schemedisplay}
+(fac 2)
+(if (zero? 2) ... (wcm 2 (* 2 (fac (- 2 1)))))
+(wcm 2 (* 2 (fac (- 2 1))))
+(wcm 2 (* 2 (fac 1)))
+(wcm 2 (* 2 (if (zero? 1) ... (wcm 1 (* 1 (fac (- 1 1)))))))
+(wcm 2 (* 2 (wcm 1 (* 1 (fac (- 1 1))))))
+(wcm 2 (* 2 (wcm 1 (* 1 (fac 0)))))
+(wcm 2 (* 2 (wcm 1 (* 1 (if (zero? 0) (begin (print (ccm)) 1) ...)))))
+(wcm 2 (* 2 (wcm 1 (* 1 (begin (print (ccm)) 1)))))
+(wcm 2 (* 2 (wcm 1 (* 1 (begin (print '(1 2)) 1)))))
+'(1 2)
+(wcm 2 (* 2 (wcm 1 (* 1 (begin 1)))))
+(wcm 2 (* 2 (wcm 1 (* 1 1))))
+(wcm 2 (* 2 (wcm 1 1)))
+(wcm 2 (* 2 1))
+(wcm 2 2)
+2
+\end{schemedisplay}
+
+\begin{schemedisplay}
+(define (fac n acc)
+  (if (zero? n)
+      (begin
+        (print (ccm))
+        acc)
+      (wcm n (fac (- n 1) (* n acc)))))
+\end{schemedisplay}
+
+\begin{schemedisplay}
+(fac 2 1)
+(if (zero? 2) ... (wcm 2 (fac (- 2 1) (* 2 1))))
+(wcm 2 (fac (- 2 1) (* 2 1)))
+(wcm 2 (fac 1 (* 2 1)))
+(wcm 2 (fac 1 2))
+(wcm 2 (if (zero? 1) ... (wcm 1 (fac (- 1 1) (* 1 2)))))
+(wcm 2 (wcm 1 (fac (- 1 1) (* 1 2))))
+(wcm 1 (fac (- 1 1) (* 1 2)))
+(wcm 1 (fac 0 (* 1 2)))
+(wcm 1 (fac 0 2))
+(wcm 1 (if (zero? 0) (begin (print (ccm)) 2) ...))
+(wcm 1 (begin (print (ccm)) 2))
+(wcm 1 (begin (print '(1)) 2))
+'(1)
+(wcm 1 (begin 2))
+(wcm 1 2)
+2
+\end{schemedisplay}
 

### 2012-09-26 paper.tex
Lots of additions.

@@
 For example, statistical profilers sample the stack regularly to record active functions, algebraic steppers observe the stack to represent the evaluation context of an expression, and debuggers naturally require consistent access to the stack.
-In essence, each of these examples rely on implementation-specific information and must be maintained accordingly as the instrumented language undergoes optimizations and ventures across platforms.
+Each of these examples relies on implementation-specific information and must be maintained as the instrumented language undergoes optimizations and ventures across platforms.
 This makes these tools brittle and increases the porting cost of the language ecosystem.
@@
 A \scheme'wcm' expression of the form \scheme|(wcm mark-expr body-expr)| annotates the youngest portion of the continuation with \scheme'mark-value', the evaluation of \scheme'mark-expr', before evaluating \scheme'body-expr' to \scheme'body-value'.
-The entire expression takes on \scheme'body-value'.
+If an annotation, or \emph{mark}, already exists on the youngest portion, it is replaced by the new mark.
+The entire \scheme'wcm' expression takes on \scheme'body-value'.
 
-Conceptually, a \scheme'ccm' expression of the form \scheme|(ccm)| traverses the continuation accumulating a list of annotations ascending in age.
+A \scheme'ccm' expression of the form \scheme|(ccm)| traverses the continuation accumulating a list of annotations ascending in age.
 
-For example, the program \scheme|(wcm (+ 1 2) (ccm))| evaluates as
-\begin{schemedisplay}
-(wcm (+ 1 2) (ccm))
-(wcm 3 (ccm))
-(wcm 3 '(3))
-'(3)
-\end{schemedisplay}
-
-Continuation marks respect tail-call optimization.
-
+If the factorial function is defined by
 \begin{schemedisplay}
@@
 \end{schemedisplay}
-
+then the expression \scheme|(fac 2)| evaluates as
 \begin{schemedisplay}
@@
 
+Continuation marks respect tail-call optimization.
+Recall that in a language that is tail-call optimized, the continuation doesn't grow at the occurrence of a tail call as instead the previous execution context is subsumed by the new.
+Thus, with a tail-recursive definition of the factorial function as in
 \begin{schemedisplay}
@@
 \end{schemedisplay}
-
+evaluation of \scheme|(fac 2)| proceeds as
 \begin{schemedisplay}
@@
 \end{schemedisplay}
+and the tail-call behavior is clear.
 
@@
 
+\newtheorem{defn}{Definition}
+
+\newenvironment{schemedefn}[1]{\begin{defn}$#1$}{\end{defn}}
+\newenvironment{namedschemedefn}[2]{\begin{defn}\label{#1}$#2$}{\end{defn}}
+
+\setspecialsymbol{TRUE}{$\mathbf{true}$}
+\setspecialsymbol{FALSE}{$\mathbf{false}$}
+\setspecialsymbol{PAIR}{$\mathbf{pair}$}
+\setspecialsymbol{FST}{$\mathbf{fst}$}
+\setspecialsymbol{SND}{$\mathbf{snd}$}
+\setspecialsymbol{NIL}{$\mathbf{nil}$}
+
+\newcommand{\true}{\mathbf{true}}
+\newcommand{\false}{\mathbf{false}}
+\newcommand{\nil}{\mathbf{nil}}
+\newcommand{\pair}[2]{((\mathbf{pair}\,#1)\,#2)}
+\newcommand{\eval}[1]{\overline{#1}}
+
+\begin{defn}
+$\true=\abs{x}{\abs{y}{x}}$
+\end{defn}
+
+\begin{defn}
+$\false=\abs{x}{\abs{y}{y}}$
+\end{defn}
+
+\begin{defn}
+$\mathbf{pair}=\abs{a}{\abs{b}{\abs{z}{\app{\app{z}{a}}{b}}}}$
+\end{defn}
+
+\begin{defn}
+$\mathbf{fst}=\abs{p}{\app{p}{\true}}$
+\end{defn}
+
+\begin{defn}
+$\mathbf{snd}=\abs{p}{\app{p}{\false}}$
+\end{defn}
+
+\begin{defn}
+$\nil=\false$
+\end{defn}
+
+\begin{schemedefn}{\C{\app{e_0}{e_1}}}
+\begin{schemedisplay}
+(lambda (k)
+   (lambda (f)
+     (lambda (m)
+       (((Ce_0
+          (lambda (a)
+            (((Ce_1
+               (lambda (b)
+                 ((((a b) k) f) m)))
+              FALSE)
+             m)))
+         FALSE)
+        m))))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\C{\wcm{e_0}{e_1}}}
+\begin{schemedisplay}
+(lambda (k)
+  (lambda (f)
+    (lambda (m)
+      (((Ce_0
+         (lambda (n) ((lambda (k) 
+                        (k (((m (lambda (x) x)) (lambda (z) z)) (lambda (z) z))))
+                      (lambda (t) 
+                        ((lambda (k) (k ((f C[(SND t)]) t)))
+                         (lambda (r) 
+                           (((Ce_1 k) TRUE) C[(PAIR v r)])))))))
+        FALSE)
+       m))))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\C{\ccm}}
+\begin{schemedisplay}
+(lambda (k)
+  (lambda (f)
+    (lambda (m)
+      (((m k) (lambda (z) z)) (lambda (z) z)))))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{defn}
+$\C{v_0}=\C{\abs{x}{e'_0}}=\abs{k}{\abs{f}{\abs{m}{\app{k}{\abs{x}{\C{e'_0}}}}}}$
+\end{defn}
+
+\begin{defn}
+$\C{x}=\abs{k}{\abs{f}{\abs{m}{\app{k}{x}}}}$
+\end{defn}
+
 \subsection{Example: Factorial}
@@
 
+The correctness of the transform lies in the property that it commutes.
+In order to test for this property, we must construct evaluation models for the source language \cm\ and the target language \lv.
+Since \cm\ is an extension of \lv, many of its forms and semantics are inherited.
+Redex allows us to exploit this fact by defining a model for \lv\ first and then extending that model with the enhancements of \cm.
+
 \setkeyword{define-language define-extended-language define-metafunction reduction-relation extend-reduction-relation define-metafunction/extension in-hole hole variable-not-otherwise-mentioned}
@@
 \setconstant{error}
-\setkeyword{chi}
+\setkeyword{chi chi_p}
 
@@
 \setspecialsymbol{lambdav-subst}{$\lambda$v-subst}
+
+We first define the form of terms $e$, values $v$, and evaluation contexts $E$.
 
@@
 \end{schemedisplay}
+
+We then define the reduction relation for \lv. It consists of five rules. The first rule captures the sole semantic rule of \lv; the remaining handle error conditions.
 
@@
 \end{schemedisplay}
+
+The reduction relation requires one metafunction to perform substitution.
+We omit any capture-avoidance facilities to keep the model simple and argue that it is unnecessary as we generate fresh variables in the transform.
 
@@
 
-Since \cm\ is a superset of \lv, we need only extend the definition of the \lv\ interpreter to accommodate the additions \cm\ brings.
+We now define the forms of terms $e$, and evaluation contexts $E$ and $F$, inheriting definitions from \lv\ where appropriate.
 
@@
 
-This definition we must redefines evaluation contexts instead of merely extending them. This is to enforce the tail-call semantics [make sure those are mentioned earlier].
-
-Redex allows us to easily define a proper extension of a language, inheriting anything left unspecified.
-As similar as the \lv\ interpeter definition is to the \lv\ definition in figure \ref{lv-language-forms}, this \cm\ interpreter definition is to the \cm\ definition in figure \ref{cm-language-forms}.
+We extend the reduction relation of \lv.
+We begin by redefining application in terms of a substitution function aware of \cm.
+The following three definitions capture the remaining semantic rules of \cm.
+The remaining definitions handle error cases specific to \cm.
 
@@
    (--> (in-hole E (ccm))
-        (in-hole E (chi E (lambda (x) (lambda (y) y))))
+        (in-hole E (chi E))
         "chi")
@@
 
-The first three rules in the reduction relation correspond with the three additional semantic rules found in \ref{cm-language-semantics}.
-The remaining handle cases introduced by the the new language forms' interaction with \scheme'error'.
+We extend the definition of substitution to accommodate the additional forms introduced by \cm.
 
@@
 
-The \scheme{lambdacm-subst} metafunction is extended to accommodate the additional forms in \cm.
+The definition of \scheme'chi' and \scheme'chi_p' mirror their definitions exactly.
 
@@
 (define-metafunction lambdacm
-  chi : E v -> v
-  [(chi hole v_ms)      v_ms]
-  [(chi (E e) v_ms)     (chi E v_ms)]
-  [(chi (v E) v_ms)     (chi E v_ms)]
-  [(chi (wcm E e) v_ms) (chi E v_ms)]
-  [(chi (wcm v E) v_ms) (chi E (lambda (p) ((p v) v_ms)))])
+  chi : E -> v
+  [(chi E) (chi_p E (lambda (x) (lambda (y) y)))])
+
+(define-metafunction lambdacm
+  chi_p : E v -> v
+  [(chi_p hole v_ms)      v_ms]
+  [(chi_p (E e) v_ms)     (chi_p E v_ms)]
+  [(chi_p (v E) v_ms)     (chi_p E v_ms)]
+  [(chi_p (wcm E e) v_ms) (chi_p E v_ms)]
+  [(chi_p (wcm v E) v_ms) (chi_p E (lambda (p) ((p v) v_ms)))])
 \end{schemedisplay}
+
+We are now prepared to test various semantic properties of these two languages.
+Of course, we are interested in testing the commutativity of our transform.
+We observe that a mark-passing approach, much like a continuation-passing approach, abstracts terms and so requires a transformed term to be primed with an argument to initiate reduction.
+For this reason, we accompany each transform with an initialization function which primes a term appropriately for the corresponding transform.
 

### 2012-09-26 fac-cm.js

@@ -1,0 +1,20 @@
+var TRUE = function( x ) {
+	return function( y ) {
+		return x;
+	}
+};
+
+var FALSE = function( x ) {
+	return function( y ) {
+		return y;
+	}
+};
+
+function listChuchDecode( l ) {
+	if( l( 0 )( 'false' ) === 'false' ) {
+		return [];
+	}
+	else {
+		return [ l( TRUE ) ].concat( l( FALSE ) );
+	}
+} 

### 2012-09-26 fac-cm.js
Factorial working in JS.

@@ -1,20 +1,40 @@
 var TRUE = function( x ) {
-	return function( y ) {
-		return x;
+	return function( f ) {
+		return function( m ) {
+			return function( y ) {
+				return function( f ) {
+					return function( m ) {
+						return x;
+					}
+				}
+			}
+		}
 	}
 };
 
 var FALSE = function( x ) {
-	return function( y ) {
-		return y;
+	return function( f ) {
+		return function( m ) {
+			return function( y ) {
+				return function( f ) {
+					return function( m ) {
+						return y;
+					}
+				}
+			}
+		}
 	}
 };
 
-function listChuchDecode( l ) {
-	if( l( 0 )( 'false' ) === 'false' ) {
+function listChurchDecode( l ) {
+	try {
+		l( 'not a function' )( 'f' )( 'm' );
+
 		return [];
 	}
-	else {
-		return [ l( TRUE ) ].concat( l( FALSE ) );
+	catch( e ) {
+		return [ l( TRUE )( 'f' )( 'm' ) ].concat( listChurchDecode( l( FALSE )( 'f' )( 'm' ) ) );
 	}
 } 
+
+console.log((((function(f11814){return (function(m11815){return ((((((function(f11818){return (function(m11819){return ((((((function(f11822){return (function(m11823){return (function(f){return (function(f11826){return (function(m11827){return ((((((function(f11830){return (function(m11831){return (function(x){return (function(f11834){return (function(m11835){return ((((((function(f11838){return (function(m11839){return f;});})((function(x){return (function(y){return y;});})))(m11835))((((function(f11842){return (function(m11843){return (function(v){return (function(f11846){return (function(m11847){return ((((((function(f11850){return (function(m11851){return ((((((function(f11854){return (function(m11855){return x;});})((function(x){return (function(y){return y;});})))(m11851))((((function(f11858){return (function(m11859){return x;});})((function(x){return (function(y){return y;});})))(m11851))))(f11850))(m11851));});})((function(x){return (function(y){return y;});})))(m11847))((((function(f11862){return (function(m11863){return v;});})((function(x){return (function(y){return y;});})))(m11847))))(f11846))(m11847));});});});});})((function(x){return (function(y){return y;});})))(m11835))))(f11834))(m11835));});});});});})((function(x){return (function(y){return y;});})))(m11827))((((function(f11866){return (function(m11867){return (function(x){return (function(f11870){return (function(m11871){return ((((((function(f11874){return (function(m11875){return f;});})((func
[…]

### 2012-09-26 paper.tex
Really, the title is the best thing here.

@@
 
-\title{A Correct Compiler for Continuation Marks}
+\title{Instant Continuation Marks--Just Add $\lambda$}
 
@@
 
-A meaning-preserving transformation should commute with evaluation. By this, we mean that the following property should hold.
+A meaning-preserving transformation should commute with evaluation. By this, we mean that the following property should hold for programs $p\in\lambda_{cm}$.
 
@@
 
-This diagram captures the property that a meaning-preserving transformation should commute with evaluation. More succinctly, 
+This diagram captures the property that a meaning-preserving transformation commutes with evaluation. More succinctly, 
 \begin{equation}
@@
 information observable to the original \cm\ program.
-
-% talk about continuity of chi
 
@@
 
-\section{Related Work}
+\section{Continuation Marks in JavaScript}
 
-[Other people] have made a strong case for continuation marks.
+% add stuff to language to express factorial
+% define transformation of new constructs
+% transform factorial function
+% transliterate it to JS
+% see continuation marks in JS
+
+% we can add continuation marks to the compiler, but we can't deploy a virtual machine wherever we want. We need to leverage the existing, entrenched virtual machines. One approach would be to
+
+Given that we can express continuation marks in the \lc, we can see that the only thing necessary for continuation marks is first-class functions.
+JavaScript is a ubiquitous higher-order language that lacks continuation marks.
+Clements et al.
+experimentally added continuation marks to JavaScript by directly modifying Mozilla's Rhino implementation \cite{clements2008implementing}.
+This is a natural fit as the implementation is designed to optimize tail calls providing a venue in which the full semantics of continuation marks can be observed.
+
+Many implementations do not optimize tail calls.
+We could modify these implementations to first optimize tail calls and then support continuation marks, but this does nothing to the existing, entrenched virtual machines in use today.
+For this reason, we define a superset of JavaScript with continuation marks and a transformation to base JavaScript.
+The transformation utilizes the tail-position flag and mark-passing style of the \lc\ transform.
+
+functions receive two additional arguments
+if a call is in tail position, pass that flag.
+I don't know.
+
+We transliterate
 
 \section{Conclusion}
+
+Continuation marks support a bevy of instrumentation tools and advanced language features in a generalized, portable way.
+Despite their demonstrated utility, they have not yet found their way into most languages.
+A verified characterization of continuation marks in a pure computational language provides implementors of higher-order languages a correct compiler for continuation marks.
+
+Stack inspection is a common approach for instrumentation and dynamic
+
+Instrumentation tools abound, and stack inspection is a common method to in
+
+Stack inspection is a common approach to natural method to obtain dynamic information, but typical implementation
 

### 2012-09-26 paper.tex
Removed stuff.

@@
 
-Stack inspection is a common approach for instrumentation and dynamic
-
-Instrumentation tools abound, and stack inspection is a common method to in
-
-Stack inspection is a common approach to natural method to obtain dynamic information, but typical implementation
-
-We have provided and verified a correct transformation for continuation marks which, in effect, compiles them away.
-
 \bibliographystyle{plain}

### 2012-09-26 paper.tex
Changing intro and tone of paper.

@@
 We present direct and CPS-like transformations from the call-by-value $\lambda$-calculus augmented with continuation marks to the pure call-by-value $\lambda$-calculus.
-We discuss how such transformations would simplify the construction of compilers which treat continuation marks correctly.
+We discuss how such transformations simplify the construction of compilers which treat continuation marks correctly.
 We document the aide of Redex, a domain-specific language for exploring language semantics, in the discovery of correct transformations.
@@
 \section{Introduction}
+
+% motivate continuation marks
+% motivate continuation marks in more languages
+% discuss why they are not in more languages
+% discuss approaches to getting them in languages with pros/cons
+% ...culminating in our approach in terms of lambda
 
@@
 
-Continuation marks originated in PLT Scheme (now Racket \cite{plt-tr1}) and have been added experimentally to \emph{JavaScript} \cite{clements2008implementing}.
+Continuation marks originated in PLT Scheme (now Racket \cite{plt-tr1}).
 However, in spite of their usefulness, continuation marks have remained absent from programming languages at large.
-One reason for this may be that continuation marks do not yet enjoy a transformation to the \lc.
-With a correct transformation, higher-order language semanticists would not need to reconcile the feature with other features in the language (provided they had already done so with $\lambda$) and their compiler-writers would not need to worry about complicating their implementations (for the same reason).
+One reason for this may be that retrofitting virtual machines to accommodate the level of stack inspection continuation marks must provide is expensive, especially when the virtual machine uses the host stack for efficiency.
+
+JavaScript, for example, enjoys ubiquity as the \emph{lingua franca} of the web and is an ideal candidate for continuation marks.
+One approach to adding continuation marks to JavaScript would be to alter numerous mature implemenations--indeed, Clements et al.
+accomplished this for Mozilla's Rhino compiler \cite{clements2008implementing}--and deploy them all.
+This leaves the devices that will never upgrade in the dust.
+The effort required to add continuation marks with this approach is immense.
+Nevertheless, it could benefit from such an infusion.
+Another approach would be to express continuation marks in terms of a subset of JavaScript--the smaller, the better.
+This macro-style approach would keep in the spirit of JavaScript as the \emph{lingua franca} of the web by desugaring our enhanced language into JavaScript.
+
+but we actually did it for the lambda calculus
+
+With such a transformation, higher-order language semanticists would not need to reconcile the feature with other features in the language (provided they had already done so with $\lambda$) and their compiler writers would not need to worry about complicating their implementations (for the same reason).
 

### 2012-09-27 paper.tex
Intro

--- 2012-09-26 paper.tex
@@
 
-Continuation marks originated in PLT Scheme (now Racket \cite{plt-tr1}) and have been added experimentally to \emph{JavaScript} \cite{clements2008implementing}.
+Continuation marks originated in PLT Scheme (now Racket \cite{plt-tr1}).
 However, in spite of their usefulness, continuation marks have remained absent from programming languages at large.
-One reason for this may be that continuation marks do not yet enjoy a transformation to the \lc.
-With a correct transformation, higher-order language semanticists would not need to reconcile the feature with other features in the language (provided they had already done so with $\lambda$) and their compiler-writers would not need to worry about complicating their implementations (for the same reason).
+One reason for this is that retrofitting virtual machines to accommodate the level of stack inspection continuation marks must provide is expensive, especially when the virtual machines use the host stack for efficiency.
 
+For example, the ubiquitous JavaScript is an ideal candidate for the addition of continuation marks.
+However, as the \emph{lingua franca} of the web, it has numerous mature implementations which have been heavily optimized.
+To add continuation marks to JavaScript amounts to modifying each implementation upstream, to say nothing of amending the JavaScript standard.
+(Clements et al.
+successfully added continuation marks in Mozilla's Rhino compiler \cite{clements2008implementing}, but it remains a proof-of-concept.)
+
+To avoid this roadblock, we instead take a macro-style approach; that is, we enhance the core language with facilities to manipulate continuation marks and desugar the enhanced language back to the core.
+To make our desugaring transformation portable to other languages, we define it over the \lc, the common core of most higher-order languages.
+With such a transformation, language semanticists do not need to reconcile the feature with other features in the language (provided they have already done so with $\lambda$) and their compiler writers do not need to worry about complicating their implementations (for the same reason).
+
+We begin this paper exposing continuation marks further.
+We then present a core and enhanced language along with Redex models for each.
+We discuss the necessary properties for a meaningful transform.
 We present two transformations of continuation marks--one direct and one CPS-like--to the call-by-value \lc.
-We employ Redex, a domain-specific semantic meta-language, to provide feedback through randomized testing and reduction trace visualization.
+We then prove a meaning-preservation theorem.
 
@@
 
-\section{Continuation Marks in JavaScript}
-
-% add stuff to language to express factorial
-% define transformation of new constructs
-% transform factorial function
-% transliterate it to JS
-% see continuation marks in JS
-
-% we can add continuation marks to the compiler, but we can't deploy a virtual machine wherever we want. We need to leverage the existing, entrenched virtual machines. One approach would be to
-
-Given that we can express continuation marks in the \lc, we can see that the only thing necessary for continuation marks is first-class functions.
-JavaScript is a ubiquitous higher-order language that lacks continuation marks.
-Clements et al.
-experimentally added continuation marks to JavaScript by directly modifying Mozilla's Rhino implementation \cite{clements2008implementing}.
-This is a natural fit as the implementation is designed to optimize tail calls providing a venue in which the full semantics of continuation marks can be observed.
-
-Many implementations do not optimize tail calls.
-We could modify these implementations to first optimize tail calls and then support continuation marks, but this does nothing to the existing, entrenched virtual machines in use today.
-For this reason, we define a superset of JavaScript with continuation marks and a transformation to base JavaScript.
-The transformation utilizes the tail-position flag and mark-passing style of the \lc\ transform.
-
-functions receive two additional arguments
-if a call is in tail position, pass that flag.
-I don't know.
-
-We transliterate
-
 \section{Conclusion}

### 2012-09-28 paper.tex
I would rather work on this.

@@
 \usepackage{amsmath}
+%\usepackage{amsthm}
 \usepackage{amssymb}
@@
 
-There are numerous programming language instruments which rely on stack inspection to function.
-For example, statistical profilers sample the stack regularly to record active functions, algebraic steppers observe the stack to represent the evaluation context of an expression, and debuggers naturally require consistent access to the stack.
-Each of these examples relies on implementation-specific information and must be maintained as the instrumented language undergoes optimizations and ventures across platforms.
+Numerous programming language instruments rely on stack inspection to function.
+Statistical profilers sample the stack regularly to record active functions, algebraic steppers observe the stack to represent the evaluation context of an expression, and debuggers naturally require consistent access to the stack.
+Each of these relies on implementation-specific information and must be maintained as the instrumented language undergoes optimizations and ventures across platforms.
 This makes these tools brittle and increases the porting cost of the language ecosystem.
 Each of these examples would benefit from a generalized stack-inspection mechanism available within the instrumented language itself.
-If written in such an enhanced language, each instrument would be more robust and port for free.
+If written in such an enhanced language, each instrument would be more robust, more easily modified, and would port for free.
 
@@
 
-\section{Background}
-
-\subsection{Continuation Marks}
+\section{Continuation Marks}
 
@@
 
-Continuation marks respect tail-call optimization.
-Recall that in a language that is tail-call optimized, the continuation doesn't grow at the occurrence of a tail call as instead the previous execution context is subsumed by the new.
+Continuation marks respect tail-call optimization (TCO).
+A tail call in a language with TCO does not enlarge the continuation.
+Instead, the previous portion of the execution context is subsumed by the new.
+Conceptually, the continuation does not grow to accommodate new marks; if one is made, the previous is overwritten.
 Thus, with a tail-recursive definition of the factorial function as in
@@
 
-Language transformations must be meaningful: they must preserve meaning or provide it. Continuation marks have special meaning in the presence of tail-call optimization.
+%Language transformations must be meaningful: they must preserve meaning or provide it. Continuation marks have special meaning in the presence of tail-call optimization.
 % uses of language transformation:
@@
 
-Language transformations are used for many reasons, but they he transformation must be meaningful: it must either preserve meaning or provide it.
+%Language transformations are used for many reasons, but they he transformation must be meaningful: it must either preserve meaning or provide it.
 
-Language transformations either provide meaning or preserve meaning.
+%Language transformations either provide meaning or preserve meaning.
 
-Language transformation is useful for a variety of purposes.
+%Language transformation is useful for a variety of purposes.
 It can be used to define one language in terms of another, It can be used to give meaning to the source language in terms of the target language [cite something], provide a characterization more amenable to analysis or optimization.
@@
 This
-Transforming has the following benefits:
-reduce core code (desugaring, simplifies implementations)
-give meaning to constructs (semantically, as well as, say, macros)
-local transformations - macros, for example
-global transformations - compilers of all kinds, for example
+%Transforming has the following benefits:
+%reduce core code (desugaring, simplifies implementations)
+%give meaning to constructs (semantically, as well as, say, macros)
+%local transformations - macros, for example
+%global transformations - compilers of all kinds, for example
 
-talk about the common practice of analyzing an extension of the lambda calculus, and a subpractice of doing so via a transformation to the core.
+%talk about the common practice of analyzing an extension of the lambda calculus, and a subpractice of doing so via a transformation to the core.
 
-While continuation marks have had significant treatment, they do not currently enjoy a transformation to the plain $\lambda$-calculus.
+%While continuation marks have had significant treatment, they do not currently enjoy a transformation to the plain $\lambda$-calculus.
 
-\section{Continuation Marks}
+%\section{Continuation Marks}
 
-Continuation marks generalize stack inspection, allowing a bevy of features that rely on some form of it to be defined at the language level.
+%Continuation marks generalize stack inspection, allowing a bevy of features that rely on some form of it to be defined at the language level.
 This includes algebraic steppers \cite{clements2001modeling}, security policies \cite{clements2004tail}, and aspects \cite{tucker2003pointcuts}.
@@
 
-Despite their profound utility, few languages offer continuation marks.
+%Despite their profound utility, few languages offer continuation marks.
 Their inception lays in Racket \cite{plt-tr1} and they have been added experimentally to \emph{JavaScript} \cite{clements2008implementing}, but their earnest adoption by other languages has been nonexistent.
@@
 
-Racket offers desugaring facilities to the user via a powerful macro system. This requires and is limited to local transformations limited to expression boundaries.
+%Racket offers desugaring facilities to the user via a powerful macro system. This requires and is limited to local transformations limited to expression boundaries.
 
-A transformation-based compiler, such as a CPS compiler \cite{appel2007compiling}, effects a global transformation of the program.
+%A transformation-based compiler, such as a CPS compiler \cite{appel2007compiling}, effects a global transformation of the program.
 This effect discards original source or transforms it so radically that it is suitable only for the levels beneath the programmer's regular interaction.
 
-\section{\cm}
+\section{Language Core and Extension}
 
-We define an extension of the call-by-value \lc\ with facilities for continuation marks. To begin, we briefly formalize the definition of the call-by-value \lc which we refer to as \lv.
+We begin by formally defining the core language the transformation produces and the extended language over which the transformation is defined.
+
+\subsection{\lv}
+
+The target language of the transformation, \lv, is simply the call-by-value \lc.
 
@@
 
+\subsection{\cm}
+
+The source language of the transformation, \cm, is an extension of \lv\ with facilities for continuation marks.
+
 Because \cm\ is an extension of \lv, the definition of terms $e$ in \cm\ 
@@
 
-The sole novelty of \cm\ above the plain \lc\ is the ability to annotate and observe the continuation.
-Our evaluation model of \cm, introduced by Pettyjohn et al.
+The sole novelty of \cm\ above the plain \lv\ is the ability to annotate and observe the continuation.
+Our evaluation model of \cm, given by Pettyjohn et al.
 \cite{pettyjohn2005continuations}, utilizes the current evaluation context for that purpose.
@@
 \begin{align}
-E[\app{\abs{x}{e}}{v}]  &\rr E[e[x\leftarrow v]]\\
-E[\wcm{v}{\wcm{v'}{e}}] &\rr E[\wcm{v'}{e}]\\
-E[\wcm{v}{v'}]          &\rr E[v']\\
-E[\ccm]                 &\rr E[\chi(E)]
+E[\app{\abs{x}{e}}{v}]  &\rr E[e[x\leftarrow v]]\label{reduction}\\
+E[\wcm{v}{\wcm{v'}{e}}] &\rr E[\wcm{v'}{e}]\label{tail-enforce}\\
+E[\wcm{v}{v'}]          &\rr E[v']\label{wcm-body}\\
+E[\ccm]                 &\rr E[\chi(E)]\label{ccm-in-context}
 \end{align}
@@
 \end{equation}
-and where $\chi'$ is defined in turn by
+and $\chi'$ is defined in turn by
 \begin{align}
@@
 \end{align}
+
+As an extension of \lv, \cm\ inherits rule \ref{reduction} directly.
+Rule \ref{tail-enforce} expresses the tail-call behavior of \cm\ in that a new mark made in tail position overwrites the previous mark.
+Rule \ref{wcm-body} defines the value of a \scheme'wcm' expression to take on the value of the body.
+Finally, rule \ref{ccm-in-context} defines the value of a \scheme'ccm' expression in terms of its context via the metafunction $\chi$.
+The value of the function is a list of the current continuation marks.
+
 Notice that $\chi'$ is defined in accumulator-passing style which has the effect of ordering the marks from most- to least-recently placed.
@@
 
-\subsection{\cm\ Example}
-
-\scheme|(wcm 4|
-
 \section{Transformation}
 
-A meaning-preserving transformation should commute with evaluation. By this, we mean that the following property should hold for programs $p\in\lambda_{cm}$.
+Language transformations must be meaningful: they must preserve meaning or provide it.
+As transformations, desugarings put the meaning of one construct in terms of another.
+If the meaning of the former was defined only informally, the desugaring offers a new definition as formal as the definition of the latter.
+In this way, desugarings offer a convenient way to provide meaning to a construct and avoid enlarging the language.
+If, however, the sugared language already has a formal meaning, the desugaring must preserve it.
+This will be made more precise shortly.
 
+The transformation $\mathcal{C}$ is defined over each syntactic form of \cm.
+In order to preserve the meaning of \cm, $\mathcal{C}$ must commute with evaluation.
+More precisely, the following property should hold for programs $p\in\lambda_{cm}$:
 \[
@@
 \]
-
-This diagram captures the property that a meaning-preserving transformation commutes with evaluation. More succinctly, 
+or, succinctly,
 \begin{equation}
@@
 
-Our first approach was to use CPS because it treats all data and control flow uniformly [cite Sabry].
+%Our first approach was to use CPS because it treats all data and control flow uniformly [cite Sabry].
 We passed an additional parameter (or two) in the spirit of double-barrelled continuations (but only in spirit).
@@
 
-Almost immediately after, we realized that CPS was wholly unnecessary and landed back where we started with a direct-style definition.
+%Almost immediately after, we realized that CPS was wholly unnecessary and landed back where we started with a direct-style definition.
 
-A comprehensive language transformation need only be defined over each syntactic form of the language.
+%A comprehensive language transformation need only be defined over each syntactic form of the language.
 Thus, our full transformation definition will comprise transformation definitions for each of the five syntactic forms of \cm.
 
-As variables, abstractions, and applications exist in the target language, we imagine that their transformation will be relatively straightforward.
-The essence of \cm\ is that programs within can apply information to and observe information about the context in which they are evaluated.
-The work of the transform is then to propagate contextual information to each expression, manipulating it appropriately as the context changes.
+As variables, abstractions, and applications exist in the target language, their transformation is relatively straightforward.
+The essence of \cm\ is that programs can apply information to and observe information about the context in which they are evaluated.
+This characterization makes the work of the transform clear: to propagate contextual information to each expression, manipulating it appropriately as the context changes.
 
@@
 The treatment of \scheme'ccm' simply needs access to the evaluation of $\chi$ on the current context.
-As \lv\ lacks any facility to observe the context, the transformation needs to explicitly provide this information to each \scheme'ccm' expression.
+As \lv\ lacks any facility to observe the context, the transformation explicitly provides this information to each \scheme'ccm' expression.
 The $\chi$ metafunction is continuous in the sense that a small change in the prefix of the input effects only a small change in the prefix of the output, a consequence of its particular recursive definition.
 This fact means that it is possible to compute $\chi$ incrementally.
-The transform would then need to ensure that each expression had access to information about its own context.
+The transform then ensures that each expression had access to information about its own context.
 
-Finally, we consider the preservation of proper tail-call behavior.
-Even though tail calls exist in \lv--indeed, they are prevalent--we cannot rely on their behavior to honor proper tail-call behavior through the transform.
-Thus, we would need to maintain information about the immediate context of the program: specifically, whether evaluation is occurring directly within the body of a \scheme'wcm' expression or not.
+Finally, we address the preservation of proper tail-call behavior.
+Even though tail calls exist in \lv--indeed, they are prevalent--their behavior does not honor proper tail-call behavior through the transform.
+Thus, we maintain information about the immediate context of the program: specifically, whether evaluation is occurring directly within the body of a \scheme'wcm' expression.
 
-To accomplish each of these, the transform is defined in the spirit of CPS. Recall that in
-CPS, every function receives an additional formal parameter and every call site receives a
-corresponding argument. The definition of $\mathcal{C}$ takes this a step further under
-which every function receives \emph{two} additional formal parameters and every call site
-two additional arguments. As in CPS, the first parameter encapsulates the continuation or,
-for our purposes, the evaluation context. The second encapsulates all contextual
-information observable to the original \cm\ program.
+All together, the transformation adds three arguments to every term $e$: the continuation $k$ representative of the current context $E$; a flag $f$ encoding whether evaluation is occurring with the body of a \scheme'wcm' expression; and the transformed value of $\chi$ for the current context $\C{\chi(E)}$.
 
-\newtheorem{defn}{Definition}
+The formal definition of $\mathcal{C}$ follows.
 
-\newenvironment{schemedefn}[1]{\begin{defn}$#1$}{\end{defn}}
-\newenvironment{namedschemedefn}[2]{\begin{defn}\label{#1}$#2$}{\end{defn}}
+\newenvironment{schemedefn}[1]{\begin{definition}$#1$}{\end{definition}}
+\newenvironment{namedschemedefn}[2]{\begin{definition}\label{#1}$#2$}{\end{definition}}
+
+\setspecialsymbol{lambda}{$\lambda$}
+\setspecialsymbol{e_0}{$e_0$}
+\setspecialsymbol{e_0p}{$e'_0$}
+\setspecialsymbol{Ce_0}{$\C{e_0}$}
+\setspecialsymbol{Ce_0p}{$\C{e'_0}$}
+\setspecialsymbol{e_1}{$e_1$}
+\setspecialsymbol{Ce_1}{$\C{e_1}$}
+\setspecialsymbol{Ce_1p}{$\C{e'_1}$}
+\setspecialsymbol{Cv}{$\C{v}$}
+\setspecialsymbol{v_0}{$v_0$}
+\setspecialsymbol{Cv_0}{$\C{v_0}$}
+\setspecialsymbol{Cv_0p}{$\C{v'_0}$}
+\setspecialsymbol{Cpv_0}{$\Cp{v_0}$}
+\setspecialsymbol{Cpv_1}{$\Cp{v_1}$}
+\setspecialsymbol{v_1}{$v_1$}
+\setspecialsymbol{T}{$\mathcal{A}$}
+\setspecialsymbol{C}{$\mathcal{C}$}
+\setspecialsymbol{Cp}{$\mathcal{C}'$}
+\setspecialsymbol{hole}{$\hole$}
+\setspecialsymbol{bottom}{$\perp$}
+\setspecialsymbol{Ce_0p_x_Cpv_1}{$\C{e'_0}[x\leftarrow\Cp{v_1}]$}
+\setspecialsymbol{e_0p_x_v_1}{$e'_0[x\leftarrow v_1]$}
+\setspecialsymbol{Cz}{$\C{z}$}
+\setspecialsymbol{CE}{$\C{E}$}
+\setspecialsymbol{xiE}{$\xi(E)$}
+\setspecialsymbol{chiE}{$\chi(E)$}
+\setspecialsymbol{CchiE}{$\C{\chi(E)}$}
 
@@
 
-\begin{defn}
+\subsection{Useful Shorthand}
+
+We first define a useful shorthand.
+
+\begin{definition}
 $\true=\abs{x}{\abs{y}{x}}$
-\end{defn}
+\end{definition}
 
-\begin{defn}
+\begin{definition}
 $\false=\abs{x}{\abs{y}{y}}$
-\end{defn}
+\end{definition}
 
-\begin{defn}
+\begin{definition}
 $\mathbf{pair}=\abs{a}{\abs{b}{\abs{z}{\app{\app{z}{a}}{b}}}}$
-\end{defn}
+\end{definition}
 
-\begin{defn}
+\begin{definition}
 $\mathbf{fst}=\abs{p}{\app{p}{\true}}$
-\end{defn}
+\end{definition}
 
-\begin{defn}
+\begin{definition}
 $\mathbf{snd}=\abs{p}{\app{p}{\false}}$
-\end{defn}
+\end{definition}
 
-\begin{defn}
+\begin{definition}
 $\nil=\false$
-\end{defn}
+\end{definition}
+
+\subsection{Definition of $\mathcal{C}$}
+
+The definition of $\mathcal{C}$ over the five syntactic forms of \cm\ follows.
 
@@
 
+The evaluation of operator and operand in an application enlarges the continuation, so marks made within said evaluation will not interfere with pre-existent marks, and this is reflected by the $\false$ flag passed to each.
+Once the arguments are evaluated, the new context is unnecessary and discarded.
+The application itself inherits the flag of the original context.
+
 \begin{schemedefn}{\C{\wcm{e_0}{e_1}}}
@@
 
+The definition over \scheme'wcm' expressions is the only one to make explicit use of the value of the flag.
+Following the evaluation of the mark expression, rule \ref{collapse} is applied, if necessary.
+The continuity of $\chi$ allows us to construct $\chi(E'[\wcm{v_0}{e_1}])$ from $\chi(E)$ where
+
+\begin{equation}
+E'=\begin{cases}
+E' &\text{if $E=E'[\wcm{v'}{\hole}]$ for some $E'$ and $v'$}\\
+E &\text{otherwise}
+\end{cases}
+\end{equation}
+
 \begin{schemedefn}{\C{\ccm}}
@@
 
-\begin{defn}
+The simplicity of the definition over \scheme`ccm` expressions offsets the work performed by the other definitions: it simply returns the list passed in.
+
+\begin{definition}
 $\C{v_0}=\C{\abs{x}{e'_0}}=\abs{k}{\abs{f}{\abs{m}{\app{k}{\abs{x}{\C{e'_0}}}}}}$
-\end{defn}
+\end{definition}
 
-\begin{defn}
+\begin{definition}
 $\C{x}=\abs{k}{\abs{f}{\abs{m}{\app{k}{x}}}}$
-\end{defn}
+\end{definition}
 
-\subsection{Example: Factorial}
+%\subsection{Example: Factorial}
 
-For convenience, we will augment the base language with direct forms for expressing factorial. Define terms $e$ in \lv\ by
-\begin{equation}
-e=\app{e}{e}\,|\,(*\,e\,e)\,|\,\app{\mathrm{pred}}{e}\,|\,\app{\mathrm{zero?}}{e}\,|\,v\,|\,x
-\end{equation}
-and values $v$ by
-\begin{equation}
-v=\abs{x}{e}\,|\,\mathbb{N}
-\end{equation}
-.
+%For convenience, we will augment the base language with direct forms for expressing factorial. Define terms $e$ in \lv\ by
+%\begin{equation}
+%e=\app{e}{e}\,|\,(*\,e\,e)\,|\,\app{\mathrm{pred}}{e}\,|\,\app{\mathrm{zero?}}{e}\,|\,v\,|\,x
+%\end{equation}
+%and values $v$ by
+%\begin{equation}
+%v=\abs{x}{e}\,|\,\mathbb{N}
+%\end{equation}
+%.
 
-We can express the factorial function by the following
-\begin{schemedisplay}
-(Y (lambda (f)
-     (lambda (n)
-       (((zero? n)
-         1)
-        (* n (f (pred n)))))))
-\end{schemedisplay}
+%We can express the factorial function by the following
+%\begin{schemedisplay}
+%(Y (lambda (f)
+%     (lambda (n)
+%       (((zero? n)
+%         1)
+%        (* n (f (pred n)))))))
+%\end{schemedisplay}
 
-and with continuation marks by
-\begin{schemedisplay}
-(Y (lambda (f)
-     (lambda (n)
-       (((zero? n)
-         (print (ccm) 1))
-        (* n (wcm n (f (pred n))))))))
-\end{schemedisplay}
+%and with continuation marks by
+%\begin{schemedisplay}
+%(Y (lambda (f)
+%     (lambda (n)
+%       (((zero? n)
+%         (print (ccm) 1))
+%        (* n (wcm n (f (pred n))))))))
+%\end{schemedisplay}
 

### 2012-09-28 paper.tex
Alpha.

@@
 For example, the ubiquitous JavaScript is an ideal candidate for the addition of continuation marks.
-However, as the \emph{lingua franca} of the web, it has numerous mature implementations which have been heavily optimized.
+However, as the lingua franca of the web, it has numerous mature implementations which have been heavily optimized.
 To add continuation marks to JavaScript amounts to modifying each implementation upstream, to say nothing of amending the JavaScript standard.
@@
 
-Continuation marks respect tail-call optimization (TCO).
-A tail call in a language with TCO does not enlarge the continuation.
+Continuation marks respect tail-call optimization.
+A tail call in a tail-call optimized language does not enlarge the continuation.
 Instead, the previous portion of the execution context is subsumed by the new.
@@
 
-All together, the transformation adds three arguments to every term $e$: the continuation $k$ representative of the current context $E$; a flag $f$ encoding whether evaluation is occurring with the body of a \scheme'wcm' expression; and the transformed value of $\chi$ for the current context $\C{\chi(E)}$.
+All together, the transformation adds three arguments to every term $e$: the continuation $k$ representative of the current context $E$; a flag $f$ encoding whether evaluation is occurring with the body of a \scheme'wcm' expression; and the transformed value $m$ of $\chi$ for the current context $\C{\chi(E)}$.
 
@@
 The definition over \scheme'wcm' expressions is the only one to make explicit use of the value of the flag.
-Following the evaluation of the mark expression, rule \ref{collapse} is applied, if necessary.
-The continuity of $\chi$ allows us to construct $\chi(E'[\wcm{v_0}{e_1}])$ from $\chi(E)$ where
+Following the evaluation of the mark expression, rule \ref{tail-enforce} is applied, if necessary.
+The continuity of $\chi$ allows us to construct $\chi(E''[\wcm{v_0}{e_1}])$ from $\chi(E)$ where
 
 \begin{equation}
-E'=\begin{cases}
+E''=\begin{cases}
 E' &\text{if $E=E'[\wcm{v'}{\hole}]$ for some $E'$ and $v'$}\\
@@
 \setconstant{error}
-\setkeyword{chi chi_p}
+\setkeyword{chi chi_p number apply-reduction-relation*}
 
@@
 
+\setkeyword{meaning-preserved? alpha-eq? transform eval init redex-check}
+
 We can test that the property described by equation \ref{meaning-preservation-property} holds for a given program \scheme'p' with 
@@
 
+\begin{definition}
+\[
+\mathrm{eval}_{cm}(p)=\begin{cases}
+v     &\text{if $p\cmrrs v$}\\
+\perp &\text{if $p\cmrrs\cdots$}
+\end{cases}
+\]
+\end{definition}
+
+\begin{definition}
+\[
+\mathrm{eval}_{v}(p)=\begin{cases}
+v     &\text{if $p\lvrrs v$}\\
+\perp &\text{if $p\lvrrs\cdots$}
+\end{cases}
+\]
+\end{definition}
+
+\newtheorem{maintheorem}{Correctness of $\mathcal{C}$}
+\begin{maintheorem}
+For all programs $p\in\lambda_{cm}$, $\C{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\C{p})$.
+\end{maintheorem}
+
+\section{Back to JavaScript}
+
+With a correct transformation of continuation marks defined over the \lc, it requires little to add continuation marks to an eager, higher-order language such as JavaScript.
+
+We add two keywords to manipulate continuation marks: the familiar \texttt{wcm} and \texttt{ccm}.
+We chose block syntax for \texttt{wcm} to convey that it is a special form, like a conditional, instead of a function-like interface.
+A lone \texttt{ccm} evaluates to the current continuation marks.
+
+Factorial in this JavaScript extension can be implemented like the following:
+\begin{verbatim}
+var fac = function( n ) {
+    if( n == 0 ) {
+        console.log( ccm );
+        return 1;
+    }
+    else {
+        return wcm { n }
+                   { n * fac( n - 1 ) };
+    }
+}
+\end{verbatim}
+
+A transformation of this function desugars these constructs into valid JavaScript.
+\begin{verbatim}
+var fac = function( f, m ) {
+    return function( n ) {
+        if( n == 0 ) {
+            console.log( (function( f, m ) { return m; })( false, m ) );
+            return 1;
+        }
+        else {
+            return (function( f, m ) {
+                return (function( v ) {
+                    return n * fac( false, [ v ].concat( f ? m.slice( 1 ) : m ) )( n - 1 );
+                })( n );
+            })( f, m );
+        }
+    }
+}
+\end{verbatim}
+
 \section{Conclusion}

### 2012-09-29 paper.tex
Addressed a few of Jay's points.

@@
 Statistical profilers sample the stack regularly to record active functions, algebraic steppers observe the stack to represent the evaluation context of an expression, and debuggers naturally require consistent access to the stack.
-Each of these relies on implementation-specific information and must be maintained as the instrumented language undergoes optimizations and ventures across platforms.
+Each scenario relies on implementation-specific information and must be maintained as the instrumented language undergoes optimizations and ventures across platforms.
 This makes these tools brittle and increases the porting cost of the language ecosystem.
-Each of these examples would benefit from a generalized stack-inspection mechanism available within the instrumented language itself.
+Each scenario would benefit from a generalized stack-inspection mechanism available within the instrumented language itself.
 If written in such an enhanced language, each instrument would be more robust, more easily modified, and would port for free.
@@
 
-We begin this paper exposing continuation marks further.
+We begin by explaining continuation marks further.
 We then present a core and enhanced language along with Redex models for each.
@@
 If an annotation, or \emph{mark}, already exists on the youngest portion, it is replaced by the new mark.
-The entire \scheme'wcm' expression takes on \scheme'body-value'.
+This newly-added mark is available within \scheme'body-expr' via \scheme'ccm' during its evaluation.
+Once complete, the entire \scheme'wcm' expression takes on \scheme'body-value'.
 
@@
 
-If the factorial function is defined by
+\subsection{Example}
+
+The traditional, properly-recursive factorial function can be expressed with continuation marks as follows:
 \begin{schemedisplay}
@@
 \end{schemedisplay}
-then the expression \scheme|(fac 2)| evaluates as
+In this definition, a multiplication pends following the recursive call, so the continuation grows with each.
+
+The evaluation of the factorial of 2, expressed by \scheme|(fac 2)|, proceeds as
 \begin{schemedisplay}
@@
 (wcm 2 (* 2 (wcm 1 (* 1 (begin (print (ccm)) 1)))))
+\end{schemedisplay}
+At this point, the term \scheme|(ccm)| is evaluated.
+This happens, in effect, by walking the continuation and building a list of the current continuation marks which appear in the \scheme'mark-value' position of a \scheme'wcm' term.
+In this example, \scheme|(ccm)| evaluates to \scheme|'(1 2)|.
+Evaluation continues as
+\begin{schemedisplay}
 (wcm 2 (* 2 (wcm 1 (* 1 (begin (print '(1 2)) 1)))))
-'(1 2)
 (wcm 2 (* 2 (wcm 1 (* 1 (begin 1)))))
@@
 \end{schemedisplay}
+The visible effect of the call is
+\begin{verbatim}
+> (fac 2)
+(1 2)
+2
+\end{verbatim}
 
-Continuation marks respect tail-call optimization.
-A tail call in a tail-call optimized language does not enlarge the continuation.
-Instead, the previous portion of the execution context is subsumed by the new.
-Conceptually, the continuation does not grow to accommodate new marks; if one is made, the previous is overwritten.
-Thus, with a tail-recursive definition of the factorial function as in
+In a tail-call optimized language, a call made in tail position does not enlarge the continuation.
+Instead, the portion of the evaluation context dedicated to the calling function is repurposed for the called function.
+Continuation marks are also subject to this optimization; if the continuation is marked in tail position, the previous mark there is replaced, if it exists.
+
+This behavior is apparent in the tail-recursive definition of the factorial function, which follows.
 \begin{schemedisplay}
@@
 \end{schemedisplay}
-evaluation of \scheme|(fac 2)| proceeds as
+In this definition, the multiplication is performed before the recursive call and an accumulated value is passed.
+With no pending computation, the evaluation context devoted to this function is no longer necessary, and is subsumed by the continuation mark directive.
+
+The evaluation of the factorial of 2, here expressed by \scheme|(fac 2 1)| proceeds as
 \begin{schemedisplay}
@@
 (wcm 2 (wcm 1 (fac (- 1 1) (* 1 2))))
+\end{schemedisplay}
+At this point, there is one \scheme'wcm' directive directly nested within another.
+Behaviorally, the outer one is collapsed \emph{before} evaluation of the inner body proceeds.
+In effect, it is overwritten.
+Following this, evaluation continues as
+\begin{schemedisplay}
 (wcm 1 (fac (- 1 1) (* 1 2)))
@@
 (wcm 1 (begin (print '(1)) 2))
-'(1)
 (wcm 1 (begin 2))
@@
 \end{schemedisplay}
-and the tail-call behavior is clear.
+The visible effect of the call is
+\begin{verbatim}
+> (fac 2 1)
+(1)
+2
+\end{verbatim}
+and the influence of the tail-call behavior is evident.
 

### 2012-10-01 paper.tex
In the middle of more points.

@@
 Statistical profilers sample the stack regularly to record active functions, algebraic steppers observe the stack to represent the evaluation context of an expression, and debuggers naturally require consistent access to the stack.
-Each scenario relies on implementation-specific information and must be maintained as the instrumented language undergoes optimizations and ventures across platforms.
-This makes these tools brittle and increases the porting cost of the language ecosystem.
-Each scenario would benefit from a generalized stack-inspection mechanism available within the instrumented language itself.
+In each of these scenarios, the instrument relies on implementation-specific information and must be maintained as the instrumented language undergoes optimizations and ventures across platforms.
+This makes these instruments brittle and increases the porting cost of the language ecosystem.
+Each would benefit from a generalized stack-inspection mechanism available within the instrumented language itself.
 If written in such an enhanced language, each instrument would be more robust, more easily modified, and would port for free.
@@
 
-We begin by explaining continuation marks further.
-We then present a core and enhanced language along with Redex models for each.
+We begin by further explaining continuation marks.
+We then formalize them via the introduction of a core and enhanced language with Redex models for each.
 We discuss the necessary properties for a meaningful transform.
@@
 
-The evaluation of the factorial of 2, here expressed by \scheme|(fac 2 1)| proceeds as
+The evaluation of the factorial of 2, here expressed by \scheme|(fac 2 1)|, proceeds as
 \begin{schemedisplay}
@@
 \end{verbatim}
-and the influence of the tail-call behavior is evident.
+from which the overwriting which occurred is evidenced the singleton list obtained by \scheme|(ccm)|.
 
-%Language transformations must be meaningful: they must preserve meaning or provide it. Continuation marks have special meaning in the presence of tail-call optimization.
-% uses of language transformation:
-% meaning of one language in terms of another (desugaring)
-% - reduce core code which simplifies implementations
-% easier to reason/optimize
+Using this mechanism, the principal onus of a statistical profiler, algebraic stepper, or debugger can be met by a straightforward language transformation which recursively wraps each term in a \scheme'wcm' directive annotating the continuation with a description of the wrapped term.
 
-%Language transformations are used for many reasons, but they he transformation must be meaningful: it must either preserve meaning or provide it.
-
-%Language transformations either provide meaning or preserve meaning.
-
-%Language transformation is useful for a variety of purposes.
-It can be used to define one language in terms of another, It can be used to give meaning to the source language in terms of the target language [cite something], provide a characterization more amenable to analysis or optimization.
-Language transformation can also be used to \emph{desugar} a language into a simpler core.
-This
-%Transforming has the following benefits:
-%reduce core code (desugaring, simplifies implementations)
-%give meaning to constructs (semantically, as well as, say, macros)
-%local transformations - macros, for example
-%global transformations - compilers of all kinds, for example
-
-%talk about the common practice of analyzing an extension of the lambda calculus, and a subpractice of doing so via a transformation to the core.
-
-%While continuation marks have had significant treatment, they do not currently enjoy a transformation to the plain $\lambda$-calculus.
-
-%\section{Continuation Marks}
-
-%Continuation marks generalize stack inspection, allowing a bevy of features that rely on some form of it to be defined at the language level.
-This includes algebraic steppers \cite{clements2001modeling}, security policies \cite{clements2004tail}, and aspects \cite{tucker2003pointcuts}.
-By defining such tools at the language level, such tools are more robust, cheaper to port, more easily changed, etc.
-
-%Despite their profound utility, few languages offer continuation marks.
-Their inception lays in Racket \cite{plt-tr1} and they have been added experimentally to \emph{JavaScript} \cite{clements2008implementing}, but their earnest adoption by other languages has been nonexistent.
-
-% transformation means that they can be compiled away and any higher order language can have them
-% transformation also means that they do not need to be built into the language
-
-% talk about composability of this transform to others
-
-%Racket offers desugaring facilities to the user via a powerful macro system. This requires and is limited to local transformations limited to expression boundaries.
-
-%A transformation-based compiler, such as a CPS compiler \cite{appel2007compiling}, effects a global transformation of the program.
-This effect discards original source or transforms it so radically that it is suitable only for the levels beneath the programmer's regular interaction.
+With this understanding of continuation marks, we can now formalize them behaviorally.
 
@@
 
-We begin by formally defining the core language the transformation produces and the extended language over which the transformation is defined.
+In order to formalize continuation marks, we introduce an extension of the \lc with facilities to manipulate continuation marks and present its semantics in the style of Felleisen and Hieb \cite{felleisen1992revised}.
+Because our ultimate goal is a desugaring transform, we first introduce the target language of the transform, the \lc, in the same way.
 
@@
 
-The target language of the transformation, \lv, is simply the call-by-value \lc.
+The target language of the transform is Plotkin's call-by-value \lc, \lv\ \cite{plotkin1975call}.
 
@@
 
-The source language of the transformation, \cm, is an extension of \lv\ with facilities for continuation marks.
+The source language of the transformation is an extension of \lv\ with facilities for continuation marks which we term \cm.
 
@@
 \cite{pettyjohn2005continuations}, utilizes the current evaluation context for that purpose.
-The mutually inductive valuation contexts $E$, $F$ are defined by
+Evaluation contexts $E$ are defined by
 \begin{align}
@@
 \end{align}
-where $\hole$ denotes a ``hole'' in the evaluation context, the ultimate destination of the evaluation of the expression that previously resided there.
+The definition of $E$ is crafted to prevent directly nested \scheme'wcm' directives from occurring in valid evaluation contexts.
 
@@
 \end{align}
-where $\chi$ is defined by
+
+As an extension of \lv, \cm\ inherits rule \ref{reduction} with substitution carried through \scheme'wcm' terms.
+
+If directly-nested \scheme'wcm' directives are introduced into the evaluation context, invalidating it (as in the tail-recursive factorial example), rule \ref{tail-enforce} collapses the outer into the inner.
+
+Rule \ref{wcm-body} defines the value of a \scheme'wcm' expression to take on the value of the body, once obtained.
+
+The definition of rule \ref{ccm-in-context} is given in terms of a metafunction $\chi$ defined by
 \begin{equation}
-\chi(E)=\chi'(E,\mathbf{nil})
+\chi(E)=\chi'(E,\abs{x}{\abs{y}{y}})
 \end{equation}
-and $\chi'$ is defined in turn by
+where $\chi'$ is defined by
 \begin{align}
@@
 \end{align}
+This definition formalizes the intuition given earlier regarding the behavior of \scheme'ccm'.
+Previously, we saw that the value of a \scheme'ccm' directive was a list; however, lists do not strictly exist as values in the \lc, nor in our extension.
+Instead of adding lists as primitive values to \cm, we employ Church-encodings to represent them natively \cite{barendregt1984lambda}.
 
-As an extension of \lv, \cm\ inherits rule \ref{reduction} directly.
-Rule \ref{tail-enforce} expresses the tail-call behavior of \cm\ in that a new mark made in tail position overwrites the previous mark.
-Rule \ref{wcm-body} defines the value of a \scheme'wcm' expression to take on the value of the body.
-Finally, rule \ref{ccm-in-context} defines the value of a \scheme'ccm' expression in terms of its context via the metafunction $\chi$.
-The value of the function is a list of the current continuation marks.
+%The accumulator-passing style definition of $\chi$ has the effect of ordering the marks from most- to least-recently placed.
 
-Notice that $\chi'$ is defined in accumulator-passing style which has the effect of ordering the marks from most- to least-recently placed.
-The direct style formulation reverses this order.
-In the end, the ordering chosen is immaterial; we use this formulation to make the transform simpler.
+With the source and target languages formally specified, we can now examine a language transformation in earnest.
 
@@
 
-Language transformations must be meaningful: they must preserve meaning or provide it.
 As transformations, desugarings put the meaning of one construct in terms of another.
-If the meaning of the former was defined only informally, the desugaring offers a new definition as formal as the definition of the latter.
+If the meaning of the former was given only informally, the desugaring offers a new definition as formal as the definition of the latter.
 In this way, desugarings offer a convenient way to provide meaning to a construct and avoid enlarging the language.
-If, however, the sugared language already has a formal meaning, the desugaring must preserve it.
-This will be made more precise shortly.
+If, however, the sugared language already has a formal meaning, as does \cm, the desugaring must preserve it.
 
+We term our transformation from \cm\ to \lv\ $\mathcal{C}$, as in \emph{compile}, since we are, in essence, compiling away continuation marks.
 The transformation $\mathcal{C}$ is defined over each syntactic form of \cm.
@@
 
-%Our first approach was to use CPS because it treats all data and control flow uniformly [cite Sabry].
-We passed an additional parameter (or two) in the spirit of double-barrelled continuations (but only in spirit).
-Our first attempt passed a pair as the second argument that needed to be deconstructed and constructed in every recursive definition.
-This was a lot of mindless manipulation that had little to do with the essence of the transform.
-In what took far too long a time, we realized the extent of this manipulation and finally confronted the subconscious aversion to passing \emph{two} arguments which we found was baseless.
+The essence of \cm\ is that programs can apply information to and observe information about the context in which they are evaluated.
+Programs in \lv\ have no such facility.
+We can simulate this facility by explicitly passing contextual information to each term as it is evaluated.
+We can define $\mathcal{C}$ to transform \scheme'wcm' directives to manipulate this information and \scheme'ccm' directives to access it.
+In other words, we can transform \cm\ programs to mark-passing style.
 
-%Almost immediately after, we realized that CPS was wholly unnecessary and landed back where we started with a direct-style definition.
+However, marks alone do not account for the tail-call behavior specified by rule \ref{wcm-collapse}.
+Since tail-call behavior is observable (if indirectly) to \cm\ programs, we must also provide to each term information about the position in which it is evaluated.
+Specifically, each \scheme'wcm' directive must be
 
-%A comprehensive language transformation need only be defined over each syntactic form of the language.
-Thus, our full transformation definition will comprise transformation definitions for each of the five syntactic forms of \cm.
+can frame this information in terms of the evaluation context and
 
-As variables, abstractions, and applications exist in the target language, their transformation is relatively straightforward.
-The essence of \cm\ is that programs can apply information to and observe information about the context in which they are evaluated.
-This characterization makes the work of the transform clear: to propagate contextual information to each expression, manipulating it appropriately as the context changes.
+This realization guides the intuition behind the transform: to propagate contextual information to each expression, manipulating it appropriately as the context changes.
 

### 2012-10-01 paper.tex
Making progress.

@@
 
-The semantics of \lv\ are defined simply by
+The reduction relation of \lv\ is defined simply by
 \begin{align}
@@
 
-The semantics of \cm\ are
+The reduction relation of \cm\ is defined by
 \begin{align}
@@
 
+Let $\lvrr$ denote the reduction relation for \lv\ and $\lvrrs$ its transitive closure.
+Similarly, let $\cmrr$ denote the reduction relation for \cm\ and $\cmrrs$ its transitive closure.
+We will use the more specific notation when $\rightarrow$ is potentially ambiguous.
+
 With the source and target languages formally specified, we can now examine a language transformation in earnest.
 
-\section{Transformation}
-
-As transformations, desugarings put the meaning of one construct in terms of another.
-If the meaning of the former was given only informally, the desugaring offers a new definition as formal as the definition of the latter.
-In this way, desugarings offer a convenient way to provide meaning to a construct and avoid enlarging the language.
-If, however, the sugared language already has a formal meaning, as does \cm, the desugaring must preserve it.
-
-We term our transformation from \cm\ to \lv\ $\mathcal{C}$, as in \emph{compile}, since we are, in essence, compiling away continuation marks.
-The transformation $\mathcal{C}$ is defined over each syntactic form of \cm.
-In order to preserve the meaning of \cm, $\mathcal{C}$ must commute with evaluation.
-More precisely, the following property should hold for programs $p\in\lambda_{cm}$:
-\[
-\begin{array}{ccc}
-p & \cmrrs & v\\
-\downarrow_\mathcal{C} & & \downarrow_\mathcal{C}\\
-\C{p} & \lvrrs & \C{v}
-\end{array}
-\]
-or, succinctly,
-\begin{equation}
-\label{meaning-preservation-property}
-p\cmrrs v\implies\C{p}\lvrrs\C{v}
-\end{equation}
-
-The essence of \cm\ is that programs can apply information to and observe information about the context in which they are evaluated.
-Programs in \lv\ have no such facility.
-We can simulate this facility by explicitly passing contextual information to each term as it is evaluated.
-We can define $\mathcal{C}$ to transform \scheme'wcm' directives to manipulate this information and \scheme'ccm' directives to access it.
-In other words, we can transform \cm\ programs to mark-passing style.
-
-However, marks alone do not account for the tail-call behavior specified by rule \ref{wcm-collapse}.
-Since tail-call behavior is observable (if indirectly) to \cm\ programs, we must also provide to each term information about the position in which it is evaluated.
-Specifically, each \scheme'wcm' directive must be
-
-can frame this information in terms of the evaluation context and
-
-This realization guides the intuition behind the transform: to propagate contextual information to each expression, manipulating it appropriately as the context changes.
-
-The treatment of \scheme'wcm' will require care as it is especially sensitive to evaluation order. Assuming termination for all constituent evaluations, consider
-\begin{align*}
-       &E[\wcm{e_0}{\wcm{e_1}{e_2}}]\\
-\cmrrs &E[\wcm{v_0}{\wcm{e_1}{e_2}}]\\
-\cmrrs &E[\wcm{v_0}{\wcm{v_1}{e_2}}]\\
-\cmrr  &E[\wcm{v_1}{e_2}]\\
-\cmrrs &E[\wcm{v_1}{v_2}]\\
-\cmrr  &E[v_2]
-\end{align*}
-We will need to arrange arguments in such a way that the natural call-by-value evaluation order of \lv\ accomplishes correct evaluation order according to the semantics of \cm.
-
-The treatment of \scheme'ccm' simply needs access to the evaluation of $\chi$ on the current context.
-As \lv\ lacks any facility to observe the context, the transformation explicitly provides this information to each \scheme'ccm' expression.
-The $\chi$ metafunction is continuous in the sense that a small change in the prefix of the input effects only a small change in the prefix of the output, a consequence of its particular recursive definition.
-This fact means that it is possible to compute $\chi$ incrementally.
-The transform then ensures that each expression had access to information about its own context.
-
-Finally, we address the preservation of proper tail-call behavior.
-Even though tail calls exist in \lv--indeed, they are prevalent--their behavior does not honor proper tail-call behavior through the transform.
-Thus, we maintain information about the immediate context of the program: specifically, whether evaluation is occurring directly within the body of a \scheme'wcm' expression.
-
-All together, the transformation adds three arguments to every term $e$: the continuation $k$ representative of the current context $E$; a flag $f$ encoding whether evaluation is occurring with the body of a \scheme'wcm' expression; and the transformed value $m$ of $\chi$ for the current context $\C{\chi(E)}$.
-
-The formal definition of $\mathcal{C}$ follows.
-
-\newenvironment{schemedefn}[1]{\begin{definition}$#1$}{\end{definition}}
-\newenvironment{namedschemedefn}[2]{\begin{definition}\label{#1}$#2$}{\end{definition}}
-
+\setkeyword{cons snd}
 \setspecialsymbol{lambda}{$\lambda$}
@@
 \setspecialsymbol{Ce_0p}{$\C{e'_0}$}
+\setspecialsymbol{x_1}{$x_1$}
+\setspecialsymbol{x_n}{$x_n$}
 \setspecialsymbol{e_1}{$e_1$}
+\setspecialsymbol{e_n}{$e_n$}
 \setspecialsymbol{Ce_1}{$\C{e_1}$}
@@
 \setspecialsymbol{PAIR}{$\mathbf{pair}$}
+\setspecialsymbol{CONS}{$\mathbf{cons}$}
 \setspecialsymbol{FST}{$\mathbf{fst}$}
@@
 
+\section{Transformation}
+
+As transformations, desugarings put the meaning of one construct in terms of another.
+If the meaning of the former was given only informally, the desugaring offers a new definition as formal as the definition of the latter.
+In this way, desugarings offer a convenient way to provide meaning to a construct and avoid enlarging the language.
+If, however, the sugared language already has a formal meaning, as does \cm, the desugaring must preserve it.
+
+We will define a transformation from \cm\ to \lv\ and term it $\mathcal{C}$, as in \emph{compile}, since we are, in essence, compiling away continuation marks.
+In order to preserve the meaning of \cm, $\mathcal{C}$ must commute with evaluation.
+More precisely, for programs $p\in\lambda_{cm}$,
+\[
+\begin{array}{ccc}
+p & \cmrrs & v\\
+\downarrow_\mathcal{C} & & \downarrow_\mathcal{C}\\
+\C{p} & \lvrrs & \C{v}
+\end{array}
+\]
+should hold, or, more succinctly,
+\begin{equation}
+\label{meaning-preservation-property}
+p\cmrrs v\implies\C{p}\lvrrs\C{v}
+\end{equation}
+
+\subsection{Intuition}
+
+The essence of \cm\ is that programs can apply information to and observe information about the context in which they are evaluated.
+Programs in \lv\ have no such facility.
+We can simulate this facility by explicitly passing contextual information to each term as it is evaluated.
+We can define $\mathcal{C}$ to transform \scheme'wcm' directives to manipulate this information and \scheme'ccm' directives to access it.
+Intuitively, we can transform \cm\ programs to mark-passing style.
+
+However, marks alone do not account for the tail-call behavior specified by rule \ref{tail-enforce}.
+Since tail-call behavior is observable (if indirectly) by \cm\ programs, we must also provide to each term information about the position in which it is evaluated.
+Specifically, each transformed \scheme'wcm' directive must be notified whether it is evaluated in tail position of an enclosing \scheme'wcm' directive as it must behave specially if so.
+Thus, in addition to passing the current continuation marks, the transform should pass a flag to each term indicating whether it is evaluated in tail position of a \scheme'wcm' directive.
+
+%Since this is a dynamic property of the evaluation, we do not attempt to infer it statically.
+
+These two pieces of information suffice to correctly simulate continuation marks.
+
+%These two functions of the context are sufficient to simulate continuation marks, and they characterize the purpose/approach of the transform: to propagate contextual information to each expression, manipulating it appropriately as the context changes.
+With this in mind, we now formally define $\mathcal{C}$.
+
+%The transformation $\mathcal{C}$ is inductively defined over each syntactic form of \cm.
+
+%The treatment of \scheme'wcm' will require care as it is especially sensitive to evaluation order. Assuming termination for all constituent evaluations, consider
+%\begin{align*}
+%       &E[\wcm{e_0}{\wcm{e_1}{e_2}}]\\
+%\cmrrs &E[\wcm{v_0}{\wcm{e_1}{e_2}}]\\
+%\cmrrs &E[\wcm{v_0}{\wcm{v_1}{e_2}}]\\
+%\cmrr  &E[\wcm{v_1}{e_2}]\\
+%\cmrrs &E[\wcm{v_1}{v_2}]\\
+%\cmrr  &E[v_2]
+%\end{align*}
+%We will need to arrange arguments in such a way that the natural call-by-value evaluation order of \lv\ accomplishes correct evaluation order according to the semantics of \cm.
+
+%The treatment of \scheme'ccm' simply needs access to the evaluation of $\chi$ on the current context.
+As \lv\ lacks any facility to observe the context, the transformation explicitly provides this information to each \scheme'ccm' expression.
+The $\chi$ metafunction is continuous in the sense that a small change in the prefix of the input effects only a small change in the prefix of the output, a consequence of its particular recursive definition.
+This fact means that it is possible to compute $\chi$ incrementally.
+The transform then ensures that each expression had access to information about its own context.
+
+%Finally, we address the preservation of proper tail-call behavior.
+Even though tail calls exist in \lv--indeed, they are prevalent--their behavior does not honor proper tail-call behavior through the transform.
+Thus, we maintain information about the immediate context of the program: specifically, whether evaluation is occurring directly within the body of a \scheme'wcm' expression.
+
+%All together, the transformation adds three arguments to every term $e$: the continuation $k$ representative of the current context $E$; a flag $f$ encoding whether evaluation is occurring with the body of a \scheme'wcm' expression; and the transformed value $m$ of $\chi$ for the current context $\C{\chi(E)}$.
+
+%The formal definition of $\mathcal{C}$ follows.
+
+With this in mind, consider a conceptual transformation \scheme|C[(rator-expr rand-expr)]| as
+\begin{schemedisplay}
+(lambda (flag)
+  (lambda (marks)
+    (let ((rator-value ((C[rator-expr] FALSE) marks))
+          (rand-value ((C[rand-expr] FALSE) marks))
+      (((rator-value rand-value) flag) marks))))
+\end{schemedisplay}
+ignoring for the moment that \scheme'let' is in neither \lv\ or \cm. This definition captures that
+\begin{enumerate}
+\item before evaluation, we expect \scheme'flag' to indicate tail position information and \scheme'marks' to provide a list of the current continuation marks,
+\item we would like to evaluate \scheme'C[rator-expr]' and \scheme'C[rand-expr]' in the same manner, providing to each its contextual information--specifically that neither is evaluated in tail position of a \scheme'wcm' directive and the continuation marks for each are unchanged from the parent context,
+\item and, following evaluation of operator and operand and application, evaluation of the resultant term is performed with the original contextual information.
+\end{enumerate}
+
+Now consider a conceptual transformation \scheme|C[(wcm mark-expr body-expr)]| as
+\begin{schemedisplay}
+(lambda (flag)
+  (lambda (marks)
+    ((C[body-expr] TRUE) (let ((mark-value ((C[mark-expr] FALSE) marks))
+                               (rest-marks (if flag (snd marks) marks)))
+                           (cons mark-value rest-marks)))))
+\end{schemedisplay}
+with similar caveats as the previous case. This definition captures that
+\begin{enumerate}
+\item as in application, we expect \scheme'flag' to indicate tail position information and \scheme'marks' to provide a list of the current continuation marks,
+\item we evaluate \scheme'mark-expr' with correct contextual information,
+\item we discard the first continuation mark of the parent context if evaluation is occurring in tail position of a \scheme'wcm' directive,
+\item and we evaluate \scheme'C[body-expr]' with the correct tail-position flag and current continuation marks.
+\end{enumerate}
+
+Finally, consider the conceptual transformation \scheme|C[(ccm)]| as
+\begin{schemedisplay}
+(lambda (flag)
+  (lambda (marks)
+    marks))
+\end{schemedisplay}
+wherein we reap the fruits of simplicity from our laborious passing: this definition is gratifyingly direct.
+
+The conceptual transformation of variables \scheme'x' and values \scheme'(lambda (x) e)' is straightforward.
+
+We now address the absence of \scheme'let', \scheme'if', \scheme'cons', etc. from \lv.
+
+We can express the \scheme'let' construct in \lv\ with application. An expression such as
+\begin{schemedisplay}
+(let ((x_1 e_1)
+      ...
+      (x_n e_n))
+  e)
+\end{schemedisplay}
+can be interpreted as
+\begin{schemedisplay}
+(...((lambda (x_1) ... (lambda (x_n)
+           e)...) e_1) ...) e_n)
+\end{schemedisplay}
+which is the curried form of
+\begin{schemedisplay}
+((lambda (x_1 ... x_n)
+           e) e_1 ... e_n)
+\end{schemedisplay}
+
+We use the Church encodings of \scheme'if' and conditionals as well as for list-manipulation operators \scheme{cons}, etc. \cite{barendregt1984lambda}.
+
+\subsection{One final concern}
+
+Our choice to keep the core language small by omitting lists as primitive values has the consequence of complicating our transform somewhat.
+Because lists are defined in terms of \lc\ values which are themselves touched by the transform and because of the commutativity property that $\mathcal{C}$ must satisfy, we cannot deal with a list of continuation marks directly--we must instead deal with a transformed list of transformed continuation marks.
+
+\newenvironment{schemedefn}[1]{\begin{definition}$#1$}{\end{definition}}
+\newenvironment{namedschemedefn}[2]{\begin{definition}\label{#1}$#2$}{\end{definition}}
+
 \subsection{Useful Shorthand}
@@
 
-\begin{definition}
-$\true=\abs{x}{\abs{y}{x}}$
-\end{definition}
-
-\begin{definition}
-$\false=\abs{x}{\abs{y}{y}}$
-\end{definition}
-
-\begin{definition}
-$\mathbf{pair}=\abs{a}{\abs{b}{\abs{z}{\app{\app{z}{a}}{b}}}}$
-\end{definition}
-
-\begin{definition}
-$\mathbf{fst}=\abs{p}{\app{p}{\true}}$
-\end{definition}
-
-\begin{definition}
-$\mathbf{snd}=\abs{p}{\app{p}{\false}}$
-\end{definition}
-
-\begin{definition}
-$\nil=\false$
-\end{definition}
+\begin{figure}
+\begin{align*}
+\true         &= \abs{x}{\abs{y}{x}}\\
+\false        &= \abs{x}{\abs{y}{y}}\\
+\mathbf{pair} &= \abs{a}{\abs{b}{\abs{z}{\app{\app{z}{a}}{b}}}}\\
+\mathbf{fst}  &= \abs{p}{\app{p}{\true}}\\
+\mathbf{snd}  &= \abs{p}{\app{p}{\false}}\\
+\nil          &= \false
+\end{align*}
+\label{chuch-encodings}
+\caption{Church encodings of booleans and lists}
+\end{figure}
 
@@
        (((Ce_0
-          (lambda (a)
+          (lambda (v_0)
             (((Ce_1
-               (lambda (b)
-                 ((((a b) k) f) m)))
+               (lambda (v_1)
+                 ((((v_0 v_1) k) f) m)))
               FALSE)
@@
       (((Ce_0
-         (lambda (n) ((lambda (k) 
+         (lambda (v_0) ((lambda (k) 
                         (k (((m (lambda (x) x)) (lambda (z) z)) (lambda (z) z))))
@@
                          (lambda (r) 
-                           (((Ce_1 k) TRUE) C[(PAIR v r)])))))))
+                           (((Ce_1 k) TRUE) C[(PAIR v_0 r)])))))))
         FALSE)

### 2012-10-02 paper.tex
Line 430: nil won't work; that's funny.

@@
 \usepackage{slatex}
+\usepackage{color}
 
@@
 
+\spnewtheorem{plaindef}{Definition}{\bfseries}{\rmfamily}
+
 \setkeyword{wcm ccm with-continuation-mark current-continuation-marks}
@@
 \newcommand{\ccm}[0]{(\mathrm{ccm})}
+\newcommand{\wcmb}[2]{(\mathbf{wcm}\,#1\,#2)}
+\newcommand{\ccmb}[0]{(\mathbf{ccm})}
 \newcommand{\app}[2]{(#1\,#2)}
@@
 \newcommand{\C}[1]{\mathcal{C}[#1]}
+\newcommand{\Cp}[1]{\mathcal{C}'[#1]}
+\newcommand{\evalv}[1]{\mathrm{eval}_v(#1)}
+\newcommand{\evalcm}[1]{\mathrm{eval}_{cm}(#1)}
 
@@
 \]
-should hold, or, more succinctly,
+should hold, or, more concisely,
 \begin{equation}
@@
 \end{equation}
+
+If we define 
+\begin{equation}
+\evalcm{p}=\begin{cases}
+v     &\text{if $p\cmrrs v$}\\
+\perp &\text{if $p\cmrrs\cdots$}
+\end{cases}
+\end{equation}
+and
+\begin{equation}
+\evalv{p}=\begin{cases}
+v     &\text{if $p\lvrrs v$}\\
+\perp &\text{if $p\lvrrs\cdots$}
+\end{cases}
+\end{equation}
+we can state this yet more concisely by $\C{\evalcm{p}}=\evalv{\C{p}}$.
 
@@
 These two pieces of information suffice to correctly simulate continuation marks.
-
-%These two functions of the context are sufficient to simulate continuation marks, and they characterize the purpose/approach of the transform: to propagate contextual information to each expression, manipulating it appropriately as the context changes.
-With this in mind, we now formally define $\mathcal{C}$.
-
-%The transformation $\mathcal{C}$ is inductively defined over each syntactic form of \cm.
-
-%The treatment of \scheme'wcm' will require care as it is especially sensitive to evaluation order. Assuming termination for all constituent evaluations, consider
-%\begin{align*}
-%       &E[\wcm{e_0}{\wcm{e_1}{e_2}}]\\
-%\cmrrs &E[\wcm{v_0}{\wcm{e_1}{e_2}}]\\
-%\cmrrs &E[\wcm{v_0}{\wcm{v_1}{e_2}}]\\
-%\cmrr  &E[\wcm{v_1}{e_2}]\\
-%\cmrrs &E[\wcm{v_1}{v_2}]\\
-%\cmrr  &E[v_2]
-%\end{align*}
-%We will need to arrange arguments in such a way that the natural call-by-value evaluation order of \lv\ accomplishes correct evaluation order according to the semantics of \cm.
-
-%The treatment of \scheme'ccm' simply needs access to the evaluation of $\chi$ on the current context.
-As \lv\ lacks any facility to observe the context, the transformation explicitly provides this information to each \scheme'ccm' expression.
-The $\chi$ metafunction is continuous in the sense that a small change in the prefix of the input effects only a small change in the prefix of the output, a consequence of its particular recursive definition.
-This fact means that it is possible to compute $\chi$ incrementally.
-The transform then ensures that each expression had access to information about its own context.
-
-%Finally, we address the preservation of proper tail-call behavior.
-Even though tail calls exist in \lv--indeed, they are prevalent--their behavior does not honor proper tail-call behavior through the transform.
-Thus, we maintain information about the immediate context of the program: specifically, whether evaluation is occurring directly within the body of a \scheme'wcm' expression.
-
-%All together, the transformation adds three arguments to every term $e$: the continuation $k$ representative of the current context $E$; a flag $f$ encoding whether evaluation is occurring with the body of a \scheme'wcm' expression; and the transformed value $m$ of $\chi$ for the current context $\C{\chi(E)}$.
-
-%The formal definition of $\mathcal{C}$ follows.
 
@@
 
+\subsection{Initiation}
+
+Abstracting terms has the effect of suspending evaluation.
+When an entire program is transformed, all evaluation is suspended.
+It can be initiated by providing contextual information.
+At the top level, the context is empty, so we pass the contextual information for the empty context: \scheme'FALSE' indicating evaluation is \emph{not} occurring in \scheme'wcm' tail position and \scheme'NIL', an empty list of marks.
+
+We can accommodate this by defining a top-level transform $\mathcal{C}'$ in terms of $\mathcal{C}$ by
+\begin{equation}
+\Cp{p}=\app{\app{\C{p}}{\false}}{\nil}
+\end{equation}
+
+and stating our commutativity property as
+\begin{equation}
+\Cp{\evalcm{p}}=\evalv{\Cp{p}}
+\end{equation}
+which is equivalent to
+\begin{equation}
+\app{\app{\C{\evalcm{p}}}{\false}}{\nil}=\evalv{\app{\app{\C{p}}{\false}}{\nil}}
+\end{equation}
+
 \subsection{One final concern}
@@
 
-\newenvironment{schemedefn}[1]{\begin{definition}$#1$}{\end{definition}}
-\newenvironment{namedschemedefn}[2]{\begin{definition}\label{#1}$#2$}{\end{definition}}
+For instance, consider the simple program \scheme|(ccm)|. $\Cp{\evalcm{\ccmb}}=\Cp{\nil}$, but
+\begin{align*}
+  &\evalv{\app{\app{\C{\ccmb}}{\false}}{\nil}}\\
+= &\evalv{\app{\app{\abs{\mathit{flag}}{\abs{\mathit{marks}}{\mathit{marks}}}}{\false}}{\nil}}\\
+= &\nil
+\end{align*}
+This is a problem because $\nil$ is not a primitive value but shorthand for the Church encoding of \scheme'nil', $\abs{x}{\abs{y}{y}}$.
+As an abstraction, it too is transformed.
+It is clear then that we must alter the commutativity requirement, as well as the transform, once more.
+The final commutativity property we seek is
+\begin{equation}
+\app{\app{\C{\evalcm{p}}}{\false}}{\C{\nil}}=\evalv{\app{\app{\C{p}}{\false}}{\C{\nil}}}
+\end{equation}
+
+\newenvironment{schemedefn}[1]{\begin{plaindef}#1}{\end{plaindef}}
+\newenvironment{namedschemedefn}[2]{\begin{plaindef}\label{#1}$#2$}{\end{plaindef}}
 
@@
 
-\begin{schemedefn}{\C{\app{e_0}{e_1}}}
+\begin{schemedefn}{\scheme|C[(rator-expr rand-expr)]|}
 \begin{schemedisplay}
-(lambda (k)
-   (lambda (f)
-     (lambda (m)
-       (((Ce_0
-          (lambda (v_0)
-            (((Ce_1
-               (lambda (v_1)
-                 ((((v_0 v_1) k) f) m)))
+(lambda (kont)
+   (lambda (flag)
+     (lambda (marks)
+       (((C[rator-expr]
+          (lambda (rator-value)
+            (((C[rand-expr]
+               (lambda (rand-value)
+                 ((((rator-value rand-value) kont) flag) marks)))
               FALSE)
@@
 
-\begin{schemedefn}{\C{\wcm{e_0}{e_1}}}
+\begin{schemedefn}{\scheme|C[(wcm mark-expr body-expr)]|}
 \begin{schemedisplay}
-(lambda (k)
-  (lambda (f)
-    (lambda (m)
-      (((Ce_0
-         (lambda (v_0) ((lambda (k) 
-                        (k (((m (lambda (x) x)) (lambda (z) z)) (lambda (z) z))))
+(lambda (kont)
+  (lambda (flag)
+    (lambda (marks)
+      (((C[mark-expr]
+         (lambda (mark-value) ((lambda (k) 
+                                 (k (((marks (lambda (x) x)) (lambda (z) z)) (lambda (z) z))))
                                (lambda (t) 
                                  ((lambda (k) (k ((f C[(SND t)]) t)))
-                         (lambda (r) 
-                           (((Ce_1 k) TRUE) C[(PAIR v_0 r)])))))))
+                                   (lambda (rest) 
+                                     (((C[body-expr] kont) TRUE) C[(CONS mark-value rest)])))))))
         FALSE)
-       m))))
+       marks))))
 \end{schemedisplay}
@@
 
-\begin{schemedefn}{\C{\ccm}}
+\begin{schemedefn}{\scheme|C[(ccm)]|}
 \begin{schemedisplay}
-(lambda (k)
-  (lambda (f)
-    (lambda (m)
-      (((m k) (lambda (z) z)) (lambda (z) z)))))
+(lambda (kont)
+  (lambda (flag)
+    (lambda (marks)
+      (((marks kont) (lambda (z) z)) (lambda (z) z)))))
 \end{schemedisplay}

### 2012-10-02 paper.tex
About to change to direct style.

@@
 
-In order to formalize continuation marks, we introduce an extension of the \lc with facilities to manipulate continuation marks and present its semantics in the style of Felleisen and Hieb \cite{felleisen1992revised}.
+In order to formalize continuation marks, we introduce an extension of the \lc\ with facilities to manipulate continuation marks and present its semantics in the style of Felleisen and Hieb \cite{felleisen1992revised}.
 Because our ultimate goal is a desugaring transform, we first introduce the target language of the transform, the \lc, in the same way.
@@
 \setspecialsymbol{e_n}{$e_n$}
+\setspecialsymbol{k_1}{$k_1$}
+\setspecialsymbol{k_2}{$k_2$}
 \setspecialsymbol{Ce_1}{$\C{e_1}$}
@@
 
-With this in mind, consider a conceptual transformation \scheme|C[(rator-expr rand-expr)]| as
+\subsection{Concept}
+
+The definition of $\mathcal{C}$ entails a defined transformation of each syntactic form of \cm.
+
+With this in mind, consider a conceptual transformation of application, \scheme|C[(rator-expr rand-expr)]|, as
 \begin{schemedisplay}
@@
 
-Now consider a conceptual transformation \scheme|C[(wcm mark-expr body-expr)]| as
+Now consider a conceptual transformation of a \scheme'wcm' directive, \scheme|C[(wcm mark-expr body-expr)]|, as
 \begin{schemedisplay}
@@
 
-Finally, consider the conceptual transformation \scheme|C[(ccm)]| as
+Finally, consider the conceptual transformation of a \scheme'ccm' directive, \scheme|C[(ccm)]|, as
 \begin{schemedisplay}
@@
 \end{schemedisplay}
+We use this fact to construct the transformation and apply $\beta$-rule to simplify terms statically.
 
@@
 Abstracting terms has the effect of suspending evaluation.
-When an entire program is transformed, all evaluation is suspended.
-It can be initiated by providing contextual information.
+When an entire program is transformed, all evaluation is suspended, and awaits arguments representing contextual information.
 At the top level, the context is empty, so we pass the contextual information for the empty context: \scheme'FALSE' indicating evaluation is \emph{not} occurring in \scheme'wcm' tail position and \scheme'NIL', an empty list of marks.
@@
 
-For instance, consider the simple program \scheme|(ccm)|. $\Cp{\evalcm{\ccmb}}=\Cp{\nil}$, but
-\begin{align*}
-  &\evalv{\app{\app{\C{\ccmb}}{\false}}{\nil}}\\
-= &\evalv{\app{\app{\abs{\mathit{flag}}{\abs{\mathit{marks}}{\mathit{marks}}}}{\false}}{\nil}}\\
-= &\nil
-\end{align*}
-This is a problem because $\nil$ is not a primitive value but shorthand for the Church encoding of \scheme'nil', $\abs{x}{\abs{y}{y}}$.
-As an abstraction, it too is transformed.
-It is clear then that we must alter the commutativity requirement, as well as the transform, once more.
-The final commutativity property we seek is
+For this reason, we adjust $\mathcal{C}'$ so that
 \begin{equation}
-\app{\app{\C{\evalcm{p}}}{\false}}{\C{\nil}}=\evalv{\app{\app{\C{p}}{\false}}{\C{\nil}}}
+\Cp{p}=\app{\app{\C{p}}{\false}}{\C{\nil}}
 \end{equation}
+and recognize that a fully-realized transform must be made aware of this.
 
@@
 
-\subsection{Useful Shorthand}
+\subsection{Definition of $\mathcal{C}$}
 
-We first define a useful shorthand.
+The definition of $\mathcal{C}$ over the five syntactic forms of \cm\ follows.
 
@@
 \end{figure}
-
-\subsection{Definition of $\mathcal{C}$}
-
-The definition of $\mathcal{C}$ over the five syntactic forms of \cm\ follows.
 
@@
       (((C[mark-expr]
-         (lambda (mark-value) ((lambda (k) 
-                                 (k (((marks (lambda (x) x)) (lambda (z) z)) (lambda (z) z))))
+         (lambda (mark-value) ((lambda (k_1) 
+                                 (k_1 (((marks (lambda (x) x)) (lambda (z) z)) (lambda (z) z))))
                                (lambda (t) 
-                                 ((lambda (k) (k ((f C[(SND t)]) t)))
+                                 ((lambda (k_2) (k_2 ((f C[(SND t)]) t)))
                                    (lambda (rest) 

### 2012-10-02 paper.tex
Direct style.

@@
 \begin{schemedisplay}
-(lambda (kont)
-   (lambda (flag)
+(lambda (flags)
   (lambda (marks)
-       (((C[rator-expr]
-          (lambda (rator-value)
-            (((C[rand-expr]
-               (lambda (rand-value)
-                 ((((rator-value rand-value) kont) flag) marks)))
-              FALSE)
-             m)))
-         FALSE)
-        m))))
+    (((((C[rator-expr] FALSE) marks)
+       ((C[rand-expr] FALSE) marks))
+      flags)
+     marks)))
 \end{schemedisplay}
@@
 \begin{schemedisplay}
-(lambda (kont)
 (lambda (flag)
   (lambda (marks)
-      (((C[mark-expr]
-         (lambda (mark-value) ((lambda (k_1) 
-                                 (k_1 (((marks (lambda (x) x)) (lambda (z) z)) (lambda (z) z))))
-                               (lambda (t) 
-                                 ((lambda (k_2) (k_2 ((f C[(SND t)]) t)))
-                                   (lambda (rest) 
-                                     (((C[body-expr] kont) TRUE) C[(CONS mark-value rest)])))))))
-        FALSE)
-       marks))))
+    ((C[body-expr] TRUE)
+     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
+       ((C[mark-expr] FALSE) marks))
+      ((flag ((C[(SND marks)] (lambda (z) z)) (lambda (z) z))) marks)))))
 \end{schemedisplay}
@@
 \begin{schemedisplay}
-(lambda (kont)
 (lambda (flag)
   (lambda (marks)
-      (((marks kont) (lambda (z) z)) (lambda (z) z)))))
+    marks))
 \end{schemedisplay}

### 2012-10-02 paper.tex
The list decreases.

@@
 \newcommand{\nil}{\mathbf{nil}}
-\newcommand{\pair}[2]{((\mathbf{pair}\,#1)\,#2)}
+\newcommand{\cons}[2]{((\mathbf{cons}\,#1)\,#2)}
 \newcommand{\eval}[1]{\overline{#1}}
@@
 \false        &= \abs{x}{\abs{y}{y}}\\
-\mathbf{pair} &= \abs{a}{\abs{b}{\abs{z}{\app{\app{z}{a}}{b}}}}\\
-\mathbf{fst}  &= \abs{p}{\app{p}{\true}}\\
+\mathbf{cons} &= \abs{a}{\abs{b}{\abs{z}{\app{\app{z}{a}}{b}}}}\\
 \mathbf{snd}  &= \abs{p}{\app{p}{\false}}\\
@@
 \end{figure}
+
+The formal transformation of application follows the \scheme'let' version exactly except the definitions of \scheme'rator-value' and \scheme'rand-value' are folded directly in.
 
@@
 
-The evaluation of operator and operand in an application enlarges the continuation, so marks made within said evaluation will not interfere with pre-existent marks, and this is reflected by the $\false$ flag passed to each.
+%The evaluation of operator and operand in an application enlarges the continuation, so marks made within said evaluation will not interfere with pre-existent marks, and this is reflected by the $\false$ flag passed to each.
 Once the arguments are evaluated, the new context is unnecessary and discarded.
 The application itself inherits the flag of the original context.
+
+The formal transformation of a \scheme'wcm' directive is also extremely similar to the \scheme'let' version. The definition of \scheme'C[cons]' is unfolded and simplified.
 
@@
        ((C[mark-expr] FALSE) marks))
-      ((flag ((C[(SND marks)] (lambda (z) z)) (lambda (z) z))) marks)))))
+      ((flag Cp[(SND marks)]) marks)))))
 \end{schemedisplay}
@@
 
-The definition over \scheme'wcm' expressions is the only one to make explicit use of the value of the flag.
+%The definition over \scheme'wcm' expressions is the only one to make explicit use of the value of the flag.
 Following the evaluation of the mark expression, rule \ref{tail-enforce} is applied, if necessary.
@@
 
-\begin{equation}
-E''=\begin{cases}
-E' &\text{if $E=E'[\wcm{v'}{\hole}]$ for some $E'$ and $v'$}\\
-E &\text{otherwise}
-\end{cases}
-\end{equation}
+%\begin{equation}
+%E''=\begin{cases}
+%E' &\text{if $E=E'[\wcm{v'}{\hole}]$ for some $E'$ and $v'$}\\
+%E &\text{otherwise}
+%\end{cases}
+%\end{equation}
+
+The \scheme'let' version of the transformation of a \scheme'ccm' directive remains unchanged.
 
@@
 
-The simplicity of the definition over \scheme`ccm` expressions offsets the work performed by the other definitions: it simply returns the list passed in.
+%The simplicity of the definition over \scheme`ccm` expressions offsets the work performed by the other definitions: it simply returns the list passed in.
 
-\begin{definition}
-$\C{v_0}=\C{\abs{x}{e'_0}}=\abs{k}{\abs{f}{\abs{m}{\app{k}{\abs{x}{\C{e'_0}}}}}}$
-\end{definition}
+Like other terms, values are modified to receive contextual information. However, it is ignored.
+% we use this fact a bit
 
-\begin{definition}
-$\C{x}=\abs{k}{\abs{f}{\abs{m}{\app{k}{x}}}}$
-\end{definition}
+\begin{schemedefn}{\scheme|C[v]|=\scheme|C[(lambda (x) e)]|}
+\begin{schemedisplay}
+(lambda (flag)
+  (lambda (marks)
+    (lambda (x) C[e])))
+\end{schemedisplay}
+\end{schemedefn}
+
+% Variables have the property that, when substitution occurs, they reconstitute values.
+
+\begin{schemedefn}{\scheme|C[x]|}
+\begin{schemedisplay}
+(lambda (flag)
+  (lambda (marks)
+    x))
+\end{schemedisplay}
+\end{schemedefn}
 
@@
 \begin{verbatim}
-var fac = function( f, m ) {
+var fac = function( flag, marks ) {
     return function( n ) {
         if( n == 0 ) {
-            console.log( (function( f, m ) { return m; })( false, m ) );
+            console.log( (function( flag, marks ) { return marks; })( false, marks ) );
             return 1;
@@
         else {
-            return (function( f, m ) {
-                return (function( v ) {
-                    return n * fac( false, [ v ].concat( f ? m.slice( 1 ) : m ) )( n - 1 );
-                })( n );
-            })( f, m );
+            return (function( flag, marks ) {
+                return (function( mark_value, rest_marks ) {
+                    return n * fac( false, [ mark_value ].concat( rest_marks ) )( n - 1 );
+                })( n, f ? m.slice( 1 ) : m );
+            })( flag, marks );
         }

### 2012-10-03 paper.tex
All changes made, plus enhancements.

@@
 \newcommand{\Cp}[1]{\mathcal{C}'[#1]}
+\newcommand{\Ch}[1]{\hat{\mathcal{C}}[#1]}
+\newcommand{\Ccps}[1]{\mathcal{C}_{\mathrm{cps}}[#1]}
+\newcommand{\Cpcps}[1]{\mathcal{C}'_{\mathrm{cps}}[#1]}
+\newcommand{\Chcps}[1]{\hat{\mathcal{C}}_{\mathrm{cps}}[#1]}
 \newcommand{\evalv}[1]{\mathrm{eval}_v(#1)}
@@
 
-The target language of the transform is Plotkin's call-by-value \lc, \lv\ \cite{plotkin1975call}.
+The target language of the transform is Plotkin's call-by-value \lc, \lv\ \cite{plotkin1975call}, augmented with natural numbers.
 
@@
 \begin{equation}
-v=\abs{x}{e}
+v=\abs{x}{e}\,|\,n
 \end{equation}
-describes the values of the \lc.
+with $n\in\mathbb{N}$.
 
@@
 \setspecialsymbol{Cp}{$\mathcal{C}'$}
+\setspecialsymbol{Ch}{$\hat{\mathcal{C}}$}
+\setspecialsymbol{Ccps}{$\mathcal{C}_{\mathrm{cps}}$}
+\setspecialsymbol{Cpcps}{$\mathcal{C}'_{\mathrm{cps}}$}
+\setspecialsymbol{Chcps}{$\hat{\mathcal{C}}_{\mathrm{cps}}$}
 \setspecialsymbol{hole}{$\hole$}
@@
 
-The definition of $\mathcal{C}$ entails a defined transformation of each syntactic form of \cm.
+The definition of $\mathcal{C}$ entails transformation over each syntactic form of \cm.
 
@@
 
-We can accommodate this by defining a top-level transform $\mathcal{C}'$ in terms of $\mathcal{C}$ by
+We can accommodate this by defining a top-level transform $\hat{\mathcal{C}}$ in terms of $\mathcal{C}$ by
 \begin{equation}
-\Cp{p}=\app{\app{\C{p}}{\false}}{\nil}
+\Ch{p}=\app{\app{\C{p}}{\false}}{\nil}
 \end{equation}
@@
 \begin{equation}
-\Cp{\evalcm{p}}=\evalv{\Cp{p}}
+\Ch{\evalcm{p}}=\evalv{\Ch{p}}
 \end{equation}
@@
 
-\subsection{One final concern}
+\subsection{Some final subtleties}
 
@@
 
-For this reason, we adjust $\mathcal{C}'$ so that
+Additionally, [values are often truncated, etc.]
+
 \begin{equation}
-\Cp{p}=\app{\app{\C{p}}{\false}}{\C{\nil}}
+\Cp{\abs{x}{e}}=\abs{x}{\C{e}}
+\end{equation}
+
+and we adjust $\hat{\mathcal{C}}$ so that
+\begin{equation}
+\Ch{p}=\app{\app{\C{p}}{\false}}{\Cp{\nil}}
 \end{equation}
@@
 
-The definition of $\mathcal{C}$ over the five syntactic forms of \cm\ follows.
+The definition of $\mathcal{C}$ over the five syntactic forms of \cm\ follows. For reference, figure \ref{church-encodings} contains the particular Church encodings of relevant terms.
 
 \begin{figure}
+\label{church-encodings}
 \begin{align*}
@@
 \end{align*}
-\label{chuch-encodings}
 \caption{Church encodings of booleans and lists}
@@
 \end{schemedefn}
-
-%The evaluation of operator and operand in an application enlarges the continuation, so marks made within said evaluation will not interfere with pre-existent marks, and this is reflected by the $\false$ flag passed to each.
-Once the arguments are evaluated, the new context is unnecessary and discarded.
-The application itself inherits the flag of the original context.
 
@@
 
-%The simplicity of the definition over \scheme`ccm` expressions offsets the work performed by the other definitions: it simply returns the list passed in.
-
 Like other terms, values are modified to receive contextual information. However, it is ignored.
@@
     x))
+\end{schemedisplay}
+\end{schemedefn}
+
+\subsection{CPS transformation}
+
+The use of continuation-passing style as a compiler's intermediate language enables a host of optimization analyses, such as dead-code elimination, and makes the implementation of otherwise-heavyweight features trivial, such as first-class continuations \cite{appel2007compiling}.
+
+We present a continuation mark transformation integrated in continuation-passing style.
+
+\begin{schemedefn}{\scheme|Ccps[(rator-expr rand-expr)]|}
+\begin{schemedisplay}
+(lambda (kont)
+   (lambda (flag)
+     (lambda (marks)
+       (((Ccps[rator-expr]
+          (lambda (rator-value)
+            (((Ccps[rand-expr]
+               (lambda (rand-value)
+                 ((((rator-value rand-value) kont) flag) marks)))
+              FALSE)
+             marks)))
+         FALSE)
+        marks))))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|C[(wcm mark-expr body-expr)]|}
+\begin{schemedisplay}
+(lambda (kont)
+  (lambda (flag)
+    (lambda (marks)
+      (((Ccps[mark-expr]
+          (lambda (mark-value) 
+            ((lambda (rest-marks) 
+               (((Ccps[body-expr] kont) TRUE) (lambda (z) C[((z mark-value) rest-marks)])))
+             ((flag Ch_cps[(SND marks)]) marks))))
+        FALSE)
+       marks))))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|Ccps[(ccm)]|}
+\begin{schemedisplay}
+(lambda (kont)
+  (lambda (flag)
+    (lambda (marks)
+      (kont marks))))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|Ccps[v]|=\scheme|Ccps[(lambda (x) e)]|}
+\begin{schemedisplay}
+(lambda (kont)
+  (lambda (flag)
+    (lambda (marks)
+      (kont (lambda (x) Ccps[e])))))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|Ccps[x]|}
+\begin{schemedisplay}
+(lambda (kont)
+  (lambda (flag)
+    (lambda (marks)
+      (kont x))))
+\end{schemedisplay}
+\end{schemedefn}
+
+We include corresponding definitions for $\mathcal{C}'$ and $\hat{\mathcal{C}}$.
+
+\begin{schemedefn}{\scheme|Cpcps[(lambda (x) e)]|}
+\begin{schemedisplay}
+(lambda (x) Ccps[e])
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|Chcps[p]|}
+\begin{schemedisplay}
+(((Ccps[p] (lambda (x) x)) FALSE) Cpcps[nil])
 \end{schemedisplay}
@@
 %\end{schemedisplay}
+
+\subsection{Example}
+
+Consider the \cm\ program \scheme|(wcm 0 ((lambda (x) (wcm x (ccm))) 1))| which reduces as
+\begin{schemedisplay}
+(wcm 0 ((lambda (x) (wcm x (ccm))) 1))
+(wcm 0 (wcm 1 (ccm)))
+(wcm 1 (ccm))
+(wcm 1 (lambda (z) ((z 1) (lambda (x) (lambda (y) y)))))
+(lambda (z) ((z 1) (lambda (x) (lambda (y) y))))
+\end{schemedisplay}
+and transforms as
+\begin{schemedisplay}
+Ch[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))]
+((C[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))] FALSE) Cp[NIL])
+(((lambda (flag)
+  (lambda (marks)
+    ((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
+     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
+       ((C[0] FALSE) marks))
+      ((flag Cp[(SND marks)]) marks)))))
+ FALSE)
+Cp[NIL])
+(((lambda (flag)
+  (lambda (marks)
+    ((C[(lambda (flags)
+  (lambda (marks)
+    (((((C[(lambda (x) (wcm x (ccm)))] FALSE) marks)
+       ((C[1] FALSE) marks))
+      flags)
+     marks)))] TRUE)
+     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
+       (((lambda (flag) (lambda (marks) 0)) FALSE) marks))
+      ((flag Cp[(SND marks)]) marks)))))
+ FALSE)
+Cp[NIL])
+(((lambda (flag)
+  (lambda (marks)
+    ((C[(lambda (flags)
+  (lambda (marks)
+    ((((((lambda (flag)
+           (lambda (marks)
+             (lambda (x)
+               C[(wcm x (ccm))]))) FALSE) marks)
+       (((lambda (flag) (lambda (marks) 1)) FALSE) marks))
+      flags)
+     marks)))] TRUE)
+     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
+       (((lambda (flag) (lambda (marks) 0)) FALSE) marks))
+      ((flag Cp[(SND marks)]) marks)))))
+ FALSE)
+Cp[NIL])
+(((lambda (flag)
+  (lambda (marks)
+    ((C[(lambda (flags)
+  (lambda (marks)
+    ((((((lambda (flag)
+           (lambda (marks)
+             (lambda (x)
+               (lambda (flag)
+  (lambda (marks)
+    ((C[(ccm)] TRUE)
+     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
+       ((C[x] FALSE) marks))
+      ((flag Cp[(SND marks)]) marks)))))))) FALSE) marks)
+       (((lambda (flag) (lambda (marks) 1)) FALSE) marks))
+      flags)
+     marks)))] TRUE)
+     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
+       (((lambda (flag) (lambda (marks) 0)) FALSE) marks))
+      ((flag Cp[(SND marks)]) marks)))))
+ FALSE)
+Cp[NIL])
+
+(((lambda (flag)
+  (lambda (marks)
+    ((C[(lambda (flags)
+  (lambda (marks)
+    ((((((lambda (flag)
+           (lambda (marks)
+             (lambda (x)
+               (lambda (flag)
+  (lambda (marks)
+    (((lambda (flag) (lambda (marks) marks)) TRUE)
+     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
+       (((lambda (flag) (lambda (marks) x)) FALSE) marks))
+      ((flag Cp[(SND marks)]) marks)))))))) FALSE) marks)
+       (((lambda (flag) (lambda (marks) 1)) FALSE) marks))
+      flags)
+     marks)))] TRUE)
+     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
+       (((lambda (flag) (lambda (marks) 0)) FALSE) marks))
+      ((flag Cp[(SND marks)]) marks)))))
+ FALSE)
+Cp[NIL])
+\end{schemedisplay}
+
+[This example is way too long. Are trivial examples generally uninteresting?]
 
@@
 
-We are now prepared to test various semantic properties of these two languages.
-Of course, we are interested in testing the commutativity of our transform.
-We observe that a mark-passing approach, much like a continuation-passing approach, abstracts terms and so requires a transformed term to be primed with an argument to initiate reduction.
+We are now prepared to test various semantic properties of these two languages. Of course, we are particularly interested in testing the commutativity of our transformation.
+
+%We observe that a mark-passing approach, much like a continuation-passing approach, abstracts terms and so requires a transformed term to be primed with an argument to initiate reduction.
 For this reason, we accompany each transform with an initialization function which primes a term appropriately for the corresponding transform.
@@
 (define (meaning-preserved? p)
-  (alpha-eq? (transform (eval lambdacm p)) (eval lambdav (init (transform p)))))
+  (alpha-eq? (eval lambdav (transform (eval lambdacm p))) (eval lambdav (transform p)))
 \end{schemedisplay}
@@
 \scheme'redex-check' generates random terms according to the grammar of the given language (\scheme'lambdacm') and category (\scheme'e') in search of counterexamples to the predicate.
+It gradually increases the size of the terms it generates, which we found useful in obtaining minimal test cases.
+Once the transformation withstood 10,000 random tests, we considered it for proof (though no incorrect transform withstood more than 500).
 
@@
 
-\begin{definition}
-\[
-\mathrm{eval}_{cm}(p)=\begin{cases}
-v     &\text{if $p\cmrrs v$}\\
-\perp &\text{if $p\cmrrs\cdots$}
-\end{cases}
-\]
-\end{definition}
+The theorem we prove guarantees the same property we have strived to preserve: commutativity. Formally, we would like to prove the correctness of $\hat{\mathcal{C}}$.
 
-\begin{definition}
-\[
-\mathrm{eval}_{v}(p)=\begin{cases}
-v     &\text{if $p\lvrrs v$}\\
-\perp &\text{if $p\lvrrs\cdots$}
-\end{cases}
-\]
-\end{definition}
+%\spnewtheorem*{maintheoremeu}{Correctness of $\hat{\mathcal{C}}$}{a}{x}
+\begin{theorem}
+For all programs $p\in\lambda_{cm}$, $\Ch{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Ch{p})$.
+\end{theorem}
 
-\newtheorem{maintheorem}{Correctness of $\mathcal{C}$}
-\begin{maintheorem}
-For all programs $p\in\lambda_{cm}$, $\C{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\C{p})$.
-\end{maintheorem}
+We do this by overloading $\mathcal{C}$ to accommodate contexts which allows us to formally relate $E[e]$ and $\C{E[e]}$.
+
+This in turn allows us to express the following lemmas.
+
+\begin{lemma}
+For all contexts $E$ and terms $e$, $E[e]\cmrrs v\implies\C{E[e]}\lvrrs\Cp{v}$.
+\end{lemma}
+
+To accomplish this, we first prove by structural induction that for all terms $e$ and context $E$ arbitrary but fixed, $E[e]\cmrrs E[v]\implies\C{E[e]}\lvrrs\C{E[v]}$.
+
+We then prove that for all contexts $E$, $E[v]\cmrrs v'\implies\C{E[v]}\lvrrs\Cp{v'}$.
+
+The theorem follows.
 
@@
 
-With a correct transformation of continuation marks defined over the \lc, it requires little to add continuation marks to an eager, higher-order language such as JavaScript.
+With a correct transformation of continuation marks defined over the \lc, little effort is required to add continuation marks to an eager, higher-order language such as JavaScript.
 
@@
 
+\section{Future Work}
+
+Although we used some level of mechanization in the proof of the theorem, a formal proof assistant, such as Coq \cite{barras1997coq}, would increase confidence out of the gate that the transformation was correct.
+
 \section{Conclusion}

### 2012-10-04 paper.tex
Added todo.
Good changes.

@@
 
-% motivate continuation marks
-% motivate continuation marks in more languages
-% discuss why they are not in more languages
-% discuss approaches to getting them in languages with pros/cons
-% ...culminating in our approach in terms of lambda
-
 Numerous programming language instruments rely on stack inspection to function.
@@
 
-We begin by further explaining continuation marks.
-We then formalize them via the introduction of a core and enhanced language with Redex models for each.
-We discuss the necessary properties for a meaningful transform.
-We present two transformations of continuation marks--one direct and one CPS-like--to the call-by-value \lc.
-We then prove a meaning-preservation theorem.
+We begin by further explaining continuation marks in section \ref{sec-continuation-marks}.
+We then formalize them via the introduction of a core and enhanced language in section \ref{sec-core-and-extension}.
+We discuss the necessary properties of a meaningful transformation between these languages.
+We present two transformations of continuation marks--one direct and one CPS-like--to the call-by-value \lc in section \ref{sec-transformation}.
+This is followed by a lightweight mechanized model for random testing in section \ref{sec-testing} and a proof of a meaning-preservation theorem in section \ref{sec-proof}.
+We finally apply it to JavaScript in section \ref{sec-javascript}.
 
 \section{Continuation Marks}
+\label{sec-continuation-marks}
 
@@
 \section{Language Core and Extension}
+\label{sec-core-and-extension}
 
@@
 \section{Transformation}
+\label{sec-transformation}
 
@@
 Our choice to keep the core language small by omitting lists as primitive values has the consequence of complicating our transform somewhat.
-Because lists are defined in terms of \lc\ values which are themselves touched by the transform and because of the commutativity property that $\mathcal{C}$ must satisfy, we cannot deal with a list of continuation marks directly--we must instead deal with a transformed list of transformed continuation marks.
-
-Additionally, [values are often truncated, etc.]
-
+Because lists are defined in terms of \lc\ values which are themselves touched by the transform and because of the commutativity property that $\mathcal{C}$ must satisfy, we cannot deal with a list of continuation marks directly--we must instead deal with a transformed list of transformed continuation marks, and manipulation of this list within transformed terms must occur at the transformed level.
+
+Additionally, after evaluation, values are ``truncated'' with their leading abstractions applied away.
+For instance, the transformation of the value \scheme|(lambda (x) x)| to \scheme|(lambda (flag) (lambda (marks) (lambda (x) (lambda (flag) (lambda (marks) x)))))| will yield, following evaluation, \scheme|(lambda (x) (lambda (flag) (lambda (marks) x)))|.
+For convenience, we define
 \begin{equation}
@@
 \end{equation}
-
 and we adjust $\hat{\mathcal{C}}$ so that
@@
 \end{equation}
-and recognize that a fully-realized transform must be made aware of this.
 
@@
 \end{schemedefn}
-
-%\subsection{Example: Factorial}
-
-%For convenience, we will augment the base language with direct forms for expressing factorial. Define terms $e$ in \lv\ by
-%\begin{equation}
-%e=\app{e}{e}\,|\,(*\,e\,e)\,|\,\app{\mathrm{pred}}{e}\,|\,\app{\mathrm{zero?}}{e}\,|\,v\,|\,x
-%\end{equation}
-%and values $v$ by
-%\begin{equation}
-%v=\abs{x}{e}\,|\,\mathbb{N}
-%\end{equation}
-%.
-
-%We can express the factorial function by the following
-%\begin{schemedisplay}
-%(Y (lambda (f)
-%     (lambda (n)
-%       (((zero? n)
-%         1)
-%        (* n (f (pred n)))))))
-%\end{schemedisplay}
-
-%and with continuation marks by
-%\begin{schemedisplay}
-%(Y (lambda (f)
-%     (lambda (n)
-%       (((zero? n)
-%         (print (ccm) 1))
-%        (* n (wcm n (f (pred n))))))))
-%\end{schemedisplay}
 
@@
 
+% TODO
+%explain transforms for values and variables
+%fix references
+%add example program
+%fix introduction and handling of CPS
+%talk about proof
+%fix conclusion
+
 \section{Testing}
+\label{sec-testing}
 
@@
 \section{Proof}
+\label{sec-proof}
 
@@
 \section{Back to JavaScript}
+\label{sec-javascript}
 
@@
 
-Factorial in this JavaScript extension can be implemented like the following:
+Tail-call elimination is not part of the JavaScript specification which complicates the treatment of the tail-call behavior.
+We have chosen to perform a simple static analysis which determines if the recursive call is in tail position.
+We simulate proper tail-call behavior by assigning the static value of the flag to reflect this.
+
+A properly-recursive factorial can be implemented in this JavaScript extension as
 \begin{verbatim}
@@
 
-A transformation of this function desugars these constructs into valid JavaScript.
+A transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
 \begin{verbatim}
@@
           return n * fac( false, [ mark_value ].concat( rest_marks ) )( n - 1 );
-                })( n, f ? m.slice( 1 ) : m );
+        })( n, flag ? marks.slice( 1 ) : marks );
       })( flag, marks );
@@
 }
+\end{verbatim}
+
+As expected, the output of a manual initiation of this function is
+\begin{verbatim}
+> fac( false, [] )( 5 )
+[ 1, 2, 3, 4, 5 ]
+120
+\end{verbatim}
+
+A tail-recursive factorial is expressible in this extension as
+\begin{verbatim}
+var fac = function( n, acc ) {
+  if( n == 0 ) {
+    console.log( ccm );
+    return acc;
+  }
+  else {
+    return wcm { n }
+               { fac( n - 1, n * acc ) };
+  }
+}
+\end{verbatim}
+and transforms to
+\begin{verbatim}
+var fac = function( flag, marks ) {
+  return function( n, acc ) {
+    if( n == 0 ) {
+      console.log( (function( flag, marks ) { return marks; })( false, marks ) );
+      return acc;
+    }
+    else {
+      return (function( flag, marks ) {
+        return (function( mark_value, rest_marks ) {
+          return fac( true, [ mark_value ].concat( rest_marks ) )( n - 1, n * acc );
+        })( n, flag ? marks.slice( 1 ) : marks );
+      })( flag, marks );
+    }
+  }
+}
+\end{verbatim}
+
+The output of a manual initiation of this function is
+\begin{verbatim}
+> fac( false, [] )( 5, 1 )
+[ 1 ]
+120
 \end{verbatim}

### 2012-10-04 paper.tex
Springer guidelines compliance.
Mark II.

@@
 \author{Kimball Germane \and Jay McCarthy}
-\institute{Brigham Young University}
+\institute{Brigham Young University, Provo, Utah}
 
@@
 We discuss how such transformations simplify the construction of compilers which treat continuation marks correctly.
-We document the aide of Redex, a domain-specific language for exploring language semantics, in the discovery of correct transformations.
+We document how Redex, a domain-specific language for exploring language semantics, aided the discovery of correct transformations.
 Finally, we verify the transformations with the proof of a meaning-preservation theorem.
@@
 Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail} and to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}.
+[Mention parameters?]
 
@@
 We discuss the necessary properties of a meaningful transformation between these languages.
-We present two transformations of continuation marks--one direct and one CPS-like--to the call-by-value \lc in section \ref{sec-transformation}.
-This is followed by a lightweight mechanized model for random testing in section \ref{sec-testing} and a proof of a meaning-preservation theorem in section \ref{sec-proof}.
+We present two transformations of continuation marks--one direct and one CPS-like--to the call-by-value \lc\ in section \ref{sec-transformation}.
+This is followed by randomized testing using a lightweight mechanized model in section \ref{sec-testing} and a proof of a meaning-preservation theorem in section \ref{sec-proof}.
 We finally apply it to JavaScript in section \ref{sec-javascript}.
@@
 Previously, we saw that the value of a \scheme'ccm' directive was a list; however, lists do not strictly exist as values in the \lc, nor in our extension.
-Instead of adding lists as primitive values to \cm, we employ Church-encodings to represent them natively \cite{barendregt1984lambda}.
-
-%The accumulator-passing style definition of $\chi$ has the effect of ordering the marks from most- to least-recently placed.
-
-Let $\lvrr$ denote the reduction relation for \lv\ and $\lvrrs$ its transitive closure.
-Similarly, let $\cmrr$ denote the reduction relation for \cm\ and $\cmrrs$ its transitive closure.
-We will use the more specific notation when $\rightarrow$ is potentially ambiguous.
-
-With the source and target languages formally specified, we can now examine a language transformation in earnest.
-
-\setkeyword{cons snd}
+Instead of adding lists as primitive values to \cm, we employ Church encodings to represent them natively \cite{barendregt1984lambda}.
+To keep things clear, we preserve the intention of Church-encoded terms by representing them with the useful shorthand of fig.
+\ref{church-encodings}.
+
+\setkeyword{cons snd nil}
 \setspecialsymbol{lambda}{$\lambda$}
@@
 
-\section{Transformation}
-\label{sec-transformation}
-
-As transformations, desugarings put the meaning of one construct in terms of another.
-If the meaning of the former was given only informally, the desugaring offers a new definition as formal as the definition of the latter.
-In this way, desugarings offer a convenient way to provide meaning to a construct and avoid enlarging the language.
-If, however, the sugared language already has a formal meaning, as does \cm, the desugaring must preserve it.
-
-We will define a transformation from \cm\ to \lv\ and term it $\mathcal{C}$, as in \emph{compile}, since we are, in essence, compiling away continuation marks.
-In order to preserve the meaning of \cm, $\mathcal{C}$ must commute with evaluation.
-More precisely, for programs $p\in\lambda_{cm}$,
-\[
-\begin{array}{ccc}
-p & \cmrrs & v\\
-\downarrow_\mathcal{C} & & \downarrow_\mathcal{C}\\
-\C{p} & \lvrrs & \C{v}
-\end{array}
-\]
-should hold, or, more concisely,
-\begin{equation}
-\label{meaning-preservation-property}
-p\cmrrs v\implies\C{p}\lvrrs\C{v}
-\end{equation}
-
-If we define 
-\begin{equation}
-\evalcm{p}=\begin{cases}
-v     &\text{if $p\cmrrs v$}\\
-\perp &\text{if $p\cmrrs\cdots$}
-\end{cases}
-\end{equation}
-and
-\begin{equation}
-\evalv{p}=\begin{cases}
-v     &\text{if $p\lvrrs v$}\\
-\perp &\text{if $p\lvrrs\cdots$}
-\end{cases}
-\end{equation}
-we can state this yet more concisely by $\C{\evalcm{p}}=\evalv{\C{p}}$.
-
-\subsection{Intuition}
-
-The essence of \cm\ is that programs can apply information to and observe information about the context in which they are evaluated.
-Programs in \lv\ have no such facility.
-We can simulate this facility by explicitly passing contextual information to each term as it is evaluated.
-We can define $\mathcal{C}$ to transform \scheme'wcm' directives to manipulate this information and \scheme'ccm' directives to access it.
-Intuitively, we can transform \cm\ programs to mark-passing style.
-
-However, marks alone do not account for the tail-call behavior specified by rule \ref{tail-enforce}.
-Since tail-call behavior is observable (if indirectly) by \cm\ programs, we must also provide to each term information about the position in which it is evaluated.
-Specifically, each transformed \scheme'wcm' directive must be notified whether it is evaluated in tail position of an enclosing \scheme'wcm' directive as it must behave specially if so.
-Thus, in addition to passing the current continuation marks, the transform should pass a flag to each term indicating whether it is evaluated in tail position of a \scheme'wcm' directive.
-
-%Since this is a dynamic property of the evaluation, we do not attempt to infer it statically.
-
-These two pieces of information suffice to correctly simulate continuation marks.
-
-\subsection{Concept}
-
-The definition of $\mathcal{C}$ entails transformation over each syntactic form of \cm.
-
-With this in mind, consider a conceptual transformation of application, \scheme|C[(rator-expr rand-expr)]|, as
-\begin{schemedisplay}
-(lambda (flag)
-  (lambda (marks)
-    (let ((rator-value ((C[rator-expr] FALSE) marks))
-          (rand-value ((C[rand-expr] FALSE) marks))
-      (((rator-value rand-value) flag) marks))))
-\end{schemedisplay}
-ignoring for the moment that \scheme'let' is in neither \lv\ or \cm. This definition captures that
-\begin{enumerate}
-\item before evaluation, we expect \scheme'flag' to indicate tail position information and \scheme'marks' to provide a list of the current continuation marks,
-\item we would like to evaluate \scheme'C[rator-expr]' and \scheme'C[rand-expr]' in the same manner, providing to each its contextual information--specifically that neither is evaluated in tail position of a \scheme'wcm' directive and the continuation marks for each are unchanged from the parent context,
-\item and, following evaluation of operator and operand and application, evaluation of the resultant term is performed with the original contextual information.
-\end{enumerate}
-
-Now consider a conceptual transformation of a \scheme'wcm' directive, \scheme|C[(wcm mark-expr body-expr)]|, as
-\begin{schemedisplay}
-(lambda (flag)
-  (lambda (marks)
-    ((C[body-expr] TRUE) (let ((mark-value ((C[mark-expr] FALSE) marks))
-                               (rest-marks (if flag (snd marks) marks)))
-                           (cons mark-value rest-marks)))))
-\end{schemedisplay}
-with similar caveats as the previous case. This definition captures that
-\begin{enumerate}
-\item as in application, we expect \scheme'flag' to indicate tail position information and \scheme'marks' to provide a list of the current continuation marks,
-\item we evaluate \scheme'mark-expr' with correct contextual information,
-\item we discard the first continuation mark of the parent context if evaluation is occurring in tail position of a \scheme'wcm' directive,
-\item and we evaluate \scheme'C[body-expr]' with the correct tail-position flag and current continuation marks.
-\end{enumerate}
-
-Finally, consider the conceptual transformation of a \scheme'ccm' directive, \scheme|C[(ccm)]|, as
-\begin{schemedisplay}
-(lambda (flag)
-  (lambda (marks)
-    marks))
-\end{schemedisplay}
-wherein we reap the fruits of simplicity from our laborious passing: this definition is gratifyingly direct.
-
-The conceptual transformation of variables \scheme'x' and values \scheme'(lambda (x) e)' is straightforward.
-
-We now address the absence of \scheme'let', \scheme'if', \scheme'cons', etc. from \lv.
-
-We can express the \scheme'let' construct in \lv\ with application. An expression such as
-\begin{schemedisplay}
-(let ((x_1 e_1)
-      ...
-      (x_n e_n))
-  e)
-\end{schemedisplay}
-can be interpreted as
-\begin{schemedisplay}
-(...((lambda (x_1) ... (lambda (x_n)
-           e)...) e_1) ...) e_n)
-\end{schemedisplay}
-which is the curried form of
-\begin{schemedisplay}
-((lambda (x_1 ... x_n)
-           e) e_1 ... e_n)
-\end{schemedisplay}
-We use this fact to construct the transformation and apply $\beta$-rule to simplify terms statically.
-
-We use the Church encodings of \scheme'if' and conditionals as well as for list-manipulation operators \scheme{cons}, etc. \cite{barendregt1984lambda}.
-
-\subsection{Initiation}
-
-Abstracting terms has the effect of suspending evaluation.
-When an entire program is transformed, all evaluation is suspended, and awaits arguments representing contextual information.
-At the top level, the context is empty, so we pass the contextual information for the empty context: \scheme'FALSE' indicating evaluation is \emph{not} occurring in \scheme'wcm' tail position and \scheme'NIL', an empty list of marks.
-
-We can accommodate this by defining a top-level transform $\hat{\mathcal{C}}$ in terms of $\mathcal{C}$ by
-\begin{equation}
-\Ch{p}=\app{\app{\C{p}}{\false}}{\nil}
-\end{equation}
-
-and stating our commutativity property as
-\begin{equation}
-\Ch{\evalcm{p}}=\evalv{\Ch{p}}
-\end{equation}
-which is equivalent to
-\begin{equation}
-\app{\app{\C{\evalcm{p}}}{\false}}{\nil}=\evalv{\app{\app{\C{p}}{\false}}{\nil}}
-\end{equation}
-
-\subsection{Some final subtleties}
-
-Our choice to keep the core language small by omitting lists as primitive values has the consequence of complicating our transform somewhat.
-Because lists are defined in terms of \lc\ values which are themselves touched by the transform and because of the commutativity property that $\mathcal{C}$ must satisfy, we cannot deal with a list of continuation marks directly--we must instead deal with a transformed list of transformed continuation marks, and manipulation of this list within transformed terms must occur at the transformed level.
-
-Additionally, after evaluation, values are ``truncated'' with their leading abstractions applied away.
-For instance, the transformation of the value \scheme|(lambda (x) x)| to \scheme|(lambda (flag) (lambda (marks) (lambda (x) (lambda (flag) (lambda (marks) x)))))| will yield, following evaluation, \scheme|(lambda (x) (lambda (flag) (lambda (marks) x)))|.
-For convenience, we define
-\begin{equation}
-\Cp{\abs{x}{e}}=\abs{x}{\C{e}}
-\end{equation}
-and we adjust $\hat{\mathcal{C}}$ so that
-\begin{equation}
-\Ch{p}=\app{\app{\C{p}}{\false}}{\Cp{\nil}}
-\end{equation}
-
-\newenvironment{schemedefn}[1]{\begin{plaindef}#1}{\end{plaindef}}
-\newenvironment{namedschemedefn}[2]{\begin{plaindef}\label{#1}$#2$}{\end{plaindef}}
-
-\subsection{Definition of $\mathcal{C}$}
-
-The definition of $\mathcal{C}$ over the five syntactic forms of \cm\ follows. For reference, figure \ref{church-encodings} contains the particular Church encodings of relevant terms.
-
 \begin{figure}
@@
 
+%The accumulator-passing style definition of $\chi$ has the effect of ordering the marks from most- to least-recently placed.
+
+Let $\lvrr$ denote the reduction relation for \lv\ and $\lvrrs$ its transitive closure.
+Similarly, let $\cmrr$ denote the reduction relation for \cm\ and $\cmrrs$ its transitive closure.
+We will use the more specific notation when $\rightarrow$ is potentially ambiguous.
+
+With the source and target languages formally specified, we can now examine a language transformation in earnest.
+
+\section{Transformation}
+\label{sec-transformation}
+
+As transformations, desugarings put the meaning of one construct in terms of another.
+If the meaning of the former was given only informally, the desugaring offers a new definition as formal as the definition of the latter.
+In this way, desugarings offer a convenient way to provide meaning to a construct and avoid enlarging the language.
+If, however, the sugared language already has a formal meaning, as does \cm, the desugaring must preserve it.
+
+We will define a transformation from \cm\ to \lv\ and term it $\mathcal{C}$, as in \emph{compile}, since we are, in essence, compiling away continuation marks.
+In order to preserve the meaning of \cm, $\mathcal{C}$ must commute with evaluation.
+More precisely, for programs $p\in\lambda_{cm}$,
+\[
+\begin{array}{ccc}
+p & \cmrrs & v\\
+\downarrow_\mathcal{C} & & \downarrow_\mathcal{C}\\
+\C{p} & \lvrrs & \C{v}
+\end{array}
+\]
+should hold. If we define 
+\begin{equation}
+\evalcm{p}=\begin{cases}
+v     &\text{if $p\cmrrs v$}\\
+\perp &\text{if $p\cmrrs\cdots$}
+\end{cases}
+\end{equation}
+and
+\begin{equation}
+\evalv{p}=\begin{cases}
+v     &\text{if $p\lvrrs v$}\\
+\perp &\text{if $p\lvrrs\cdots$}
+\end{cases}
+\end{equation}
+we can state this more concisely by 
+\begin{equation}
+\C{\evalcm{p}}=\evalv{\C{p}}
+\end{equation}
+
+\subsection{Intuition}
+
+The essence of \cm\ is that programs can apply information to and observe information about the context in which they are evaluated.
+Programs in \lv\ have no such facility.
+We can simulate this facility by explicitly passing contextual information to each term as it is evaluated.
+We can define $\mathcal{C}$ to transform \scheme'wcm' directives to manipulate this information and \scheme'ccm' directives to access it.
+Intuitively, we can transform \cm\ programs to mark-passing style.
+
+However, marks alone do not account for the tail-call behavior specified by rule \ref{tail-enforce}.
+Since tail-call behavior is observable (if indirectly) by \cm\ programs, we must also provide to each term information about the position in which it is evaluated.
+Specifically, each transformed \scheme'wcm' directive must be notified whether it is evaluated in tail position of an enclosing \scheme'wcm' directive as it must behave specially if so.
+Thus, in addition to passing the current continuation marks, the transform should pass a flag to each term indicating whether it is evaluated in tail position of a \scheme'wcm' directive.
+
+%Since this is a dynamic property of the evaluation, we do not attempt to infer it statically.
+
+These two pieces of information suffice to correctly simulate continuation marks.
+
+\subsection{Concept}
+
+The definition of $\mathcal{C}$ entails transformation over each syntactic form of \cm.
+
+With this in mind, consider a conceptual transformation of application, \scheme|C[(rator-expr rand-expr)]|, as
+\begin{schemedisplay}
+(lambda (flag)
+  (lambda (marks)
+    (let ((rator-value ((C[rator-expr] FALSE) marks))
+          (rand-value ((C[rand-expr] FALSE) marks))
+      (((rator-value rand-value) flag) marks))))
+\end{schemedisplay}
+ignoring for the moment that \scheme'let' is in neither \lv\ or \cm. This definition captures that
+\begin{enumerate}
+\item before evaluation, we expect \scheme'flag' to indicate tail position information and \scheme'marks' to provide a list of the current continuation marks,
+\item we would like to evaluate \scheme'C[rator-expr]' and \scheme'C[rand-expr]' in the same manner, providing to each its contextual information--specifically that neither is evaluated in tail position of a \scheme'wcm' directive and the continuation marks for each are unchanged from the parent context,
+\item and, following evaluation of operator and operand and application, evaluation of the resultant term is performed with the original contextual information.
+\end{enumerate}
+
+Now consider a conceptual transformation of a \scheme'wcm' directive, \scheme|C[(wcm mark-expr body-expr)]|, as
+\begin{schemedisplay}
+(lambda (flag)
+  (lambda (marks)
+    ((C[body-expr] TRUE) (let ((mark-value ((C[mark-expr] FALSE) marks))
+                               (rest-marks (if flag (snd marks) marks)))
+                           (cons mark-value rest-marks)))))
+\end{schemedisplay}
+with similar caveats as the previous case. This definition captures that
+\begin{enumerate}
+\item as in application, we expect \scheme'flag' to indicate tail position information and \scheme'marks' to provide a list of the current continuation marks,
+\item we evaluate \scheme'mark-expr' with correct contextual information,
+\item we discard the first continuation mark of the parent context if evaluation is occurring in tail position of a \scheme'wcm' directive,
+\item and we evaluate \scheme'C[body-expr]' with the correct tail-position flag and current continuation marks.
+\end{enumerate}
+
+Finally, consider the conceptual transformation of a \scheme'ccm' directive, \scheme|C[(ccm)]|, as
+\begin{schemedisplay}
+(lambda (flag)
+  (lambda (marks)
+    marks))
+\end{schemedisplay}
+wherein we reap the fruits of simplicity from our laborious passing: this definition is gratifyingly direct.
+
+The conceptual transformation of variables \scheme'x' and values \scheme'(lambda (x) e)' is straightforward.
+
+We now address the absence of \scheme'let', \scheme'if', \scheme'cons', etc. from \lv.
+
+We can express the \scheme'let' construct in \lv\ with application. An expression such as
+\begin{schemedisplay}
+(let ((x_1 e_1)
+      ...
+      (x_n e_n))
+  e)
+\end{schemedisplay}
+can be interpreted as
+\begin{schemedisplay}
+(...((lambda (x_1) ... (lambda (x_n)
+           e)...) e_1) ...) e_n)
+\end{schemedisplay}
+which is the curried form of
+\begin{schemedisplay}
+((lambda (x_1 ... x_n)
+           e) e_1 ... e_n)
+\end{schemedisplay}
+We unfold this characterization of \scheme'let' to guide the construction of $\mathcal{C}$ before simplifying.
+[Drop everything after ``with application.''?]
+
+To achieve \scheme'if' and conditionals as well as list primitives \scheme'cons', \scheme'snd', and \scheme'nil', we use the Church encodings of fig. \ref{church-encodings}.
+
+\subsection{Initiation}
+
+Abstracting terms has the effect of suspending evaluation.
+When an entire program is transformed, all evaluation is suspended, and awaits arguments representing contextual information.
+At the top level, the context is empty, so we pass the contextual information for the empty context: \scheme'FALSE', indicating evaluation is \emph{not} occurring in \scheme'wcm' tail position and \scheme'NIL', an empty list of marks.
+
+We can accommodate this by defining a top-level transform $\hat{\mathcal{C}}$ in terms of $\mathcal{C}$ by
+\begin{equation}
+\Ch{p}=\app{\app{\C{p}}{\false}}{\nil}
+\end{equation}
+
+and stating our commutativity property as
+\begin{equation}
+\label{commutativity-property}
+\Ch{\evalcm{p}}=\evalv{\Ch{p}}
+\end{equation}
+which is equivalent to
+\begin{equation}
+\app{\app{\C{\evalcm{p}}}{\false}}{\nil}=\evalv{\app{\app{\C{p}}{\false}}{\nil}}
+\end{equation}
+
+\subsection{Some Final Subtleties}
+
+Our choice to keep the core language small by omitting lists as primitive values has the consequence of complicating our transform somewhat.
+Because lists are defined in terms of \lc\ values which are themselves touched by the transform and because of the commutativity property that $\mathcal{C}$ must satisfy, we cannot deal with a list of continuation marks directly--we must instead deal with a transformed list of transformed continuation marks, and manipulation of this list within transformed terms must occur at the transformed level.
+
+Additionally, after evaluation, values are ``truncated'' with their leading abstractions applied away.
+For instance, the transformation of the value \scheme|(lambda (x) x)| to \scheme|(lambda (flag) (lambda (marks) (lambda (x) (lambda (flag) (lambda (marks) x)))))| will yield, following evaluation, \scheme|(lambda (x) (lambda (flag) (lambda (marks) x)))|.
+For convenience, we define
+\begin{equation}
+\Cp{\abs{x}{e}}=\abs{x}{\C{e}}
+\end{equation}
+and we adjust $\hat{\mathcal{C}}$ so that
+\begin{equation}
+\Ch{p}=\app{\app{\C{p}}{\false}}{\Cp{\nil}}
+\end{equation}
+
+\newenvironment{schemedefn}[1]{\begin{plaindef}#1}{\end{plaindef}}
+\newenvironment{namedschemedefn}[2]{\begin{plaindef}\label{#1}$#2$}{\end{plaindef}}
+
+\subsection{Definition of $\mathcal{C}$}
+
+Finally, we present the definition of $\mathcal{C}$ over the five syntactic forms of \cm.
+
+\begin{schemedefn}{\scheme|C[(rator-expr rand-expr)]|}
+
 The formal transformation of application follows the \scheme'let' version exactly except the definitions of \scheme'rator-value' and \scheme'rand-value' are folded directly in.
-
-\begin{schemedefn}{\scheme|C[(rator-expr rand-expr)]|}
 \begin{schemedisplay}
@@
 
+\begin{schemedefn}{\scheme|C[(wcm mark-expr body-expr)]|}
+
 The formal transformation of a \scheme'wcm' directive is also extremely similar to the \scheme'let' version. The definition of \scheme'C[cons]' is unfolded and simplified.
-
-\begin{schemedefn}{\scheme|C[(wcm mark-expr body-expr)]|}
 \begin{schemedisplay}
@@
 
-%The definition over \scheme'wcm' expressions is the only one to make explicit use of the value of the flag.
-Following the evaluation of the mark expression, rule \ref{tail-enforce} is applied, if necessary.
-The continuity of $\chi$ allows us to construct $\chi(E''[\wcm{v_0}{e_1}])$ from $\chi(E)$ where
-
-%\begin{equation}
-%E''=\begin{cases}
-%E' &\text{if $E=E'[\wcm{v'}{\hole}]$ for some $E'$ and $v'$}\\
-%E &\text{otherwise}
-%\end{cases}
-%\end{equation}
+\begin{schemedefn}{\scheme|C[(ccm)]|}
 
 The \scheme'let' version of the transformation of a \scheme'ccm' directive remains unchanged.
-
-\begin{schemedefn}{\scheme|C[(ccm)]|}
 \begin{schemedisplay}
@@
 
-Like other terms, values are modified to receive contextual information. However, it is ignored.
-% we use this fact a bit
-
 \begin{schemedefn}{\scheme|C[v]|=\scheme|C[(lambda (x) e)]|}
+
+Like other terms, values are modified to receive contextual information. However, being unaffected by context, values discard this information.
 \begin{schemedisplay}
@@
 
-% Variables have the property that, when substitution occurs, they reconstitute values.
-
 \begin{schemedefn}{\scheme|C[x]|}
+
+Variables have the property that, when substitution occurs, they reconstitute transformed values.
+That is, in the midst of application in $\mathcal{C}$, terms of the form \scheme|(Cp[(lambda (x) x)] Cp[(lambda (y) y)])| appear, reducing to $\C{x}[x\leftarrow \Cp{\abs{y}{y}}]=\C{x[x\leftarrow\abs{y}{y}}=\C{\abs{y}{y}}$.
 \begin{schemedisplay}
@@
 
-\subsection{CPS transformation}
-
-The use of continuation-passing style as a compiler's intermediate language enables a host of optimization analyses, such as dead-code elimination, and makes the implementation of otherwise-heavyweight features trivial, such as first-class continuations \cite{appel2007compiling}.
-
-We present a continuation mark transformation integrated in continuation-passing style.
+\subsection{Definition of $\mathcal{C}$ in CPS}
+
+Many compilers transform source to continuation-passing style (CPS) in the compilation process.
+The use of CPS as an intermediate language enables a host of optimization analyses and makes the implementation of otherwise-heavyweight features trivial, such as first-class continuations \cite{appel2007compiling}.
+
+We present a continuation mark transformation integrated in CPS.
 
@@
 % TODO
-%explain transforms for values and variables
 %fix references
 %add example program
-%fix introduction and handling of CPS
-%talk about proof
 %fix conclusion
@@
 
-The correctness of the transform lies in the property that it commutes.
+The correctness of the transform lies in the property that it commutes with evaluation.
 In order to test for this property, we must construct evaluation models for the source language \cm\ and the target language \lv.
@@
 
-\setkeyword{define-language define-extended-language define-metafunction reduction-relation extend-reduction-relation define-metafunction/extension in-hole hole variable-not-otherwise-mentioned}
-\setkeyword{wcm ccm}
-
-\setspecialsymbol{lambda}{$\lambda$}
-\setspecialsymbol{->}{$\to$}
-\setspecialsymbol{-->}{$\rightarrow$}
-\setspecialsymbol{betav}{$\beta$v}
-\setconstant{error}
-\setkeyword{chi chi_p number apply-reduction-relation*}
-
-\setspecialsymbol{lambdav}{$\lambda$v}
-\setspecialsymbol{lambdav-rr}{$\lambda$v-rr}
-\setspecialsymbol{lambdav-subst}{$\lambda$v-subst}
-
-We first define the form of terms $e$, values $v$, and evaluation contexts $E$.
-
-\begin{schemedisplay}
-(define-language lambdav
-  (e (e e) x v error)
-  (x variable-not-otherwise-mentioned)
-  (v (lambda (x) e) number)
-  (E (E e) (v E) hole))
-\end{schemedisplay}
-
-We then define the reduction relation for \lv. It consists of five rules. The first rule captures the sole semantic rule of \lv; the remaining handle error conditions.
-
-\begin{schemedisplay}
-(define lambdav-rr
-  (reduction-relation lambdav
-   (--> (in-hole E ((lambda (x) e) v))
-        (in-hole E (lambdav-subst x v e))
-        "betav")
-   (--> (in-hole E (number_1 v))
-        (in-hole E error)
-        "error: number in operator position")
-   (--> (in-hole E x)
-        (in-hole E error)
-        "error: unbound identifier")
-   (--> (in-hole E (error e))
-        (in-hole E error)
-        "error in operator")
-   (--> (in-hole E (v error))
-        (in-hole E error)
-        "error in operand")))
-\end{schemedisplay}
-
-The reduction relation requires one metafunction to perform substitution.
-We omit any capture-avoidance facilities to keep the model simple and argue that it is unnecessary as we generate fresh variables in the transform.
-
-\begin{schemedisplay}
-(define-metafunction lambdav
-  lambdav-subst : x v e -> e
-  ;; 1. substitute in application
-  [(lambdav-subst x_1 v_1 (e_1 e_2))
-   ((lambdav-subst x_1 v_1 e_1) (lambdav-subst x_1 v_1 e_2))]
-  ;; 2a. substitute in variable (same)
-  [(lambdav-subst x_1 v_1 x_1)
-   v_1]
-  ;; 2b. substitute in variable (different)
-  [(lambdav-subst x_1 v_1 x_2)
-   x_2]
-  ;; 3a. substitute in abstraction (bound)
-  [(lambdav-subst x_1 v_1 (lambda (x_1) e_1))
-   (lambda (x_1) e_1)]
-  ;; 3b. substitute in abstraction (free)
-  [(lambdav-subst x_1 v_1 (lambda (x_2) e_1))
-   (lambda (x_2) (lambdav-subst x_1 v_1 e_1))]
-  ;; 4. substitute in number
-  [(lambdav-subst x_1 v_1 number_1)
-   number_1]
-  ;; 5. substitute in error
-  [(lambdav-subst x_1 v_1 error)
-   error])
-\end{schemedisplay}
-
-\setspecialsymbol{lambdacm}{$\lambda$cm}
-\setspecialsymbol{lambdacm-rr}{$\lambda$cm-rr}
-\setspecialsymbol{lambdacm-subst}{$\lambda$cm-subst}
-
-We now define the forms of terms $e$, and evaluation contexts $E$ and $F$, inheriting definitions from \lv\ where appropriate.
-
-\begin{schemedisplay}
-(define-extended-language lambdacm lambdav
-  (e .... (wcm e e) (ccm))
-  (E (wcm v F) F)
-  (F (E e) (v E) (wcm E e) hole))
-\end{schemedisplay}
-
-We extend the reduction relation of \lv.
-We begin by redefining application in terms of a substitution function aware of \cm.
-The following three definitions capture the remaining semantic rules of \cm.
-The remaining definitions handle error cases specific to \cm.
-
-\begin{schemedisplay}
-(define lambdacm-rr
-  (extend-reduction-relation lambdav-rr lambdacm
-   (--> (in-hole E ((lambda (x) e) v))
-        (in-hole E (lambdacm-subst x v e))
-        "betav")
-   (--> (in-hole E (wcm v_1 (wcm v_2 e)))
-        (in-hole E (wcm v_2 e))
-        "wcm-collapse")
-   (--> (in-hole E (wcm v_1 v_2))
-        (in-hole E v_2)
-        "wcm")
-   (--> (in-hole E (ccm))
-        (in-hole E (chi E))
-        "chi")
-   (--> (in-hole E (wcm error e))
-        (in-hole E error)
-        "error in wcm mark expression")
-   (--> (in-hole E (wcm v error))
-        (in-hole E error)
-        "error in wcm body expression")))
-\end{schemedisplay}
-
-We extend the definition of substitution to accommodate the additional forms introduced by \cm.
-
-\begin{schemedisplay}
-(define-metafunction/extension lambdav-subst lambdacm
-  lambdacm-subst : x v e -> e
-  ;; 1. substitute in wcm form
-  [(lambdacm-subst x_1 v_1 (wcm e_1 e_2))
-   (wcm (lambdacm-subst x_1 v_1 e_1) (lambdacm-subst x_1 v_1 e_2))]
-  ;; 2. substitute in ccm form
-  [(lambdacm-subst x_1 v_1 (ccm))
-   (ccm)])
-\end{schemedisplay}
-
-The definition of \scheme'chi' and \scheme'chi_p' mirror their definitions exactly.
-
-\begin{schemedisplay}
-(define-metafunction lambdacm
-  chi : E -> v
-  [(chi E) (chi_p E (lambda (x) (lambda (y) y)))])
-
-(define-metafunction lambdacm
-  chi_p : E v -> v
-  [(chi_p hole v_ms)      v_ms]
-  [(chi_p (E e) v_ms)     (chi_p E v_ms)]
-  [(chi_p (v E) v_ms)     (chi_p E v_ms)]
-  [(chi_p (wcm E e) v_ms) (chi_p E v_ms)]
-  [(chi_p (wcm v E) v_ms) (chi_p E (lambda (p) ((p v) v_ms)))])
-\end{schemedisplay}
-
-We are now prepared to test various semantic properties of these two languages. Of course, we are particularly interested in testing the commutativity of our transformation.
+After constructing the models, we are prepared to test various semantic properties of these two languages. Of course, we are particularly interested in testing the commutativity of our transformation.
 
@@
 
-\setkeyword{meaning-preserved? alpha-eq? transform eval init redex-check}
-
-We can test that the property described by equation \ref{meaning-preservation-property} holds for a given program \scheme'p' with 
+\setspecialsymbol{lambdav}{$\lambda$v}
+\setspecialsymbol{lambdacm}{$\lambda$cm}
+\setkeyword{meaning-preserved? alpha-eq? eval c-hat redex-check apply-reduction-relation*}
+
+We can test that the property described by equation \ref{commutativity-property} holds for a given program \scheme'p' with 
 \begin{schemedisplay}
 (define (meaning-preserved? p)
-  (alpha-eq? (eval lambdav (transform (eval lambdacm p))) (eval lambdav (transform p)))
+  (alpha-eq? (eval lambdav (c-hat (eval lambdacm p))) (eval lambdav (c-hat p)))
 \end{schemedisplay}
@@
 It gradually increases the size of the terms it generates, which we found useful in obtaining minimal test cases.
-Once the transformation withstood 10,000 random tests, we considered it for proof (though no incorrect transform withstood more than 500).
+We subjected both the direct and CPS transformation to random testing and each eventually withstood 10,000 random tests.
+Interestingly, no incorrect transformation withstood more than 500 random tests before failing.
 
@@
 
-The theorem we prove guarantees the same property we have strived to preserve: commutativity. Formally, we would like to prove the correctness of $\hat{\mathcal{C}}$.
+With a reasonable intuition and definitions hardened by random testing, we prove correctness.
+
+We have discovered a truly marvelous proof of this, which this paper is too short to contain.
+
+We offer a sketch of the proof here.
+
+We begin by expressing the commutativity property we have sought to preserve as a theorem:
 
@@
 We have chosen to perform a simple static analysis which determines if the recursive call is in tail position.
-We simulate proper tail-call behavior by assigning the static value of the flag to reflect this.
+We simulate proper tail-call behavior by statically assigning the value of the flag to reflect this.
 
@@
 
+\section{Related Work}
+
+Monads, introduced by Moggi \cite{moggi1989computational}, are another way to model semantics.
+Their appeal is significant: they allow one to implement a particular model of computation with with a pure computational logic, such as the \lc.
+Ager et al.
+\cite{ager2003functional} implemented a stack-inspection monad in terms of a lifted state monad, but the size of the continuation must be managed by hand within.
+
 \section{Future Work}

### 2012-10-05 paper.tex
Put example in with JIT expansion.

@@
     ((C[body-expr] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
+     (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)])))
        ((C[mark-expr] FALSE) marks))
-      ((flag Cp[(SND marks)]) marks)))))
+      ((flag Ch[(SND marks)]) marks)))))
 \end{schemedisplay}
@@
             ((lambda (rest-marks) 
-               (((Ccps[body-expr] kont) TRUE) (lambda (z) C[((z mark-value) rest-marks)])))
+               (((Ccps[body-expr] kont) TRUE) Cp[((CONS mark-value) rest-marks)])))
              ((flag Ch_cps[(SND marks)]) marks))))
@@
 Ch[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))]
+
 ((C[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))] FALSE) Cp[NIL])
+
 (((lambda (flag)
@@
       ((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
-       ((C[0] FALSE) marks))
-      ((flag Cp[(SND marks)]) marks)))))
- FALSE)
+       (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+         ((C[0] FALSE) marks)) ((flag Ch[(SND marks)]) marks)))))
+  FALSE) Cp[NIL])
+
+((lambda (marks)
+   ((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
+    (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+      ((C[0] FALSE) marks)) ((FALSE Ch[(SND marks)]) marks))))
  Cp[NIL])
+
+((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   ((C[0] FALSE) Cp[NIL])) ((FALSE Ch[(SND NIL)]) Cp[NIL])))
+
+((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   (((lambda (flag) (lambda (marks) 0)) FALSE) Cp[NIL])) ((FALSE Ch[(SND NIL)]) Cp[NIL])))
+
+((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   ((lambda (marks) 0) Cp[NIL])) ((FALSE Ch[(SND NIL)]) Cp[NIL])))
+
+((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   0) ((FALSE Ch[(SND NIL)]) Cp[NIL])))
+
+((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
+ ((lambda (rest-marks) Cp[((CONS 0) rest-marks)])
+  ((FALSE Ch[(SND NIL)]) Cp[NIL])))
+
+((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
+ ((lambda (rest-marks) Cp[((CONS 0) rest-marks)])
+  Cp[NIL]))
+
+((C[((lambda (x) (wcm x (ccm))) 1)] TRUE) Cp[((CONS 0) NIL)])
+
 (((lambda (flag)
-  (lambda (marks)
-    ((C[(lambda (flags)
     (lambda (marks)
@@
          ((C[1] FALSE) marks))
-      flags)
-     marks)))] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
-       (((lambda (flag) (lambda (marks) 0)) FALSE) marks))
-      ((flag Cp[(SND marks)]) marks)))))
- FALSE)
-Cp[NIL])
-(((lambda (flag)
-  (lambda (marks)
-    ((C[(lambda (flags)
-  (lambda (marks)
-    ((((((lambda (flag)
-           (lambda (marks)
-             (lambda (x)
-               C[(wcm x (ccm))]))) FALSE) marks)
-       (((lambda (flag) (lambda (marks) 1)) FALSE) marks))
-      flags)
-     marks)))] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
-       (((lambda (flag) (lambda (marks) 0)) FALSE) marks))
-      ((flag Cp[(SND marks)]) marks)))))
- FALSE)
-Cp[NIL])
-(((lambda (flag)
-  (lambda (marks)
-    ((C[(lambda (flags)
-  (lambda (marks)
-    ((((((lambda (flag)
-           (lambda (marks)
-             (lambda (x)
-               (lambda (flag)
-  (lambda (marks)
-    ((C[(ccm)] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
-       ((C[x] FALSE) marks))
-      ((flag Cp[(SND marks)]) marks)))))))) FALSE) marks)
-       (((lambda (flag) (lambda (marks) 1)) FALSE) marks))
-      flags)
-     marks)))] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
-       (((lambda (flag) (lambda (marks) 0)) FALSE) marks))
-      ((flag Cp[(SND marks)]) marks)))))
- FALSE)
-Cp[NIL])
+        flag)
+       marks))) TRUE) Cp[((CONS 0) NIL)])
+
+(((((C[(lambda (x) (wcm x (ccm)))] FALSE) Cp[((CONS 0) NIL)])
+   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
+  TRUE) Cp[((CONS 0) NIL)])
+
+((((((lambda (flag) (lambda (marks) (lambda (x) C[(wcm x (ccm))]))) FALSE) Cp[((CONS 0) NIL)])
+   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
+  TRUE) Cp[((CONS 0) NIL)])
+
+(((((lambda (marks) (lambda (x) C[(wcm x (ccm))])) Cp[((CONS 0) NIL)])
+   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
+  TRUE) Cp[((CONS 0) NIL)])
+
+((((lambda (x) C[(wcm x (ccm))])
+   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
+  TRUE) Cp[((CONS 0) NIL)])
+
+((((lambda (x) C[(wcm x (ccm))])
+   (((lambda (flag) (lambda (marks) 1)) FALSE) Cp[((CONS 0) NIL)]))
+  TRUE) Cp[((CONS 0) NIL)])
+
+((((lambda (x) C[(wcm x (ccm))])
+   ((lambda (marks) 1) Cp[((CONS 0) NIL)]))
+  TRUE) Cp[((CONS 0) NIL)])
+
+((((lambda (x) C[(wcm x (ccm))])
+   1) TRUE) Cp[((CONS 0) NIL)])
+
+((C[(wcm 1 (ccm))]
+   TRUE) Cp[((CONS 0) NIL)])
 
@@
     (lambda (marks)
-    ((C[(lambda (flags)
-  (lambda (marks)
-    ((((((lambda (flag)
-           (lambda (marks)
-             (lambda (x)
-               (lambda (flag)
-  (lambda (marks)
-    (((lambda (flag) (lambda (marks) marks)) TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
-       (((lambda (flag) (lambda (marks) x)) FALSE) marks))
-      ((flag Cp[(SND marks)]) marks)))))))) FALSE) marks)
-       (((lambda (flag) (lambda (marks) 1)) FALSE) marks))
-      flags)
-     marks)))] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) (lambda (z) C[((z mark-value) rest-marks)])))
-       (((lambda (flag) (lambda (marks) 0)) FALSE) marks))
-      ((flag Cp[(SND marks)]) marks)))))
- FALSE)
-Cp[NIL])
+      ((C[(ccm)] TRUE)
+       (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+         ((C[1] FALSE) marks))
+        ((flag Ch[(SND marks)]) marks)))))
+  TRUE) Cp[((CONS 0) NIL)])
+
+((lambda (marks)
+   ((C[(ccm)] TRUE)
+    (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+      ((C[1] FALSE) marks))
+     ((TRUE Ch[(SND marks)]) marks))))
+ Cp[((CONS 0) NIL)])
+
+((C[(ccm)] TRUE)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
+  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
+
+(((lambda (flag) (lambda (marks) marks))
+  TRUE)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
+  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
+
+((lambda (marks) marks)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
+  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
+
+((lambda (marks) marks)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   (((lambda (flag) (lambda (marks) 1)) FALSE) Cp[((CONS 0) NIL)]))
+  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
+
+((lambda (marks) marks)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   ((lambda (marks) 1) Cp[((CONS 0) NIL)]))
+  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
+
+((lambda (marks) marks)
+ (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+   1)
+  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
+
+((lambda (marks) marks)
+ ((lambda (rest-marks) Cp[((CONS 1) rest-marks)])
+  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
+
+((lambda (marks) marks)
+ ((lambda (rest-marks) Cp[((CONS 1) rest-marks)])
+  Ch[(SND ((CONS 0) NIL))]))
+
+((lambda (marks) marks)
+ ((lambda (rest-marks) Cp[((CONS 1) rest-marks)])
+  Cp[NIL]))
+
+((lambda (marks) marks)
+ Cp[((CONS 1) NIL)])
+
+Cp[((CONS 1) NIL)]
 \end{schemedisplay}

### 2012-10-05 paper.tex
CM monad in it.

@@
 
+Expressing continuation marks in a monadic way is simple enough. In Haskell,
+\begin{verbatim}
+module CM where
+import Control.Monad
+
+data CM m a = CM ((Bool,[m]) -> a)
+
+instance Monad (CM m) where
+  return x = CM (\_ -> x)
+  (CM m) >>= f = CM (\(flag,vs) -> let (CM m') = f (m (False,vs)) in m' (flag,vs))
+
+wcm :: m -> CM m a -> CM m a
+wcm v (CM m) = CM (\(f,vs) -> m (True, v:(if f then (tail vs) else vs)))
+
+ccm :: CM m [m]
+ccm = CM (\(_,vs) -> vs)
+
+runCM :: CM m a -> a
+runCM (CM m) = m (False,[])
+\end{verbatim}
+
+\begin{verbatim}
+module Main where
+import CM
+
+fac :: Int -> CM Int (Int,[Int])
+fac 0 = do
+  vs <- ccm;
+  return (1,vs)
+
+fac n = do
+  (acc,vs) <- wcm n (fac (n - 1));
+  return (n * acc,vs)
+
+fac_tr :: Int -> Int -> CM Int (Int,[Int])
+fac_tr 0 acc = do
+  vs <- ccm;
+  return (acc,vs)
+
+fac_tr n acc = wcm n (fac_tr (n - 1) (n * acc))
+
+main = do
+  print $ runCM (fac 5);
+  print $ runCM (fac_tr 5 1)
+\end{verbatim}
+
+\begin{verbatim}
+(120,[1,2,3,4,5])
+(120,[1])
+\end{verbatim}
+[Say more with less.]
+
 \section{Future Work}

### 2012-10-05 paper.tex
Small abstract change.

--- 2012-10-05 paper.tex
@@
 We document how Redex, a domain-specific language for exploring language semantics, aided the discovery of correct transformations.
-Finally, we verify the transformations with the proof of a meaning-preservation theorem.
+We verify the transformations with the proof of a meaning-preservation theorem.
+Finally, we apply the transformation to JavaScript.
 \end{abstract}
@@
 Tail-call elimination is not part of the JavaScript specification which complicates the treatment of the tail-call behavior.
-We have chosen to perform a simple static analysis which determines if the recursive call is in tail position.
-We simulate proper tail-call behavior by statically assigning the value of the flag to reflect this.
+We simulate proper tail-call behavior by using the flag to encode whether a function call is in tail position.
 
@@
 
-A transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
+A direct-style transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
 \begin{verbatim}
@@
 
+In this example, the API to compute factorials has changed, but the change is localized.
+With top-level control of a program, we can transform the relevant part of the system, leaving the rest the same.
+This preserves interoperability with foreign functions, e.g., third-party JavaScript libraries.
+
 \section{Related Work}
@@
 Their appeal is significant: they allow one to implement a particular model of computation with with a pure computational logic, such as the \lc.
+
+Expressing continuation marks in a monadic way is simple enough. In Haskell,
+\begin{verbatim}
+data CM m a = CM ((Bool,[m]) -> a)
+
+instance Monad (CM m) where
+  return x = CM (\_ -> x)
+  (CM m) >>= f = CM (\(flag,vs) -> let (CM m') = f (m (False,vs)) in m' (flag,vs))
+
+wcm :: m -> CM m a -> CM m a
+wcm v (CM m) = CM (\(f,vs) -> m (True, v:(if f then (tail vs) else vs)))
+
+ccm :: CM m [m]
+ccm = CM (\(_,vs) -> vs)
+
+runCM :: CM m a -> a
+runCM (CM m) = m (False,[])
+\end{verbatim}
+is one way, using our approach.
+The factorial functions leveraging this implementation exhibit the correct continuation mark behavior but this is \emph{not} a monad.
+Specifically, the right identity law
+\begin{verbatim}
+m >>= return = m
+\end{verbatim}
+does not hold since \texttt{CM} conceptually adds a stack frame in bind (\texttt{>>=}).
 Ager et al.
-\cite{ager2003functional} implemented a stack-inspection monad in terms of a lifted state monad, but the size of the continuation must be managed by hand within.
+\cite{ager2003functional} implemented a similar stack-inspection monad in terms of a lifted state monad, but resorted to manual management of the stack frames.
 

### 2012-10-09 paper.tex
Changes.

@@
 \setspecialsymbol{CE}{$\C{E}$}
+\setspecialsymbol{XI}{$\xi$}
 \setspecialsymbol{xiE}{$\xi(E)$}
@@
     ((C[body-expr] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)])))
+     (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
        ((C[mark-expr] FALSE) marks))
@@
             ((lambda (rest-marks) 
-               (((Ccps[body-expr] kont) TRUE) Cp[((CONS mark-value) rest-marks)])))
-             ((flag Ch_cps[(SND marks)]) marks))))
+               (((Ccps[body-expr] kont) TRUE) Cpcps[((CONS mark-value) rest-marks)]))
+             ((flag Chcps[(SND marks)]) marks))))
         FALSE)
@@
 
-% TODO
-%fix references
-%add example program
-%fix conclusion
-
 \section{Testing}
@@
 
-With a reasonable intuition and definitions hardened by random testing, we prove correctness.
+With a reasonable intuition and definitions hardened by random testing, we prove correctness of the continuation-passing variant of the transform $\mathcal{C}_{\mathrm{cps}}$.
 
@@
 We do this by overloading $\mathcal{C}$ to accommodate contexts which allows us to formally relate $E[e]$ and $\C{E[e]}$.
+
+\begin{schemedefn}{\scheme|hole|}
+\begin{schemedisplay}
+(lambda (value)
+  value)
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|E[(hole rand-value)]|}
+\begin{schemedisplay}
+(lambda (rator-value)
+  (((Ccps[rand-expr]
+     (lambda (rand-value)
+       ((((rator-value rand-value) Ccps[E]) xiE) Cpcps[chiE])))
+    FALSE)
+  Cpcps[chiE]))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|E[(v_0 hole)]|}
+\begin{schemedisplay}
+(lambda (rand-value)
+  ((((v_0 rand-value) Ccps[E]) xiE) Cpcps[chiE]))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|E[(wcm hole body-expr)]|}
+\begin{schemedisplay}
+(lambda (mark-value) 
+  ((lambda (rest-marks) 
+      (((Ccps[body-expr] Ccps[E]) TRUE) Cpcps[((CONS mark-value) rest-marks)]))
+    ((xiE Chcps[(SND chiE)]) Cpcps[chiE])))
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|E[(wcm v_0 hole)]|}
+\begin{schemedisplay}
+Ccps[E]
+\end{schemedisplay}
+\end{schemedefn}
+
+\begin{schemedefn}{\scheme|Ccps[E[e]]|}
+\begin{schemedisplay}
+(((Ccps[e] C[E]) xiE) Cpcps[E])
+\end{schemedisplay}
+\end{schemedefn}
+
+From this definition, it is apparent that \scheme|Chcps[p]|=\scheme|(((Ccps[p] Ccps[hole]) XI(hole)) Cpcps[hole])|.
 

### 2012-10-09 paper.tex
Better example intro.

@@
 
-Consider the \cm\ program \scheme|(wcm 0 ((lambda (x) (wcm x (ccm))) 1))| which reduces as
+To better illustrate what the transformation does, we step through the reduction of a program which exhibits its more interesting aspects.
+One \cm\ program suited to this purpose is \scheme|(wcm 0 ((lambda (x) (wcm x (ccm))) 1))|.
+It reduces according to \cm\ semantics as
 \begin{schemedisplay}
@@
 \end{schemedisplay}
-and transforms as
+
+Now consider the transformation of the same program. We apply the transformation just-in-time as we reduce which has the effect of reducing term size and promoting clarity.
 \begin{schemedisplay}

### 2012-10-10 paper.tex
Small corrections.
Small correction.
Latest draft.

@@
 Behaviorally, the outer one is collapsed \emph{before} evaluation of the inner body proceeds.
-In effect, it is overwritten.
+In essence, it is overwritten.
 Following this, evaluation continues as
@@
 
-Abstracting terms has the effect of suspending evaluation.
-When an entire program is transformed, all evaluation is suspended, and awaits arguments representing contextual information.
+In our call-by-value language, abstracting terms has the effect of suspending evaluation.
+When an entire program is transformed, all evaluation is suspended, awaiting arguments representing contextual information.
 At the top level, the context is empty, so we pass the contextual information for the empty context: \scheme'FALSE', indicating evaluation is \emph{not} occurring in \scheme'wcm' tail position and \scheme'NIL', an empty list of marks.
@@
 
-The formal transformation of a \scheme'wcm' directive is also extremely similar to the \scheme'let' version. The definition of \scheme'C[cons]' is unfolded and simplified.
+The formal transformation of a \scheme'wcm' directive is also extremely similar to the \scheme'let' version.
+The Church-encoded conditional eagerly evaluates both branches, but this still achieves correct behavior.
 \begin{schemedisplay}
@@
     ((C[body-expr] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
+     (((lambda (mark-value) (lambda (rest-marks) Ch[((CONS mark-value) rest-marks)]))
        ((C[mark-expr] FALSE) marks))
@@
             ((lambda (rest-marks) 
-               (((Ccps[body-expr] kont) TRUE) Cpcps[((CONS mark-value) rest-marks)]))
+               (((Ccps[body-expr] kont) TRUE) Chcps[((CONS mark-value) rest-marks)]))
              ((flag Chcps[(SND marks)]) marks))))
@@
 
-Now consider the transformation of the same program. We apply the transformation just-in-time as we reduce which has the effect of reducing term size and promoting clarity.
+Now consider the reduction of the same program transformed. We apply the transformation just-in-time as we reduce to reduce term size and promote clarity.
 \begin{schemedisplay}
@@
 \begin{theorem}
-For all programs $p\in\lambda_{cm}$, $\Ch{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Ch{p})$.
+For all programs $p\in\lambda_{cm}$, $\Chcps{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Chcps{p})$.
 \end{theorem}
 
-We do this by overloading $\mathcal{C}$ to accommodate contexts which allows us to formally relate $E[e]$ and $\C{E[e]}$.
-
-\begin{schemedefn}{\scheme|hole|}
+We do this by overloading $\mathcal{C}_{\mathrm{cps}}$ to accommodate contexts which allows us to formally relate $E[e]$ and $\Ccps{E[e]}$. We first define
+\begin{plaindef}
+\[
+\xi(E)=\begin{cases}
+\mathbf{true} &\text{if $E=E'[\wcm{v'}{\hole}]$ for some $E'$ and $v'$}\\
+\mathbf{false} &\text{otherwise}
+\end{cases}
+\]
+\end{plaindef}
+to denote the \scheme'flags' argument and assume that the \scheme'marks' argument is \scheme|Cpcps[chiE]|. We can now define $\mathcal{C}_{\mathrm{cps}}$ over contexts:
+
+\begin{schemedefn}{\scheme|Ccps[hole]|}
 \begin{schemedisplay}
@@
 
-\begin{schemedefn}{\scheme|E[(hole rand-value)]|}
+\begin{schemedefn}{\scheme|Ccps[E[(hole rand-value)]]|}
 \begin{schemedisplay}
@@
 
-\begin{schemedefn}{\scheme|E[(v_0 hole)]|}
+\begin{schemedefn}{\scheme|Ccps[E[(v_0 hole)]]|}
 \begin{schemedisplay}
@@
 
-\begin{schemedefn}{\scheme|E[(wcm hole body-expr)]|}
+\begin{schemedefn}{\scheme|Ccps[E[(wcm hole body-expr)]]|}
 \begin{schemedisplay}
@@
   ((lambda (rest-marks) 
-      (((Ccps[body-expr] Ccps[E]) TRUE) Cpcps[((CONS mark-value) rest-marks)]))
+      (((Ccps[body-expr] Ccps[E]) TRUE) Chcps[((CONS mark-value) rest-marks)]))
     ((xiE Chcps[(SND chiE)]) Cpcps[chiE])))
@@
 
-\begin{schemedefn}{\scheme|E[(wcm v_0 hole)]|}
+\begin{schemedefn}{\scheme|Ccps[E[(wcm v_0 hole)]]|}
 \begin{schemedisplay}
@@
 
+This allows us to define $\mathcal{C}_{\mathrm{cps}}$ over a context-expression pair with mildly unorthodox notation.
+
 \begin{schemedefn}{\scheme|Ccps[E[e]]|}
 \begin{schemedisplay}
-(((Ccps[e] C[E]) xiE) Cpcps[E])
-\end{schemedisplay}
-\end{schemedefn}
-
-From this definition, it is apparent that \scheme|Chcps[p]|=\scheme|(((Ccps[p] Ccps[hole]) XI(hole)) Cpcps[hole])|.
+(((Ccps[e] Ccps[E]) xiE) Cpcps[chiE])
+\end{schemedisplay}
+\end{schemedefn}
+
+From this definition, it is apparent that \scheme|Chcps[p]|=\scheme|(((Ccps[p] Ccps[hole]) XI(hole)) Cpcps[hole])|=\scheme|Ccps[hole[p]]|.
 
@@
 With top-level control of a program, we can transform the relevant part of the system, leaving the rest the same.
-This preserves interoperability with foreign functions, e.g., third-party JavaScript libraries.
+This preserves interoperability with foreign functions, i.e., third-party JavaScript libraries.
 

### 2012-10-10 paper.tex
Winding it up.

@@
 \usepackage{slatex}
-\usepackage{color}
+\usepackage{listings}
 
@@
 Continuation marks \cite{clements2006portable} are a programming language feature which generalizes stack inpection.
-Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail} and to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}.
-[Mention parameters?]
+Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail}, to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}, and to implement a form of dynamic binding in parameters.
 
@@
 
-We now address the absence of \scheme'let', \scheme'if', \scheme'cons', etc. from \lv.
-
-We can express the \scheme'let' construct in \lv\ with application. An expression such as
-\begin{schemedisplay}
-(let ((x_1 e_1)
-      ...
-      (x_n e_n))
-  e)
-\end{schemedisplay}
-can be interpreted as
-\begin{schemedisplay}
-(...((lambda (x_1) ... (lambda (x_n)
-           e)...) e_1) ...) e_n)
-\end{schemedisplay}
-which is the curried form of
-\begin{schemedisplay}
-((lambda (x_1 ... x_n)
-           e) e_1 ... e_n)
-\end{schemedisplay}
-We unfold this characterization of \scheme'let' to guide the construction of $\mathcal{C}$ before simplifying.
-[Drop everything after ``with application.''?]
-
-To achieve \scheme'if' and conditionals as well as list primitives \scheme'cons', \scheme'snd', and \scheme'nil', we use the Church encodings of fig. \ref{church-encodings}.
+Now, to address the absence of \scheme'let', \scheme'if', \scheme'cons', etc.
+from \lv: We can express the \scheme'let' construct in \lv\ with application.
+To achieve \scheme'if' and conditionals as well as list primitives \scheme'cons', \scheme'snd', and \scheme'nil', we use the Church encodings of fig.
+\ref{church-encodings}.
 
@@
 
+\noindent
+\begin{minipage}{0.5\linewidth}
 \begin{schemedefn}{\scheme|Ccps[v]|=\scheme|Ccps[(lambda (x) e)]|}
@@
 \end{schemedefn}
-
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
 \begin{schemedefn}{\scheme|Ccps[x]|}
@@
 \end{schemedefn}
+\end{minipage}
+\\
+\\
 
 We include corresponding definitions for $\mathcal{C}'$ and $\hat{\mathcal{C}}$.
-
+\\
+\\
+\noindent
+\begin{minipage}{0.5\linewidth}
 \begin{schemedefn}{\scheme|Cpcps[(lambda (x) e)]|}
@@
 \end{schemedefn}
-
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
 \begin{schemedefn}{\scheme|Chcps[p]|}
@@
 \end{schemedefn}
+\end{minipage}
 
@@
 
-Now consider the reduction of the same program transformed. We apply the transformation just-in-time as we reduce to reduce term size and promote clarity.
+Now consider the reduction of the same program transformed.
+We apply the transformation just-in-time as we reduce to prevent term size explosion and promote clarity and omit the uninteresting reductions.
 \begin{schemedisplay}
 Ch[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))]
-
+; By definition this is
 ((C[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))] FALSE) Cp[NIL])
@@
 \end{schemedisplay}
-
-[This example is way too long. Are trivial examples generally uninteresting?]
 
@@
 
-We have discovered a truly marvelous proof of this, which this paper is too short to contain.
-
-We offer a sketch of the proof here.
+Because of space limitations, we offer only a sketch of the proof here. [The full proof can be found in a technical report at ...]
 
@@
 
-%\spnewtheorem*{maintheoremeu}{Correctness of $\hat{\mathcal{C}}$}{a}{x}
 \begin{theorem}
@@
 
-We do this by overloading $\mathcal{C}_{\mathrm{cps}}$ to accommodate contexts which allows us to formally relate $E[e]$ and $\Ccps{E[e]}$. We first define
-\begin{plaindef}
-\[
-\xi(E)=\begin{cases}
-\mathbf{true} &\text{if $E=E'[\wcm{v'}{\hole}]$ for some $E'$ and $v'$}\\
-\mathbf{false} &\text{otherwise}
-\end{cases}
-\]
-\end{plaindef}
-to denote the \scheme'flags' argument and assume that the \scheme'marks' argument is \scheme|Cpcps[chiE]|. We can now define $\mathcal{C}_{\mathrm{cps}}$ over contexts:
-
-\begin{schemedefn}{\scheme|Ccps[hole]|}
-\begin{schemedisplay}
-(lambda (value)
-  value)
-\end{schemedisplay}
-\end{schemedefn}
-
-\begin{schemedefn}{\scheme|Ccps[E[(hole rand-value)]]|}
-\begin{schemedisplay}
-(lambda (rator-value)
-  (((Ccps[rand-expr]
-     (lambda (rand-value)
-       ((((rator-value rand-value) Ccps[E]) xiE) Cpcps[chiE])))
-    FALSE)
-  Cpcps[chiE]))
-\end{schemedisplay}
-\end{schemedefn}
-
-\begin{schemedefn}{\scheme|Ccps[E[(v_0 hole)]]|}
-\begin{schemedisplay}
-(lambda (rand-value)
-  ((((v_0 rand-value) Ccps[E]) xiE) Cpcps[chiE]))
-\end{schemedisplay}
-\end{schemedefn}
-
-\begin{schemedefn}{\scheme|Ccps[E[(wcm hole body-expr)]]|}
-\begin{schemedisplay}
-(lambda (mark-value) 
-  ((lambda (rest-marks) 
-      (((Ccps[body-expr] Ccps[E]) TRUE) Chcps[((CONS mark-value) rest-marks)]))
-    ((xiE Chcps[(SND chiE)]) Cpcps[chiE])))
-\end{schemedisplay}
-\end{schemedefn}
-
-\begin{schemedefn}{\scheme|Ccps[E[(wcm v_0 hole)]]|}
-\begin{schemedisplay}
-Ccps[E]
-\end{schemedisplay}
-\end{schemedefn}
-
-This allows us to define $\mathcal{C}_{\mathrm{cps}}$ over a context-expression pair with mildly unorthodox notation.
-
+We prove this by overloading $\mathcal{C}_{\mathrm{cps}}$ to accommodate first evaluation contexts and then context-term pairs (denoted in unorthodox style) by
 \begin{schemedefn}{\scheme|Ccps[E[e]]|}
@@
 \end{schemedefn}
-
-From this definition, it is apparent that \scheme|Chcps[p]|=\scheme|(((Ccps[p] Ccps[hole]) XI(hole)) Cpcps[hole])|=\scheme|Ccps[hole[p]]|.
-
-This in turn allows us to express the following lemmas.
-
-\begin{lemma}
-For all contexts $E$ and terms $e$, $E[e]\cmrrs v\implies\C{E[e]}\lvrrs\Cp{v}$.
+\noindent
+which allows us to formally relate $E[e]$ and $\Ccps{E[e]}$. We then prove a lemma
+\begin{lemma}[Simulation]
+For all contexts $E\in\lambda_{cm}$ and expressions $e\in\lambda_{cm}$, $E[e]\cmrr E'[e']\implies\C{E[e]}\lvrrs\C{E'[e']}$
 \end{lemma}
-
-To accomplish this, we first prove by structural induction that for all terms $e$ and context $E$ arbitrary but fixed, $E[e]\cmrrs E[v]\implies\C{E[e]}\lvrrs\C{E[v]}$.
-
-We then prove that for all contexts $E$, $E[v]\cmrrs v'\implies\C{E[v]}\lvrrs\Cp{v'}$.
-
-The theorem follows.
+by induction over contexts $E$ and terms $e$. This lemma justifies the evolution of the evaluation context and each rule in the reduction relation except rule \ref{reduction}. We express this with
+\begin{lemma}[Reduction]
+For all values $v_0,v_1\in\lambda_{cm}$, $\app{v_0}{v_1}\cmrr e'\implies\app{\Cpcps{v_0}}{\Cpcps{v_1}}\lvrr\Ccps{e'}$.
+\end{lemma}
+
+With these lemmas, the theorem follows quickly.
 
@@
 Expressing continuation marks in a monadic way is simple enough. In Haskell,
-\begin{verbatim}
+\begin{lstlisting}[language=Haskell]
 data CM m a = CM ((Bool,[m]) -> a)
@@
   return x = CM (\_ -> x)
-  (CM m) >>= f = CM (\(flag,vs) -> let (CM m') = f (m (False,vs)) in m' (flag,vs))
+  (CM m) >>= f = CM (\(flag,vs) ->
+      let (CM m') = f (m (False,vs)) in m' (flag,vs))
 
@@
 runCM (CM m) = m (False,[])
-\end{verbatim}
+\end{lstlisting}
 is one way, using our approach.
 The factorial functions leveraging this implementation exhibit the correct continuation mark behavior but this is \emph{not} a monad.
-Specifically, the right identity law
-\begin{verbatim}
-m >>= return = m
-\end{verbatim}
-does not hold since \texttt{CM} conceptually adds a stack frame in bind (\texttt{>>=}).
+Specifically, the right identity law, \texttt{m >>= return = m}, does not hold since \texttt{CM} conceptually adds a stack frame in bind (\texttt{>>=}).
 Ager et al.
@@
 
-\section{Future Work}
+\section{Future Work and Conclusion}
 
 Although we used some level of mechanization in the proof of the theorem, a formal proof assistant, such as Coq \cite{barras1997coq}, would increase confidence out of the gate that the transformation was correct.
-
-\section{Conclusion}
 

### 2012-10-11 paper.tex
Substitution lemma.

@@
 When an entire program is transformed, all evaluation is suspended, awaiting arguments representing contextual information.
-At the top level, the context is empty, so we pass the contextual information for the empty context: \scheme'FALSE', indicating evaluation is \emph{not} occurring in \scheme'wcm' tail position and \scheme'NIL', an empty list of marks.
+At the top level, the context is empty, so we pass the contextual information for the empty context: \scheme'FALSE', indicating evaluation is \emph{not} occurring in \scheme'wcm' tail position, and \scheme'NIL', an empty list of marks.
 
@@
 and stating our commutativity property as
+[make this equivalence since the left side needs to be reduced?]
 \begin{equation}
@@
 by induction over contexts $E$ and terms $e$. This lemma justifies the evolution of the evaluation context and each rule in the reduction relation except rule \ref{reduction}. We express this with
-\begin{lemma}[Reduction]
-For all values $v_0,v_1\in\lambda_{cm}$, $\app{v_0}{v_1}\cmrr e'\implies\app{\Cpcps{v_0}}{\Cpcps{v_1}}\lvrr\Ccps{e'}$.
+\begin{lemma}[Substitution]
+\label{lem:substitution}
+For all $e,x,v\in\lambda_{cm}$, $\C{e[x\leftarrow v]}=\C{e}[x\leftarrow \Cp{v}]$.
 \end{lemma}

### 2012-10-12 paper.tex
Syntax highlighting and small corrections.

@@
 \usepackage{amsmath}
-%\usepackage{amsthm}
 \usepackage{amssymb}
@@
 \usepackage{listings}
+\usepackage[usenames,dvipsnames]{xcolor}
+
+\lstset{
+	basicstyle=\small\ttfamily,
+	keywordstyle=\color{RoyalBlue},
+	identifierstyle=\color{gray}
+}
 
@@
 
-To avoid this roadblock, we instead take a macro-style approach; that is, we enhance the core language with facilities to manipulate continuation marks and desugar the enhanced language back to the core.
+To avoid this roadblock, we instead take a macro-style approach; that is, we enhance a core language with facilities to manipulate continuation marks and desugar the enhanced language back to the core.
 To make our desugaring transformation portable to other languages, we define it over the \lc, the common core of most higher-order languages.
@@
 
-With a reasonable intuition and definitions hardened by random testing, we prove correctness of the continuation-passing variant of the transform $\mathcal{C}_{\mathrm{cps}}$.
+With a reasonable intuition and definitions hardened by random testing, we prove correctness of the continuation-passing variant of the transform, $\mathcal{C}_{\mathrm{cps}}$.
 
@@
 A properly-recursive factorial can be implemented in this JavaScript extension as
-\begin{verbatim}
+\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
 var fac = function( n ) {
@@
 }
-\end{verbatim}
+\end{lstlisting}
 
 A direct-style transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
-\begin{verbatim}
+\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
 var fac = function( flag, marks ) {
@@
 }
-\end{verbatim}
+\end{lstlisting}
 
 As expected, the output of a manual initiation of this function is
-\begin{verbatim}
+\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
 > fac( false, [] )( 5 )
@@
 120
-\end{verbatim}
+\end{lstlisting}
 
 A tail-recursive factorial is expressible in this extension as
-\begin{verbatim}
+\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
 var fac = function( n, acc ) {
@@
 }
-\end{verbatim}
+\end{lstlisting}
 and transforms to
-\begin{verbatim}
+\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
 var fac = function( flag, marks ) {
@@
 }
-\end{verbatim}
+\end{lstlisting}
 
 The output of a manual initiation of this function is
-\begin{verbatim}
+\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
 > fac( false, [] )( 5, 1 )
@@
 120
-\end{verbatim}
+\end{lstlisting}
 

### 2012-10-12 paper.tex
Changed to equivalence.

@@
 \end{equation}
-
-and stating our commutativity property as
-[make this equivalence since the left side needs to be reduced?]
+This changes our commutativity property somewhat, since applying $\hat{\mathcal{C}}$ to a value--say, the result of $\mathrm{eval}_{cm}$--will instigate further, if benign, reduction.
+In light of this, we must alter our statement about commutativity to reflect that we no longer guarantee term equality, but term equivalence, in the sense that they have the same normal form, or share a lack thereof.
+We can state this modified property as
 \begin{equation}
 \label{commutativity-property}
-\Ch{\evalcm{p}}=\evalv{\Ch{p}}
-\end{equation}
-which is equivalent to
-\begin{equation}
-\app{\app{\C{\evalcm{p}}}{\false}}{\nil}=\evalv{\app{\app{\C{p}}{\false}}{\nil}}
+\Ch{\evalcm{p}}\equiv\evalv{\Ch{p}}
+\end{equation}
+which, expanded, is
+\begin{equation}
+\app{\app{\C{\evalcm{p}}}{\false}}{\nil}\equiv\evalv{\app{\app{\C{p}}{\false}}{\nil}}
 \end{equation}
@@
 ((C[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))] FALSE) Cp[NIL])
-
+; which explodes upon expansion to
 (((lambda (flag)

### 2012-10-12 paper.tex
Penultimate draft.

@@
 This changes our commutativity property somewhat, since applying $\hat{\mathcal{C}}$ to a value--say, the result of $\mathrm{eval}_{cm}$--will instigate further, if benign, reduction.
-In light of this, we must alter our statement about commutativity to reflect that we no longer guarantee term equality, but term equivalence, in the sense that they have the same normal form, or share a lack thereof.
+In light of this, we must alter our statement about commutativity to reflect that we no longer guarantee term equality, but term equivalence, in the sense that they share a normal form (or lack of one).
 We can state this modified property as
@@
 Ch[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))]
-; By definition this is
+\end{schemedisplay}
+
+By definition this is
+\begin{schemedisplay}
 ((C[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))] FALSE) Cp[NIL])
-; which explodes upon expansion to
+\end{schemedisplay}
+which explodes upon expansion to
+\begin{schemedisplay}
 (((lambda (flag)
@@
   FALSE) Cp[NIL])
-
-((lambda (marks)
-   ((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
-    (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-      ((C[0] FALSE) marks)) ((FALSE Ch[(SND marks)]) marks))))
- Cp[NIL])
-
+\end{schemedisplay}
+
+After the application of contextual information, we reach
+\begin{schemedisplay}
 ((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
@@
    ((C[0] FALSE) Cp[NIL])) ((FALSE Ch[(SND NIL)]) Cp[NIL])))
-
-((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
- (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-   (((lambda (flag) (lambda (marks) 0)) FALSE) Cp[NIL])) ((FALSE Ch[(SND NIL)]) Cp[NIL])))
-
-((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
- (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-   ((lambda (marks) 0) Cp[NIL])) ((FALSE Ch[(SND NIL)]) Cp[NIL])))
-
-((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
- (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-   0) ((FALSE Ch[(SND NIL)]) Cp[NIL])))
-
-((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
- ((lambda (rest-marks) Cp[((CONS 0) rest-marks)])
-  ((FALSE Ch[(SND NIL)]) Cp[NIL])))
-
-((C[((lambda (x) (wcm x (ccm))) 1)] TRUE)
- ((lambda (rest-marks) Cp[((CONS 0) rest-marks)])
-  Cp[NIL]))
-
+\end{schemedisplay}
+the transformation of a \scheme'wcm' body.
+No evaluation of this term occurs until the contextual information has been evaluated and applied.
+In the process of evaluation, the \scheme'mark-expr' is evaluated and prepended to the list of continutation marks.
+
+This ultimately results in the following:
+\begin{schemedisplay}
 ((C[((lambda (x) (wcm x (ccm))) 1)] TRUE) Cp[((CONS 0) NIL)])
-
+\end{schemedisplay}
+
+It is evident that this term will behave exactly as a top-level term except as this contextual information influences it, and this is exactly the property we have strived for.
+Expansion of this term yields
+\begin{schemedisplay}
 (((lambda (flag)
@@
        marks))) TRUE) Cp[((CONS 0) NIL)])
-
-(((((C[(lambda (x) (wcm x (ccm)))] FALSE) Cp[((CONS 0) NIL)])
-   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
-  TRUE) Cp[((CONS 0) NIL)])
-
-((((((lambda (flag) (lambda (marks) (lambda (x) C[(wcm x (ccm))]))) FALSE) Cp[((CONS 0) NIL)])
-   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
-  TRUE) Cp[((CONS 0) NIL)])
-
-(((((lambda (marks) (lambda (x) C[(wcm x (ccm))])) Cp[((CONS 0) NIL)])
-   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
-  TRUE) Cp[((CONS 0) NIL)])
-
-((((lambda (x) C[(wcm x (ccm))])
-   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
-  TRUE) Cp[((CONS 0) NIL)])
-
-((((lambda (x) C[(wcm x (ccm))])
-   (((lambda (flag) (lambda (marks) 1)) FALSE) Cp[((CONS 0) NIL)]))
-  TRUE) Cp[((CONS 0) NIL)])
-
-((((lambda (x) C[(wcm x (ccm))])
-   ((lambda (marks) 1) Cp[((CONS 0) NIL)]))
-  TRUE) Cp[((CONS 0) NIL)])
-
+\end{schemedisplay}
+the expansion of an application. Both these terms are values, so are unaffected by the contextual information.
+
+The elided reduction has the effect of preparing the terms for application.
+\begin{schemedisplay}
 ((((lambda (x) C[(wcm x (ccm))])
    1) TRUE) Cp[((CONS 0) NIL)])
-
+\end{schemedisplay}
+reduces to
+\begin{schemedisplay}
 ((C[(wcm 1 (ccm))]
    TRUE) Cp[((CONS 0) NIL)])
-
+\end{schemedisplay}
+
+This expands and reduces as the \scheme'wcm' term seen previously:
+\begin{schemedisplay}
 (((lambda (flag)
@@
   TRUE) Cp[((CONS 0) NIL)])
-
-((lambda (marks)
-   ((C[(ccm)] TRUE)
-    (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-      ((C[1] FALSE) marks))
-     ((TRUE Ch[(SND marks)]) marks))))
- Cp[((CONS 0) NIL)])
-
-((C[(ccm)] TRUE)
- (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
-  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
-
-(((lambda (flag) (lambda (marks) marks))
-  TRUE)
- (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
-  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
-
-((lambda (marks) marks)
- (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-   ((C[1] FALSE) Cp[((CONS 0) NIL)]))
-  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
-
-((lambda (marks) marks)
- (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-   (((lambda (flag) (lambda (marks) 1)) FALSE) Cp[((CONS 0) NIL)]))
-  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
-
-((lambda (marks) marks)
- (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-   ((lambda (marks) 1) Cp[((CONS 0) NIL)]))
-  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
-
-((lambda (marks) marks)
- (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)]))
-   1)
-  ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
-
+\end{schemedisplay}
+
+Of interest in this process is the effective collapse of the previous \scheme'mark' context by virtue of the value of \scheme'flag'. When we reach
+\begin{schemedisplay}
 ((lambda (marks) marks)
@@
   ((TRUE Ch[(SND ((CONS 0) NIL))]) Cp[((CONS 0) NIL)])))
-
+\end{schemedisplay}
+the list is beheaded to simulate mark overwriting:
+\begin{schemedisplay}
 ((lambda (marks) marks)
@@
   Ch[(SND ((CONS 0) NIL))]))
-
-((lambda (marks) marks)
- ((lambda (rest-marks) Cp[((CONS 1) rest-marks)])
-  Cp[NIL]))
-
+\end{schemedisplay}
+
+With the contextual information given, the evaluation of \scheme'ccm' is simple:
+\begin{schemedisplay}
 ((lambda (marks) marks)
  Cp[((CONS 1) NIL)])
-
+\end{schemedisplay}
+reduces to
+\begin{schemedisplay}
 Cp[((CONS 1) NIL)]
 \end{schemedisplay}
+and we are left with just what we hoped for.
 

### 2012-10-12 paper.tex
What changes?

--- 2012-10-12 paper.tex
@@
 \usepackage{amsmath}
+%\usepackage{amsthm}
 \usepackage{amssymb}
 \usepackage{slatex}
-\usepackage{listings}
-\usepackage[usenames,dvipsnames]{xcolor}
-
-\lstset{
-	basicstyle=\small\ttfamily,
-	keywordstyle=\color{RoyalBlue},
-	identifierstyle=\color{gray}
-}
+\usepackage{color}
 
@@
 We document how Redex, a domain-specific language for exploring language semantics, aided the discovery of correct transformations.
-We verify the transformations with the proof of a meaning-preservation theorem.
-Finally, we apply the transformation to JavaScript.
+Finally, we verify the transformations with the proof of a meaning-preservation theorem.
 \end{abstract}
@@
 Continuation marks \cite{clements2006portable} are a programming language feature which generalizes stack inpection.
-Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail}, to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}, and to implement a form of dynamic binding in parameters.
+Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail} and to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}.
+[Mention parameters?]
 
@@
 
-To avoid this roadblock, we instead take a macro-style approach; that is, we enhance a core language with facilities to manipulate continuation marks and desugar the enhanced language back to the core.
+To avoid this roadblock, we instead take a macro-style approach; that is, we enhance the core language with facilities to manipulate continuation marks and desugar the enhanced language back to the core.
 To make our desugaring transformation portable to other languages, we define it over the \lc, the common core of most higher-order languages.
@@
 Behaviorally, the outer one is collapsed \emph{before} evaluation of the inner body proceeds.
-In essence, it is overwritten.
+In effect, it is overwritten.
 Following this, evaluation continues as
@@
 \setspecialsymbol{CE}{$\C{E}$}
-\setspecialsymbol{XI}{$\xi$}
 \setspecialsymbol{xiE}{$\xi(E)$}
@@
 
-Now, to address the absence of \scheme'let', \scheme'if', \scheme'cons', etc.
-from \lv: We can express the \scheme'let' construct in \lv\ with application.
-To achieve \scheme'if' and conditionals as well as list primitives \scheme'cons', \scheme'snd', and \scheme'nil', we use the Church encodings of fig.
-\ref{church-encodings}.
+We now address the absence of \scheme'let', \scheme'if', \scheme'cons', etc. from \lv.
+
+We can express the \scheme'let' construct in \lv\ with application. An expression such as
+\begin{schemedisplay}
+(let ((x_1 e_1)
+      ...
+      (x_n e_n))
+  e)
+\end{schemedisplay}
+can be interpreted as
+\begin{schemedisplay}
+(...((lambda (x_1) ... (lambda (x_n)
+           e)...) e_1) ...) e_n)
+\end{schemedisplay}
+which is the curried form of
+\begin{schemedisplay}
+((lambda (x_1 ... x_n)
+           e) e_1 ... e_n)
+\end{schemedisplay}
+We unfold this characterization of \scheme'let' to guide the construction of $\mathcal{C}$ before simplifying.
+[Drop everything after ``with application.''?]
+
+To achieve \scheme'if' and conditionals as well as list primitives \scheme'cons', \scheme'snd', and \scheme'nil', we use the Church encodings of fig. \ref{church-encodings}.
 
@@
 
-In our call-by-value language, abstracting terms has the effect of suspending evaluation.
-When an entire program is transformed, all evaluation is suspended, awaiting arguments representing contextual information.
-At the top level, the context is empty, so we pass the contextual information for the empty context: \scheme'FALSE', indicating evaluation is \emph{not} occurring in \scheme'wcm' tail position, and \scheme'NIL', an empty list of marks.
+Abstracting terms has the effect of suspending evaluation.
+When an entire program is transformed, all evaluation is suspended, and awaits arguments representing contextual information.
+At the top level, the context is empty, so we pass the contextual information for the empty context: \scheme'FALSE', indicating evaluation is \emph{not} occurring in \scheme'wcm' tail position and \scheme'NIL', an empty list of marks.
 
@@
 and stating our commutativity property as
-[make this equivalence since the left side needs to be reduced?]
 \begin{equation}
@@
 
-The formal transformation of a \scheme'wcm' directive is also extremely similar to the \scheme'let' version.
-The Church-encoded conditional eagerly evaluates both branches, but this still achieves correct behavior.
+The formal transformation of a \scheme'wcm' directive is also extremely similar to the \scheme'let' version. The definition of \scheme'C[cons]' is unfolded and simplified.
 \begin{schemedisplay}
@@
     ((C[body-expr] TRUE)
-     (((lambda (mark-value) (lambda (rest-marks) Ch[((CONS mark-value) rest-marks)]))
+     (((lambda (mark-value) (lambda (rest-marks) Cp[((CONS mark-value) rest-marks)])))
        ((C[mark-expr] FALSE) marks))
@@
             ((lambda (rest-marks) 
-               (((Ccps[body-expr] kont) TRUE) Chcps[((CONS mark-value) rest-marks)]))
-             ((flag Chcps[(SND marks)]) marks))))
+               (((Ccps[body-expr] kont) TRUE) Cp[((CONS mark-value) rest-marks)])))
+             ((flag Ch_cps[(SND marks)]) marks))))
         FALSE)
@@
 
-\noindent
-\begin{minipage}{0.5\linewidth}
 \begin{schemedefn}{\scheme|Ccps[v]|=\scheme|Ccps[(lambda (x) e)]|}
@@
 \end{schemedefn}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
+
 \begin{schemedefn}{\scheme|Ccps[x]|}
@@
 \end{schemedefn}
-\end{minipage}
-\\
-\\
 
 We include corresponding definitions for $\mathcal{C}'$ and $\hat{\mathcal{C}}$.
-\\
-\\
-\noindent
-\begin{minipage}{0.5\linewidth}
+
 \begin{schemedefn}{\scheme|Cpcps[(lambda (x) e)]|}
@@
 \end{schemedefn}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
+
 \begin{schemedefn}{\scheme|Chcps[p]|}
@@
 \end{schemedefn}
-\end{minipage}
 
@@
 
-To better illustrate what the transformation does, we step through the reduction of a program which exhibits its more interesting aspects.
-One \cm\ program suited to this purpose is \scheme|(wcm 0 ((lambda (x) (wcm x (ccm))) 1))|.
-It reduces according to \cm\ semantics as
+Consider the \cm\ program \scheme|(wcm 0 ((lambda (x) (wcm x (ccm))) 1))| which reduces as
 \begin{schemedisplay}
@@
 \end{schemedisplay}
-
-Now consider the reduction of the same program transformed.
-We apply the transformation just-in-time as we reduce to prevent term size explosion and promote clarity and omit the uninteresting reductions.
+and transforms as
 \begin{schemedisplay}
 Ch[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))]
-; By definition this is
+
 ((C[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))] FALSE) Cp[NIL])
@@
 \end{schemedisplay}
+
+[This example is way too long. Are trivial examples generally uninteresting?]
+
+% TODO
+%fix conclusion
 
@@
 
-With a reasonable intuition and definitions hardened by random testing, we prove correctness of the continuation-passing variant of the transform, $\mathcal{C}_{\mathrm{cps}}$.
-
-Because of space limitations, we offer only a sketch of the proof here. [The full proof can be found in a technical report at ...]
+With a reasonable intuition and definitions hardened by random testing, we prove correctness.
+
+We have discovered a truly marvelous proof of this, which this paper is too short to contain.
+
+We offer a sketch of the proof here.
 
@@
 
+%\spnewtheorem*{maintheoremeu}{Correctness of $\hat{\mathcal{C}}$}{a}{x}
 \begin{theorem}
-For all programs $p\in\lambda_{cm}$, $\Chcps{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Chcps{p})$.
+For all programs $p\in\lambda_{cm}$, $\Ch{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Ch{p})$.
 \end{theorem}
 
-We prove this by overloading $\mathcal{C}_{\mathrm{cps}}$ to accommodate first evaluation contexts and then context-term pairs (denoted in unorthodox style) by
-\begin{schemedefn}{\scheme|Ccps[E[e]]|}
-\begin{schemedisplay}
-(((Ccps[e] Ccps[E]) xiE) Cpcps[chiE])
-\end{schemedisplay}
-\end{schemedefn}
-\noindent
-which allows us to formally relate $E[e]$ and $\Ccps{E[e]}$. We then prove a lemma
-\begin{lemma}[Simulation]
-For all contexts $E\in\lambda_{cm}$ and expressions $e\in\lambda_{cm}$, $E[e]\cmrr E'[e']\implies\C{E[e]}\lvrrs\C{E'[e']}$
+We do this by overloading $\mathcal{C}$ to accommodate contexts which allows us to formally relate $E[e]$ and $\C{E[e]}$.
+
+This in turn allows us to express the following lemmas.
+
+\begin{lemma}
+For all contexts $E$ and terms $e$, $E[e]\cmrrs v\implies\C{E[e]}\lvrrs\Cp{v}$.
 \end{lemma}
-by induction over contexts $E$ and terms $e$. This lemma justifies the evolution of the evaluation context and each rule in the reduction relation except rule \ref{reduction}. We express this with
-\begin{lemma}[Substitution]
-\label{lem:substitution}
-For all $e,x,v\in\lambda_{cm}$, $\C{e[x\leftarrow v]}=\C{e}[x\leftarrow \Cp{v}]$.
-\end{lemma}
-
-With these lemmas, the theorem follows quickly.
+
+To accomplish this, we first prove by structural induction that for all terms $e$ and context $E$ arbitrary but fixed, $E[e]\cmrrs E[v]\implies\C{E[e]}\lvrrs\C{E[v]}$.
+
+We then prove that for all contexts $E$, $E[v]\cmrrs v'\implies\C{E[v]}\lvrrs\Cp{v'}$.
+
+The theorem follows.
 
@@
 Tail-call elimination is not part of the JavaScript specification which complicates the treatment of the tail-call behavior.
-We simulate proper tail-call behavior by using the flag to encode whether a function call is in tail position.
+We have chosen to perform a simple static analysis which determines if the recursive call is in tail position.
+We simulate proper tail-call behavior by statically assigning the value of the flag to reflect this.
 
 A properly-recursive factorial can be implemented in this JavaScript extension as
-\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
+\begin{verbatim}
 var fac = function( n ) {
@@
 }
-\end{lstlisting}
-
-A direct-style transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
-\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
+\end{verbatim}
+
+A transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
+\begin{verbatim}
 var fac = function( flag, marks ) {
@@
 }
-\end{lstlisting}
+\end{verbatim}
 
 As expected, the output of a manual initiation of this function is
-\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
+\begin{verbatim}
 > fac( false, [] )( 5 )
@@
 120
-\end{lstlisting}
+\end{verbatim}
 
 A tail-recursive factorial is expressible in this extension as
-\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
+\begin{verbatim}
 var fac = function( n, acc ) {
@@
 }
-\end{lstlisting}
+\end{verbatim}
 and transforms to
-\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
+\begin{verbatim}
 var fac = function( flag, marks ) {
@@
 }
-\end{lstlisting}
+\end{verbatim}
 
 The output of a manual initiation of this function is
-\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
+\begin{verbatim}
 > fac( false, [] )( 5, 1 )
@@
 120
-\end{lstlisting}
-
-In this example, the API to compute factorials has changed, but the change is localized.
-With top-level control of a program, we can transform the relevant part of the system, leaving the rest the same.
-This preserves interoperability with foreign functions, i.e., third-party JavaScript libraries.
+\end{verbatim}
 
@@
 Their appeal is significant: they allow one to implement a particular model of computation with with a pure computational logic, such as the \lc.
+Ager et al.
+\cite{ager2003functional} implemented a stack-inspection monad in terms of a lifted state monad, but the size of the continuation must be managed by hand within.
 
 Expressing continuation marks in a monadic way is simple enough. In Haskell,
-\begin{lstlisting}[language=Haskell]
+\begin{verbatim}
+module CM where
+import Control.Monad
+
 data CM m a = CM ((Bool,[m]) -> a)
@@
   return x = CM (\_ -> x)
-  (CM m) >>= f = CM (\(flag,vs) ->
-      let (CM m') = f (m (False,vs)) in m' (flag,vs))
+  (CM m) >>= f = CM (\(flag,vs) -> let (CM m') = f (m (False,vs)) in m' (flag,vs))
 
@@
 runCM (CM m) = m (False,[])
-\end{lstlisting}
-is one way, using our approach.
-The factorial functions leveraging this implementation exhibit the correct continuation mark behavior but this is \emph{not} a monad.
-Specifically, the right identity law, \texttt{m >>= return = m}, does not hold since \texttt{CM} conceptually adds a stack frame in bind (\texttt{>>=}).
-Ager et al.
-\cite{ager2003functional} implemented a similar stack-inspection monad in terms of a lifted state monad, but resorted to manual management of the stack frames.
-
-\section{Future Work and Conclusion}
+\end{verbatim}
+
+\begin{verbatim}
+module Main where
+import CM
+
+fac :: Int -> CM Int (Int,[Int])
+fac 0 = do
+  vs <- ccm;
+  return (1,vs)
+
+fac n = do
+  (acc,vs) <- wcm n (fac (n - 1));
+  return (n * acc,vs)
+
+fac_tr :: Int -> Int -> CM Int (Int,[Int])
+fac_tr 0 acc = do
+  vs <- ccm;
+  return (acc,vs)
+
+fac_tr n acc = wcm n (fac_tr (n - 1) (n * acc))
+
+main = do
+  print $ runCM (fac 5);
+  print $ runCM (fac_tr 5 1)
+\end{verbatim}
+
+\begin{verbatim}
+(120,[1,2,3,4,5])
+(120,[1])
+\end{verbatim}
+[Say more with less.]
+
+\section{Future Work}
 
 Although we used some level of mechanization in the proof of the theorem, a formal proof assistant, such as Coq \cite{barras1997coq}, would increase confidence out of the gate that the transformation was correct.
+
+\section{Conclusion}
 

### 2012-10-15 paper.tex

--- 2012-10-12 paper.tex
@@
 
-Continuation marks \cite{clements2006portable} are a programming language feature which generalizes stack inpection.
-Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail}, to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}, and to implement a form of dynamic binding in parameters.
+Continuation marks \cite{clements2006portable} are a programming language feature which generalizes stack inspection.
+Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail}, to express aspect-oriented programming in higher-order languages \cite{tucker2003pointcuts}, and to implement a form of dynamic binding in parameters.
 
@@
 
-Now consider the reduction of the same program transformed.
-We apply the transformation just-in-time as we reduce to prevent term size explosion and promote clarity and omit the uninteresting reductions.
+Now consider the reduction of the same program transformed. We apply the transformation just-in-time as we reduce to prevent term size explosion and promote clarity and omit uninteresting reductions.
 \begin{schemedisplay}
@@
 \end{schemedisplay}
-the transformation of a \scheme'wcm' body.
-No evaluation of this term occurs until the contextual information has been evaluated and applied.
-In the process of evaluation, the \scheme'mark-expr' is evaluated and prepended to the list of continutation marks.
-
-This ultimately results in the following:
+the transformation of the \scheme'wcm' body.
+Terms within are arranged so that correct evaluation occurs within the native call-by-value regime.
+This evaluates \scheme'mark-expr' and and prepends its value to the list of continuation marks before proceeding with evaluation of \scheme'body-expr'.
+This reduction soon yields the following term:
 \begin{schemedisplay}
@@
 \end{schemedisplay}
-
 It is evident that this term will behave exactly as a top-level term except as this contextual information influences it, and this is exactly the property we have strived for.
@@
 \end{schemedisplay}
-the expansion of an application. Both these terms are values, so are unaffected by the contextual information.
-
-The elided reduction has the effect of preparing the terms for application.
+the expansion of an application.
+In this example, both the operator and operand are values, so are essentially unaffected by the application of contextual information; this application has the effect of preparing the terms for application:
 \begin{schemedisplay}
@@
 
-With the contextual information given, the evaluation of \scheme'ccm' is simple:
+Once given the contextual information, the evaluation of \scheme'ccm' is simple:
 \begin{schemedisplay}
@@
 
-Because of space limitations, we offer only a sketch of the proof here. [The full proof can be found in a technical report at ...]
+Because of space limitations, we offer only a sketch of the proof here.
+The full proof can be found in a technical report\footnote{\texttt{http://students.cs.byu.edu/\textasciitilde{}krgerman/esop2013-technical-report.pdf}}.
 
@@
 Expressing continuation marks in a monadic way is simple enough. In Haskell,
+\begin{minipage}{\linewidth}
 \begin{lstlisting}[language=Haskell]
@@
 \end{lstlisting}
+\end{minipage}
 is one way, using our approach.
 The factorial functions leveraging this implementation exhibit the correct continuation mark behavior but this is \emph{not} a monad.
-Specifically, the right identity law, \texttt{m >>= return = m}, does not hold since \texttt{CM} conceptually adds a stack frame in bind (\texttt{>>=}).
+Specifically, the right identity law, \texttt{{\color{gray}m} >>= {\color{RoyalBlue}return} = {\color{gray}m}}, does not hold since \texttt{{\color{gray}CM}} conceptually adds a stack frame in bind (\texttt{>>=}).
 Ager et al.
@@
 
-\section{Future Work and Conclusion}
-
-Although we used some level of mechanization in the proof of the theorem, a formal proof assistant, such as Coq \cite{barras1997coq}, would increase confidence out of the gate that the transformation was correct.
+\section{Conclusion and Future Work}
 
@@
 Despite their demonstrated utility, they have not yet found their way into most languages.
-A verified characterization of continuation marks in a pure computational language provides implementors of higher-order languages a correct compiler for continuation marks.
+A verified characterization of continuation marks in a pure computational language provides implementors of higher-order languages a correct compiler for continuation marks which we have demonstrated for JavaScript.
+
+Our macro-style approach would be more useful if deployed as a proper macro, but this requires a hygienic macro system which many languages lack.
+\texttt{sweet.js} is just such a system for JavaScript in early stages and is ideal for our transformation.
+
+Although we used some level of mechanization in the proof of the theorem, a formal proof assistant, such as Coq \cite{barras1997coq}, would increase confidence out of the gate that the transformation was correct.
 

### 2012-10-15 proof.tex
For posterity.

\chapter{Proof}

% TODO
% make overline show up in mechanized proof and definitions, where necessary
% label mechanized proof better
% justify each semantic rule with a lemma
% label transition steps
% make beta appear in Redex stuff
% make sure the paper with Jay's suggestions is satisfied

The evaluation of a \cm\ program proceeds with the evolution of an evaluation context and possibly reducible expression. When evaluation begins, the evaluation context is merely \scheme'hole', a placeholder for the eventual result, and the reducible expression, or redex, is the program itself. The evaluation of arguments--both in application and continuation mark forms--defers evaluation of the expression at hand by storing its evaluation context and evaluating subterms. As these results are applied and evaluation continues, the size of the context fluctuates until finally, if the program terminates, we are left with a single value to plug in \scheme'hole'. This value is the value of the program.

The state of evaluation at any given point can be encapsulated by a pair of an evaluation context $E$ and an expression $e$ which we write in unorthodox style as $E[e]$. In order to prove that evaluation in the transformation corresponds to native evaluation, we must relate this state with its corresponding transformation.

We do this by overloading $\mathcal{C}_{\mathrm{cps}}$ to accommodate evaluation contexts which allows us to formally relate $E[e]$ and $\Ccps{E[e]}$. We first define
\begin{definition}
\[
\xi(E)=\begin{cases}
\mathbf{true} &\text{if $E=E'[\wcm{v'}{\hole}]$ for some $E'$ and $v'$}\\
\mathbf{false} &\text{otherwise}
\end{cases}
\]
\end{definition}
\noindent
to denote the \scheme'flags' argument and assume that the \scheme'marks' argument is \scheme|Cpcps[chiE]|. We can now define $\mathcal{C}_{\mathrm{cps}}$ over contexts $E\in\lambda_{cm}$:

\begin{schemedefinition}{\scheme|Ccps[hole]|}
\begin{schemeblock}
\begin{schemedisplay}
(lambda (value)
  value)
\end{schemedisplay}
\end{schemeblock}
\end{schemedefinition}

\begin{schemedefinition}{\scheme|Ccps[E[(hole rand-value)]]|}
\begin{schemeblock}
\begin{schemedisplay}
(lambda (rator-value)
  (((Ccps[rand-expr]
     (lambda (rand-value)
       ((((rator-value rand-value) Ccps[E]) xiE) Cpcps[chiE])))
    FALSE)
  Cpcps[chiE]))
\end{schemedisplay}
\end{schemeblock}
\end{schemedefinition}

\begin{schemedefinition}{\scheme|Ccps[E[(v_0 hole)]]|}
\begin{schemeblock}
\begin{schemedisplay}
(lambda (rand-value)
  ((((v_0 rand-value) Ccps[E]) xiE) Cpcps[chiE]))
\end{schemedisplay}
\end{schemeblock}
\end{schemedefinition}

\begin{schemedefinition}{\scheme|Ccps[E[(wcm hole body-expr)]]|}
\begin{schemeblock}
\begin{schemedisplay}
(lambda (mark-value) 
  ((lambda (rest-marks) 
      (((Ccps[body-expr] Ccps[E]) TRUE) Chcps[((CONS mark-value) rest-marks)]))
    ((xiE Chcps[(SND chiE)]) Cpcps[chiE])))
\end{schemedisplay}
\end{schemeblock}
\end{schemedefinition}

\begin{schemedefinition}{\scheme|Ccps[E[(wcm v_0 hole)]]|}
\begin{schemeblock}
\begin{schemedisplay}
Ccps[E]
\end{schemedisplay}
\end{schemeblock}
\end{schemedefinition}

This allows us to define $\mathcal{C}_{\mathrm{cps}}$ over a context-expression pair.

\begin{schemedefinition}{\scheme|Ccps[E[e]]|}
\begin{schemeblock}
\begin{schemedisplay}
(((Ccps[e] Ccps[E]) xiE) Cpcps[chiE])
\end{schemedisplay}
\end{schemeblock}
\end{schemedefinition}

From this definition, it is apparent that \scheme|Chcps[p]|=\scheme|(((Ccps[p] Ccps[hole]) XI(hole)) Cpcps[hole])|=\scheme|Ccps[hole[p]]|.

Now we show that substitution is preserved by the transformation.

\begin{lemma}[Substitution]
\label{lem:substitution}
For all $e,x,v\in\lambda_{cm}$, $\C{e[x\leftarrow v]}=\C{e}[x\leftarrow \Cp{v}]$.
\end{lemma}

See section \ref{app:substitution} for proof.

Finally, we define ``filling the hole'', the insertion of a value in the context from which it came.

\begin{schemedefinition}{\scheme|Ccps[E[v]]|}
\begin{schemeblock}
\begin{schemedisplay}
(Ccps[E] Cpcps[v])
\end{schemedisplay}
\end{schemeblock}
\end{schemedefinition}

With each significant step of native evaluation formally related with the transformation, we can express a simulation lemma.

\begin{lemma}[Simulation]
For all contexts $E\in\lambda_{cm}$ and expressions $e\in\lambda_{cm}$, $E[e]\cmrr E'[e']\implies\C{E[e]}\lvrrs\C{E'[e']}$
\end{lemma}

We will reason by structural induction on both contexts $E$ and terms $e$. Instead of
nesting the induction, which requires the consideration of $|E|\cdot|e|$ cases, we will
take first $E$ and then $e$ in isolation, in each assuming the correctness of the other,
which requires the consideration of only $|E|+|e|$ cases.

First, we prove it holds for terms $e$. In each case, let $E$ be an arbitrary context.
\begin{proof}{Case $e=\app{e_0}{e_1}$}
By steps app1-app3, $\Ccps{E[\app{e_0}{e_1}]}\lvrrs\Ccps{E[\app{\hole}{e_1}][e_0]}$.
\end{proof}
\begin{proof}{Case $e=\wcm{e_0}{e_1}$}
By steps wcm1-wcm3, $\Ccps{E[\wcm{e_0}{e_1}]}\lvrrs\Ccps{E[\wcm{\hole}{e_1}][e_0]}$.
\end{proof}
\begin{proof}{Case $e=\ccm$}
By steps ccm1-ccm3, $\Ccps{E[\ccm]}\lvrrs\Ccps{E[\chi(E)]}$.
\end{proof}
\begin{proof}{Case $e=v_0$}
By steps value1-value3, $\Ccps{E[v_0]}\lvrrs\Ccps{E[v_0]}$.
\end{proof}
\begin{proof}{Case $e=x$}
By steps x1-x4, $\Ccps{E[x]}\lvrrs\Ccps{E[\mathrm{error}]}$.
\end{proof}

Now, we prove it holds for contexts $E$. In each case, let $v_0$ be an arbitrary value.
\begin{proof}{Case $E=\hole$}
This is identical to the case that $e=v_0$.
\end{proof}

\begin{proof}{Case $E=E'[\app{\hole}{e_1}]$}
By step app4, $\Ccps{E[v_0]}\lvrr\Ccps{E'[\app{v_0}{\hole}][e_1]}$.
\end{proof}

\begin{proof}{Case $E=E'[\app{v_0}{\hole}]$}
By step app5, $\Ccps{E[v_0]}\lvrr\Ccps{E'[\app{v_0}{v_1}]}$.

By step app6 and lemma \ref{lem:substitution}, $\Ccps{E'[\app{v_0}{v_1}]}\lvrr\Ccps{E'[e']}$.
\end{proof}

\begin{proof}{Case $E=E'[\wcm{\hole}{e_1}]$}
If $E'=E''[\wcm{v'}{\hole}$ for some $E''$ and $v'$, then $\Ccps{E'[\wcm{\hole}{e_1}][v_0]}\lvrrs\Ccps{E''[\wcm{v_0}{\hole}][e_1]}$ by steps wcm4tail-wcm6tail.

Otherwise, $\Ccps{E'[\wcm{\hole}{e_1}][v_0]}\lvrrs\Ccps{E'[\wcm{v_0}{\hole}][e_1]}$ by steps wcm4nontail-wcm6nontail.
\end{proof}

\begin{proof}{Case $E=E'[\wcm{v_0}{\hole}]$}
By definition, $\Ccps{E[v_1]}=\Ccps{E'[v_1]}$.
\end{proof}

\begin{corollary}
\label{cor:great}
For all contexts $E$ and terms $e\in\lambda_{cm}$, if $E[e]\cmrrs v$, then $\C{E[e]}\lvrrs\Cp{v}$.
\end{corollary}

\begin{proof}
%Let $E\in\lambda_{cm}$ and $e\in\lambda_{cm}$ be arbitrary.
By simulation.
\end{proof}

%\newtheorem{maintheorem}{Correctness of $\hat{\mathcal{C}}$}
\begin{theorem}
For all programs $p\in\lambda_{cm}$, $\Chcps{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Chcps{p})$.
\end{theorem}

\begin{proof}

\begin{align*}
p\cmrrs v &\implies\mathrm{eval}_{cm}(p)=v &\text{by definition of $\mathrm{eval}_{cm}$}\\
          &\implies\Ch{\mathrm{eval}_{cm}(p)}\lvrrs\Cp{v} &\text{by definition of $\hat{\mathcal{C}}$}\\
          &\implies\Ch{\mathrm{eval}_{cm}(p)}\equiv\Cp{v} &\text{by definition of equivalency}\\
\end{align*}

\begin{align*}
p\cmrrs v &\implies\hole[p]\cmrrs v &\text{(since $p\cmrr\hole[p]$)}\\
          &\implies\C{\hole[p]}\lvrrs\Cp{v} &\text{(by corollary \ref{cor:great})}\\
          &\implies\app{\app{\app{\C{p}}{\C{\hole}}}{\xi(\hole)}}{\Cp{\chi(\hole)}}\lvrrs\C{v}&\text{by definition of $\mathcal{C}$}\\
          &\implies\Ch{p}\lvrrs\C{v}&\text{(by definition of $\hat{\mathcal{C}}$)}\\
          &\implies\mathrm{eval}_{v}(\Ch{p})=\Cp{v} &\text{(by definition of $\mathrm{eval}_v$)}
\end{align*}

Therefore, $\C{\mathrm{eval}_{cm}(p)}\equiv\mathrm{eval}_{v}(\C{p})$.
\end{proof}

### 2012-10-15 proof.tex
Easier to follow.

@@
 \begin{lemma}[Simulation]
-For all contexts $E\in\lambda_{cm}$ and expressions $e\in\lambda_{cm}$, $E[e]\cmrr E'[e']\implies\C{E[e]}\lvrrs\C{E'[e']}$
+For all contexts $E\in\lambda_{cm}$ and expressions $e\in\lambda_{cm}$, $E[e]\cmrr E'[e']\implies\Ccps{E[e]}\lvrrs\Ccps{E'[e']}$
 \end{lemma}
@@
 
-\begin{corollary}
-\label{cor:great}
-For all contexts $E$ and terms $e\in\lambda_{cm}$, if $E[e]\cmrrs v$, then $\C{E[e]}\lvrrs\Cp{v}$.
-\end{corollary}
+\begin{lemma}
+\label{lem:great}
+For all contexts $E$ and terms $e\in\lambda_{cm}$, if $E[e]\cmrrs v$, then $\Ccps{E[e]}\lvrrs\Cpcps{v}$.
+\end{lemma}
 
@@
 %Let $E\in\lambda_{cm}$ and $e\in\lambda_{cm}$ be arbitrary.
-By simulation.
+If $E[e]\cmrrs v$, then $E[e]\cmrrs\hole[v]\cmrr v$.
+
+By simulation, $E[e]\cmrrs\hole[v]\implies\Ccps{E[e]}\lvrrs\Ccps{\hole[v]}$.
+
+By definition, $\Ccps{\hole[v]}=\app{\app{\app{\Ccps{v}}{\Ccps{\hole}}}{\xi(\hole)}}{\Cpcps{\chi(\hole)}}\lvrrs\Cpcps{v}$, which is what we sought.
 \end{proof}
@@
 %\newtheorem{maintheorem}{Correctness of $\hat{\mathcal{C}}$}
-\begin{theorem}
-For all programs $p\in\lambda_{cm}$, $\Chcps{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Chcps{p})$.
+\begin{theorem}[Correctness of $\mathcal{C}_{\mathrm{cps}}$]
+For all programs $p\in\lambda_{cm}$, $\Chcps{\mathrm{eval}_{cm}(p)}\equiv\mathrm{eval}_{v}(\Chcps{p})$.
 \end{theorem}
@@
 p\cmrrs v &\implies\hole[p]\cmrrs v &\text{(since $p\cmrr\hole[p]$)}\\
-          &\implies\C{\hole[p]}\lvrrs\Cp{v} &\text{(by corollary \ref{cor:great})}\\
+          &\implies\C{\hole[p]}\lvrrs\Cp{v} &\text{(by lemma \ref{lem:great})}\\
           &\implies\app{\app{\app{\C{p}}{\C{\hole}}}{\xi(\hole)}}{\Cp{\chi(\hole)}}\lvrrs\C{v}&\text{by definition of $\mathcal{C}$}\\

### 2012-11-29 paper.tex
Minor points addressed.

@@
 
-Numerous programming language instruments rely on stack inspection to function.
+Numerous program instruments rely on stack inspection to function.
 Statistical profilers sample the stack regularly to record active functions, algebraic steppers observe the stack to represent the evaluation context of an expression, and debuggers naturally require consistent access to the stack.
@@
 
-To avoid this roadblock, we instead take a macro-style approach; that is, we enhance a core language with facilities to manipulate continuation marks and desugar the enhanced language back to the core.
+To avoid this roadblock, we instead enhance a core language with facilities to manipulate continuation marks and desugar the enhanced language back to the core.
 To make our desugaring transformation portable to other languages, we define it over the \lc, the common core of most higher-order languages.
@@
 \item before evaluation, we expect \scheme'flag' to indicate tail position information and \scheme'marks' to provide a list of the current continuation marks,
-\item we would like to evaluate \scheme'C[rator-expr]' and \scheme'C[rand-expr]' in the same manner, providing to each its contextual information--specifically that neither is evaluated in tail position of a \scheme'wcm' directive and the continuation marks for each are unchanged from the parent context,
-\item and, following evaluation of operator and operand and application, evaluation of the resultant term is performed with the original contextual information.
+\item we would like to evaluate \scheme'C[rator-expr]' and \scheme'C[rand-expr]' in the same manner, providing to each its contextual information--specifically that neither is evaluated in tail position of a \scheme'wcm' directive and the continuation marks for each are unchanged from the parent context, and
+\item following evaluation of operator and operand and application, evaluation of the resultant term is performed with the original contextual information.
 \end{enumerate}
@@
 \item we evaluate \scheme'mark-expr' with correct contextual information,
-\item we discard the first continuation mark of the parent context if evaluation is occurring in tail position of a \scheme'wcm' directive,
-\item and we evaluate \scheme'C[body-expr]' with the correct tail-position flag and current continuation marks.
+\item we discard the first continuation mark of the parent context if evaluation is occurring in tail position of a \scheme'wcm' directive, and
+\item we evaluate \scheme'C[body-expr]' with the correct tail-position flag and current continuation marks.
 \end{enumerate}
@@
 
-The conceptual transformation of variables \scheme'x' and values \scheme'(lambda (x) e)' is straightforward.
+The conceptual transformation of variables \scheme'x' and values \scheme'(lambda (x) e)' is similar.
 
@@
 To achieve \scheme'if' and conditionals as well as list primitives \scheme'cons', \scheme'snd', and \scheme'nil', we use the Church encodings of fig.
-\ref{church-encodings}.
+1.
+%\ref{church-encodings}.
 
@@
 
-Our choice to keep the core language small by omitting lists as primitive values has the consequence of complicating our transform somewhat.
+We have not added lists as primitive values to $\cm$ but instead encode lists with other terms in the language.
+By not relying on a particular characterization of lists in the target language, the dependencies of the transformation remain few, and thus widen its possible application.
+However, this makes lists themselves subject to the transformation which complicates their manifestation in the core language.
+The effect of this is that the transformation must deal with the list of continuation marks at the transformed level.
+
+%Our choice to keep the core language small by omitting lists as primitive values has the consequence of complicating our transform somewhat.
 Because lists are defined in terms of \lc\ values which are themselves touched by the transform and because of the commutativity property that $\mathcal{C}$ must satisfy, we cannot deal with a list of continuation marks directly--we must instead deal with a transformed list of transformed continuation marks, and manipulation of this list within transformed terms must occur at the transformed level.
@@
 The formal transformation of a \scheme'wcm' directive is also extremely similar to the \scheme'let' version.
-The Church-encoded conditional eagerly evaluates both branches, but this still achieves correct behavior.
+The Church-encoded conditional eagerly evaluates both branches, but this still achieves correct behavior, as \scheme'marks' is already a value and \scheme|(SND marks)| is benign, even if \scheme'marks' is \scheme'NIL'.
 \begin{schemedisplay}
@@
 
-Many compilers transform source to continuation-passing style (CPS) in the compilation process.
+Some compilers transform source to continuation-passing style (CPS) in the compilation process.
 The use of CPS as an intermediate language enables a host of optimization analyses and makes the implementation of otherwise-heavyweight features trivial, such as first-class continuations \cite{appel2007compiling}.
@@
 \end{lemma}
-by induction over contexts $E$ and terms $e$. This lemma justifies the evolution of the evaluation context and each rule in the reduction relation except rule \ref{reduction}. We express this with
+by induction over contexts $E$ and terms $e$.
+This lemma justifies the evolution of the evaluation context and each rule in the reduction relation except rule \ref{reduction}.
+We preserve this rule with
 \begin{lemma}[Substitution]
@@
 Monads, introduced by Moggi \cite{moggi1989computational}, are another way to model semantics.
-Their appeal is significant: they allow one to implement a particular model of computation with with a pure computational logic, such as the \lc.
+Their appeal is significant: they allow one to implement a particular model of computation with a pure computational logic, such as the \lc.
 

### 2012-11-30 macro/cm.js
Macro

@@ -1,0 +1,23 @@
+macro ccm {
+	case test => {
+		function( flag, marks ) {
+			return marks;
+		}
+	}
+}
+
+macro wcm {
+	case ({ $mark:Expr } { $body:Expr }) =>
+		function( flag, marks ) {
+			
+
+
+var fac = function( n ) {
+	if( n == 0 ) {
+		console.log( ccm );
+		return 1;
+	}
+	else {
+		return wcm { n }{ n * fac( n - 1 ) }
+	}
+}

### 2012-11-30 macro/test.js

@@ -1,0 +1,53 @@
+/*macro ccm {
+	case () => {
+		function( flag, marks ) {
+			return marks;
+		}
+	}
+}
+
+macro wcm {
+	case ( $mark:Expr, $body:Expr ) => {
+		$body
+	}
+}
+
+macro def {
+	case $name:ident ( $x:ident (,) ... ) { $body:SourceElements } => {
+		function $name( $x (,) ..., flag, marks ) {
+			$body
+		}
+	}
+}
+
+macro fun {
+	case ( $x:ident (,) ... ) { $body:SourceElements } => {
+		function( $x (,) ..., flag, marks ) {
+			$body
+		}
+	}
+}
+*/
+
+macro iffy {
+	case ($body:Program) => {
+		$body
+	}
+}
+
+iffy (function(n){return n;})
+
+/*def fac( n ) {
+	return 1;
+}
+/*
+	if( n == 0 ) {
+		//console.log( ccm() );
+		return 1;
+	}
+	else {
+		return 1;
+		//return wcm { n, n * fac( n - 1 ) }
+	}
+}
+*/

### 2013-04-11 paper.tex
For TFP.

@@
 
-\title{Instant Continuation Marks--Just Add $\lambda$}
+\title{Pure Continuation Marks}
 
@@
 
-We present direct and CPS-like transformations from the call-by-value $\lambda$-calculus augmented with continuation marks to the pure call-by-value $\lambda$-calculus.
+We present a transformation from the call-by-value $\lambda$-calculus augmented with continuation marks to the pure call-by-value $\lambda$-calculus.
 We discuss how such transformations simplify the construction of compilers which treat continuation marks correctly.
 We document how Redex, a domain-specific language for exploring language semantics, aided the discovery of correct transformations.
-We verify the transformations with the proof of a meaning-preservation theorem.
+We offer the sketch proof of a meaning-preservation theorem.
 Finally, we apply the transformation to JavaScript.
@@
 We discuss the necessary properties of a meaningful transformation between these languages.
-We present two transformations of continuation marks--one direct and one CPS-like--to the call-by-value \lc\ in section \ref{sec-transformation}.
-This is followed by randomized testing using a lightweight mechanized model in section \ref{sec-testing} and a proof of a meaning-preservation theorem in section \ref{sec-proof}.
+We present a transformation of continuation marks to the call-by-value \lc\ in section \ref{sec-transformation}.
+This is followed by randomized testing using a lightweight mechanized model in section \ref{sec-testing} and a sketch proof of a meaning-preservation theorem in section \ref{sec-proof}.
 We finally apply it to JavaScript in section \ref{sec-javascript}.
@@
 
-The evaluation of the factorial of 2, expressed by \scheme|(fac 2)|, proceeds as
-\begin{schemedisplay}
-(fac 2)
-(if (zero? 2) ... (wcm 2 (* 2 (fac (- 2 1)))))
-(wcm 2 (* 2 (fac (- 2 1))))
-(wcm 2 (* 2 (fac 1)))
-(wcm 2 (* 2 (if (zero? 1) ... (wcm 1 (* 1 (fac (- 1 1)))))))
-(wcm 2 (* 2 (wcm 1 (* 1 (fac (- 1 1))))))
-(wcm 2 (* 2 (wcm 1 (* 1 (fac 0)))))
-(wcm 2 (* 2 (wcm 1 (* 1 (if (zero? 0) (begin (print (ccm)) 1) ...)))))
-(wcm 2 (* 2 (wcm 1 (* 1 (begin (print (ccm)) 1)))))
-\end{schemedisplay}
-At this point, the term \scheme|(ccm)| is evaluated.
-This happens, in effect, by walking the continuation and building a list of the current continuation marks which appear in the \scheme'mark-value' position of a \scheme'wcm' term.
-In this example, \scheme|(ccm)| evaluates to \scheme|'(1 2)|.
-Evaluation continues as
-\begin{schemedisplay}
-(wcm 2 (* 2 (wcm 1 (* 1 (begin (print '(1 2)) 1)))))
-(wcm 2 (* 2 (wcm 1 (* 1 (begin 1)))))
-(wcm 2 (* 2 (wcm 1 (* 1 1))))
-(wcm 2 (* 2 (wcm 1 1)))
-(wcm 2 (* 2 1))
-(wcm 2 2)
-2
-\end{schemedisplay}
-The visible effect of the call is
+The effect of evaluating the factorial of 2, expressed as \scheme|(fac 2)|, is 
 \begin{verbatim}
@@
 
-The evaluation of the factorial of 2, here expressed by \scheme|(fac 2 1)|, proceeds as
-\begin{schemedisplay}
-(fac 2 1)
-(if (zero? 2) ... (wcm 2 (fac (- 2 1) (* 2 1))))
-(wcm 2 (fac (- 2 1) (* 2 1)))
-(wcm 2 (fac 1 (* 2 1)))
-(wcm 2 (fac 1 2))
-(wcm 2 (if (zero? 1) ... (wcm 1 (fac (- 1 1) (* 1 2)))))
-(wcm 2 (wcm 1 (fac (- 1 1) (* 1 2))))
-\end{schemedisplay}
-At this point, there is one \scheme'wcm' directive directly nested within another.
-Behaviorally, the outer one is collapsed \emph{before} evaluation of the inner body proceeds.
-In essence, it is overwritten.
-Following this, evaluation continues as
-\begin{schemedisplay}
-(wcm 1 (fac (- 1 1) (* 1 2)))
-(wcm 1 (fac 0 (* 1 2)))
-(wcm 1 (fac 0 2))
-(wcm 1 (if (zero? 0) (begin (print (ccm)) 2) ...))
-(wcm 1 (begin (print (ccm)) 2))
-(wcm 1 (begin (print '(1)) 2))
-(wcm 1 (begin 2))
-(wcm 1 2)
-2
-\end{schemedisplay}
-The visible effect of the call is
+The effect of evaluating the factorial of 2, here expressed by \scheme|(fac 2 1)|, is
 \begin{verbatim}
@@
 
-We have not added lists as primitive values to $\cm$ but instead encode lists with other terms in the language.
+We have not added lists as primitive values to \cm but instead encode lists with other terms in the language.
 By not relying on a particular characterization of lists in the target language, the dependencies of the transformation remain few, and thus widen its possible application.
@@
 \begin{schemedisplay}
-(lambda (flags)
+(lambda (flag)
   (lambda (marks)
@@
        ((C[rand-expr] FALSE) marks))
-      flags)
+      flag)
      marks)))
@@
 \end{schemedefn}
-
-\subsection{Definition of $\mathcal{C}$ in CPS}
-
-Some compilers transform source to continuation-passing style (CPS) in the compilation process.
-The use of CPS as an intermediate language enables a host of optimization analyses and makes the implementation of otherwise-heavyweight features trivial, such as first-class continuations \cite{appel2007compiling}.
-
-We present a continuation mark transformation integrated in CPS.
-
-\begin{schemedefn}{\scheme|Ccps[(rator-expr rand-expr)]|}
-\begin{schemedisplay}
-(lambda (kont)
-   (lambda (flag)
-     (lambda (marks)
-       (((Ccps[rator-expr]
-          (lambda (rator-value)
-            (((Ccps[rand-expr]
-               (lambda (rand-value)
-                 ((((rator-value rand-value) kont) flag) marks)))
-              FALSE)
-             marks)))
-         FALSE)
-        marks))))
-\end{schemedisplay}
-\end{schemedefn}
-
-\begin{schemedefn}{\scheme|C[(wcm mark-expr body-expr)]|}
-\begin{schemedisplay}
-(lambda (kont)
-  (lambda (flag)
-    (lambda (marks)
-      (((Ccps[mark-expr]
-          (lambda (mark-value) 
-            ((lambda (rest-marks) 
-               (((Ccps[body-expr] kont) TRUE) Chcps[((CONS mark-value) rest-marks)]))
-             ((flag Chcps[(SND marks)]) marks))))
-        FALSE)
-       marks))))
-\end{schemedisplay}
-\end{schemedefn}
-
-\begin{schemedefn}{\scheme|Ccps[(ccm)]|}
-\begin{schemedisplay}
-(lambda (kont)
-  (lambda (flag)
-    (lambda (marks)
-      (kont marks))))
-\end{schemedisplay}
-\end{schemedefn}
-
-\noindent
-\begin{minipage}{0.5\linewidth}
-\begin{schemedefn}{\scheme|Ccps[v]|=\scheme|Ccps[(lambda (x) e)]|}
-\begin{schemedisplay}
-(lambda (kont)
-  (lambda (flag)
-    (lambda (marks)
-      (kont (lambda (x) Ccps[e])))))
-\end{schemedisplay}
-\end{schemedefn}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-\begin{schemedefn}{\scheme|Ccps[x]|}
-\begin{schemedisplay}
-(lambda (kont)
-  (lambda (flag)
-    (lambda (marks)
-      (kont x))))
-\end{schemedisplay}
-\end{schemedefn}
-\end{minipage}
-\\
-\\
-
-We include corresponding definitions for $\mathcal{C}'$ and $\hat{\mathcal{C}}$.
-\\
-\\
-\noindent
-\begin{minipage}{0.5\linewidth}
-\begin{schemedefn}{\scheme|Cpcps[(lambda (x) e)]|}
-\begin{schemedisplay}
-(lambda (x) Ccps[e])
-\end{schemedisplay}
-\end{schemedefn}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-\begin{schemedefn}{\scheme|Chcps[p]|}
-\begin{schemedisplay}
-(((Ccps[p] (lambda (x) x)) FALSE) Cpcps[nil])
-\end{schemedisplay}
-\end{schemedefn}
-\end{minipage}
 
@@
 
-\scheme'redex-check' generates random terms according to the grammar of the given language (\scheme'lambdacm') and category (\scheme'e') in search of counterexamples to the predicate.
+\scheme'redex-check' generates random terms according to the grammar of the given language (\scheme'lambdacm') and term category (\scheme'e') in search of counterexamples to the predicate.
 It gradually increases the size of the terms it generates, which we found useful in obtaining minimal test cases.
-We subjected both the direct and CPS transformation to random testing and each eventually withstood 10,000 random tests.
+We subjected the transformation to random testing.
+We used the specific counterexamples to guide modifications to the transform until it  withstood 10,000 random tests.
 Interestingly, no incorrect transformation withstood more than 500 random tests before failing.
@@
 
-With a reasonable intuition and definitions hardened by random testing, we prove correctness of the continuation-passing variant of the transform, $\mathcal{C}_{\mathrm{cps}}$.
+With a reasonable intuition and definitions hardened by random testing, we lightly stetch a proof of the correctness of $\mathcal{C}$.
 
-Because of space limitations, we offer only a sketch of the proof here.
+%Because of space limitations, we offer only a sketch of the proof here.
 The full proof can be found in a technical report\footnote{\texttt{http://students.cs.byu.edu/\textasciitilde{}krgerman/esop2013-technical-report.pdf}}.
@@
 \begin{theorem}
-For all programs $p\in\lambda_{cm}$, $\Chcps{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Chcps{p})$.
+For all programs $p\in\lambda_{cm}$, $\Ch{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Ch{p})$.
 \end{theorem}
 
-We prove this by overloading $\mathcal{C}_{\mathrm{cps}}$ to accommodate first evaluation contexts and then context-term pairs (denoted in unorthodox style) by
-\begin{schemedefn}{\scheme|Ccps[E[e]]|}
+We prove this by overloading $\mathcal{C}$ to accommodate first evaluation contexts and then context-term pairs (where $E[e]$ represents $(E,e)$) by
+\begin{schemedefn}{\scheme|C[E[e]]|}
 \begin{schemedisplay}
-(((Ccps[e] Ccps[E]) xiE) Cpcps[chiE])
+C[E][((C[e] xiE) Cp[chiE])]
 \end{schemedisplay}
@@
 \noindent
-which allows us to formally relate $E[e]$ and $\Ccps{E[e]}$. We then prove a lemma
+which allows us to formally relate $E[e]$ and $\C{E[e]}$.
+%For example, $\C{\app{\abs{x}{x}}{\hole}[42]}=\C{\app{\abs{x}{x}}{\hole}}[\app{\app{\C{42}}{\xi(\app{\abs{x}{x}}{\hole})}}{\Cp{\chi(\app{\abs{x}{x}}{\hole})}}]$.
+
+We then prove a lemma
 \begin{lemma}[Simulation]
@@
 
-Our macro-style approach would be more useful if deployed as a proper macro, but this requires a hygienic macro system which many languages lack.
+%Our macro-style approach would be more useful if deployed as a proper macro, but this requires a hygienic macro system which many languages lack.
 \texttt{sweet.js} is just such a system for JavaScript in early stages and is ideal for our transformation.
 
-Although we used some level of mechanization in the proof of the theorem, a formal proof assistant, such as Coq \cite{barras1997coq}, would increase confidence out of the gate that the transformation was correct.
+%Although we used some level of mechanization in the proof of the theorem, a formal proof assistant, such as Coq \cite{barras1997coq}, would increase confidence out of the gate that the transformation was correct.
 
